### 2017-01-21 Thesis/latex/acknowledgements.tex
New git repo

\chapter{Acknowledgements}

First of all I would like to thank Henrik Kalisch and Magnus Sv\"ard who carried out the selection process of candidates for the PhD position. I am sincerely thankful to Professor Henrik Kalisch, my main supervisor, for sharing knowledge and guiding me, for his enthusiasm, support and encouragement. The thesis would not have been possible without you.

In the course of the doctorate studies, I have had fruitful collaboration with Olivier Verdier, Evgueni Dinvay, Denys Dutykh, Ben Segal and others. I would like to thank them for productive discussions, advices on research questions and some inspiring ideas we came up with.

As a PhD Candidate I have been lucky to meet my colleagues and fellow students at the Department of Mathematics. They have always provided a friendly and supportive environment. I appreciate everything we have experienced during the time of our studies.

I would like to express my gratitude to my family for their continuous support and encouragement over the years of my studies. I am also thankful for all of my teachers and professors from schools and universities I studied at.

I am heartily grateful for my wife Assem. You have always been there for me. Without your love, support and understanding I would not have been able to go that far.

### 2017-01-21 Thesis/latex/outline.tex
New git repo

\chapter{Outline}

This thesis is organised in the following way. Part I contains general theoretical background on nonlinear wave models as well as description of methods used to solve the equations involved. Some properties of solutions to the equations and a summary of results are also given in the first part. Part 2 consists of the research papers that present scientific results in detail.

\section*{List of reseach papers included in Part II}

\textbf{Paper A:}

\noindent Moldabayev, D., Kalisch, H., Dutykh, D.: \textit{The Whitham equation as a model for surface water waves}, Phys. D~ \textbf{309}, 99–107 (2015),

\noindent http://dx.doi.org/10.1016/j.physd.2015.07.010.

\vspace{1em}

\noindent \textbf{Paper B:}

\noindent Dinvay, E., Moldabayev, D., Dutykh, D., Kalisch, H.:  \textit{The Whitham equation with surface tension}, Nonlinear Dynamics, 1--14 (2017),

\noindent http://dx.doi.org/10.1007/s11071-016-3299-7.

\vspace{1em}

\noindent \textbf{Paper C:}

\noindent Henrik Kalisch, Daulet Moldabayev, Olivier Verdier: \textit{A numerical study of nonlinear dispersive wave models with SpecTraVVave}, \textcolor{red}{specify status of the paper.}

\vspace{1em}

\noindent \textbf{Paper D:}

\noindent Benjamin Segal, Daulet Moldabayev, Henrik Kalisch, Bernard Deconinck: \textit{Explicit solutions for a long-wave model with constant vorticity}, submitted to European Journal of Mechanics - B/Fluids.

### 2017-01-21 Thesis/latex/papers.tex
New git repo

%\part{Scientific results}

%--------------------------------------------------------------------------------------------------
% Paper A
%--------------------------------------------------------------------------------------------------
\chapter*{Paper A}

\addcontentsline{toc}{chapter}{Paper A The Whitham equation as a model for surface water waves}
\section*{The Whitham equation as a model for surface water waves}

\noindent Daulet Moldabayev, Henrik Kalisch, Denys Dutyh \\

\noindent \textit{Physica} D \textbf{309}, (2015)

\noindent http://dx.doi.org/10.1016/j.physd.2015.07.010
\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/paper-1/page_}{9}

%--------------------------------------------------------------------------------------------------
% Paper B
%--------------------------------------------------------------------------------------------------

\chapter*{Paper B}

\addcontentsline{toc}{chapter}{Paper B The Whitham equation with surface tension}
\section*{The Whitham equation with surface tension}

\noindent Evgueni Dinvay, Daulet Moldabayev, Denys Dutykh, Henrik Kalisch\\

\noindent \textit{Nonlinear Dynamics}, 1--14 (2017)

\noindent http://dx.doi.org/10.1007/s11071-016-3299-7

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/paper-2/page_}{14}

%--------------------------------------------------------------------------------------------------
% Paper C
%--------------------------------------------------------------------------------------------------

\chapter*{Paper C}

\addcontentsline{toc}{chapter}{Paper C A numerical study of nonlinear dispersive wave models with SpecTraVVave}
\section*{A numerical study of nonlinear dispersive wave models with SpecTraVVave}

\noindent Henrik Kalisch, Daulet Moldabayev, Olivier Verdier\\

%\noindent \textit{Physical Review} B, \textbf{76}, 035303 (2007)
\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/paper-3/page_}{27}

%--------------------------------------------------------------------------------------------------
% Paper D
%--------------------------------------------------------------------------------------------------

\chapter*{Paper D}

\addcontentsline{toc}{chapter}{Paper D Explicit solutions for a long-wave model with constant vorticity}
\section*{Explicit solutions for a long-wave model with constant vorticity}

\noindent Benjamin Segal, Daulet Moldabayev, Henrik Kalisch, Bernard Deconinck\\

%\noindent \textit{Physical Review} B, \textbf{76}, 035303 (2007)
\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/paper-4/page_}{19}

### 2017-01-21 Thesis/latex/preface.tex
New git repo

\chapter{Preface}

This dissertation is submitted as a partial fulfillment of the requirements for the degree Doctor of Philosophy (PhD) at the Department of Mathematics, University of Bergen. The research was
supported by the Research Council of Norway on grant no. 213474/F20.

The advisory committee has consisted of Henrik Kalisch (University of Bergen, Bergen, Norway),  Dmitry Pelinovsky (McMaster University, Hamilton, Ontario, Canada) and Angel Dur\'an (Universidad de Valladolid, Valladolid, Spain).

### 2017-01-21 Thesis/latex/scientific_environment.tex
New git repo

\chapter{Scientific environment}

This is where you write the name of the subject/professional milieu involved in the study, faculty(faculties)/institute(s)/centre(s)/research groups/school of research.

Contractual joint venture partners should also be mentioned here. If these are represented by graphic symbols (logo etc.) the it should be on this page.

NB! Most organisations have guidelines concerning the use of their logo. These guidelines regulate who can use the logo, when the logo can be used and how. This is quite relevant when a logo is used together with information connected to other institutions/organisations. (A typical example would be rules concerning space between other written or graphic content.)

Remember to find out which guidelines apply and ensure that they are followed!
Most organisations will be able to provide a digital version of their logo in a suitable format.

### 2017-01-21 Thesis/papers.tex
New git repo

%\chapter*{Scientific results}
%\addcontentsline{toc}{chapter}{Scientific results}

%--------------------------------------------------------------------------------------------------
% Paper I
%--------------------------------------------------------------------------------------------------
\chapter*{Paper A}
\addcontentsline{toc}{chapter}{Paper A - Linear and nonlinear convection in porous media between coaxial cylinders}
\section*{Linear and nonlinear convection in porous media between coaxial cylinders}
%\addcontentsline{toc}{section}{Linear and nonlinear convection in porous media between coaxial cylinders}

\noindent C. Bringedal, I. Berre, J.M. Nordbotten, D.A.S. Rees\\

\noindent \textit{Physics of Fluids} \textbf{23}, 9, (2011). \\ http://dx.doi.org/10.1063/1.3637642
\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
%\newgeometry{hmargin=0.5cm,vmargin={0.5cm}}
\includepaperpages{../papers/Paper1/page_}{11}
%\includepdf[pages={2-12},lastpage=12]{../papers/Paper1}
%\restoregeometry

%\chapter*{Paper B}
%\section{Influence of natural convection in a porous medium when producing from borehole heat exchangers}
\chapter*{Paper B}
\addcontentsline{toc}{chapter}{Paper B - Influence of natural convection in a porous medium when producing from borehole heat exchangers}  
\section*{Influence of natural convection in a porous medium when producing from borehole heat exchangers}
%\addcontentsline{toc}{section}{Influence of natural convection in a porous medium when producing from borehole heat exchangers}

\noindent C. Bringedal, I. Berre, J.M. Nordbotten, \\

\noindent \textit{Water Resources Research} \textbf{49}, 8, (2013). \\ http://dx.doi.org/10.1002/wrcr.20388

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper2/page_}{12}

\chapter*{Paper C}
\addcontentsline{toc}{chapter}{Paper C - An Approach for Investigation of Geochemical Rock-Fluid Interactions}  
\section*{An Approach for Investigation of Geochemical Rock-Fluid Interactions}
%\addcontentsline{toc}{section}{An Approach for Investigation of Geochemical Rock-Fluid Interactions}

\noindent C. Bringedal, I. Berre, F.A. Radu, \\

\noindent \textit{Proceedings, Thirty-Ninth Workshop on Geothermal Reservoir Engineering}, Stanford University, (2014).

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper3/page_}{10}

\chapter*{Paper D}
\addcontentsline{toc}{chapter}{Paper D - A model for non-isothermal flow and mineral precipitation and dissolution in a thin strip}  
\section*{A model for non-isothermal flow and mineral precipitation and dissolution in a thin strip}
%\addcontentsline{toc}{section}{A model for non-isothermal flow and mineral precipitation and dissolution in a thin strip}

\noindent C. Bringedal, I. Berre, I.S. Pop, F.A. Radu, \\

%\noindent \textit{Journal of Computational and Applied Mathematics}, \\http://dx.doi.org/10.1016/j.cam.2014.12.009, (2014).

\noindent \textit{Journal of Computational and Applied Mathematics} \textbf{289}, (2015)\\ http://dx.doi.org/10.1016/j.cam.2014.12.009

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper4/page_}{10}

\chapter*{Paper E}
\addcontentsline{toc}{chapter}{Paper E - Upscaling of non-isothermal reactive porous media flow with changing porosity}  
\section*{Upscaling of non-isothermal reactive porous media flow with changing porosity}
%\addcontentsline{toc}{section}{Upscaling of non-isothermal reactive porous media flow with changing porosity}

\noindent C. Bringedal, I. Berre, I.S. Pop, F.A. Radu, \\

\noindent \textit{Transport in Porous Media} (2015)\\ http://dx.doi.org/10.1007/s11242-015-0530-9
%\noindent \textit{Submitted to Transport in Porous Media}

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper5/page_}{26}

\chapter*{Paper F}
\addcontentsline{toc}{chapter}{Paper F - Upscaling of non-isothermal reactive porous media flow under dominant P\'{e}clet number: the effect of changing porosity}  
\section*{Upscaling of non-isothermal reactive porous media flow under dominant P\'{e}clet number: the effect of changing porosity}
%\addcontentsline{toc}{section}{Upscaling of non-isothermal reactive porous media flow under dominant P\'{e}clet number: the effect of changing porosity}

\noindent C. Bringedal, I. Berre, I.S. Pop, F.A. Radu, \\

\noindent \textit{Submitted}% to SIAM Multiscale Modeling $\&$ Simulations}.

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper6/page_}{30}

### 2017-01-21 Thesis/papers/doc.py

@@ -1,0 +1,122 @@
+import os, shutil, sys
+from numpy import sort
+
+def PrepAllPapers(howMany=1, workdir = "."):
+	"""
+	Prepare paper PDFs for inclusion in thesis as figures: Burst all 
+	papers into one PDF per page, and crop margin whitespace.
+	"""
+	os.chdir(workdir)
+
+	#Iterate over papers, burst and crop
+	map(PrepPaper, range(1,howMany+1))
+
+class PaperDoesNotExistException(Exception):
+	pass
+
+
+def PrepPaper(paperIndex):
+	paper = "paper%i.pdf" % paperIndex
+	if not os.path.exists(paper):
+		raise PaperDoesNotExistException("Paper %i does not exist or is not correctly named, please name it %s and put it in this folder" % (paperIndex, paper))
+
+	print ("Preparing Paper %i" % paperIndex)
+
+	#Remove previous prep
+	dirName = "paper-%i" % (paperIndex)
+	if os.path.exists(dirName):
+		shutil.rmtree(dirName)
+	#create output folder
+	os.mkdir(dirName)
+	
+	#Copy paper PDF to workingdir
+	shutil.copy(paper, "%s/" % dirName) 
+	os.chdir(dirName)
+
+	#Burst it
+	BurstPDF(paper)
+
+	#Crop each page and convert to ps
+	pageFiles = sort([fname for fname in os.listdir(".") if fname[:3] == "pg_"])
+	for pageNum, pfn in enumerate(pageFiles):
+		
+		#Crop
+		croppedPDF = "%s-crop.pdf" % pfn.strip(".pdf")
+		CropPDF(pfn, croppedPDF)
+
+		#Convert to PS
+		psName = "page_%i.ps" % pageNum
+		PDFtoPS(croppedPDF, psName)
+		
+		#Remove tmp files
+		os.remove(pfn)
+		os.remove(croppedPDF)
+
+	#Remove more tmp files
+	os.remove(paper)
+	os.remove("doc_data.txt")
+
+	os.chdir("..")
+
+
+def BurstPDF(pdfName):
+	"""
+	Burst a PDF into separate files for each page
+	"""
+	os.system("pdftk %s burst" % pdfName)
+
+
+def PDFtoPS(infile, outfile):
+	"""
+	Convert a PDF to a PS file
+	"""
+	os.system("pdftops %s %s" % (infile, outfile))
+
+
[…]

### 2017-01-21 Thesis/papers/prep_papers_thesisinclude.py

@@ -1,0 +1,124 @@
+#!/usr/bin/env python
+
+import os, shutil, sys
+from numpy import sort
+
+def PrepAllPapers(howMany=1, workdir = "."):
+	"""
+	Prepare paper PDFs for inclusion in thesis as figures: Burst all 
+	papers into one PDF per page, and crop margin whitespace.
+	"""
+	os.chdir(workdir)
+
+	#Iterate over papers, burst and crop
+	map(PrepPaper, range(1,howMany+1))
+
+class PaperDoesNotExistException(Exception):
+	pass
+
+
+def PrepPaper(paperIndex):
+	paper = "paper%i.pdf" % paperIndex
+	if not os.path.exists(paper):
+		raise PaperDoesNotExistException("Paper %i does not exist or is not correctly named, please name it %s and put it in this folder" % (paperIndex, paper))
+
+	print ("Preparing Paper %i" % paperIndex)
+
+	#Remove previous prep
+	dirName = "paper-%i" % (paperIndex)
+	if os.path.exists(dirName):
+		shutil.rmtree(dirName)
+	#create output folder
+	os.mkdir(dirName)
+	
+	#Copy paper PDF to workingdir
+	shutil.copy(paper, "%s/" % dirName) 
+	os.chdir(dirName)
+
+	#Burst it
+	BurstPDF(paper)
+
+	#Crop each page and convert to ps
+	pageFiles = sort([fname for fname in os.listdir(".") if fname[:3] == "pg_"])
+	for pageNum, pfn in enumerate(pageFiles):
+		
+		#Crop
+		croppedPDF = "%s-crop.pdf" % pfn.strip(".pdf")
+		CropPDF(pfn, croppedPDF)
+
+		#Convert to PS
+		psName = "page_%i.ps" % pageNum
+		PDFtoPS(croppedPDF, psName)
+		
+		#Remove tmp files
+		os.remove(pfn)
+		os.remove(croppedPDF)
+
+	#Remove more tmp files
+	os.remove(paper)
+	os.remove("doc_data.txt")
+
+	os.chdir("..")
+
+
+def BurstPDF(pdfName):
+	"""
+	Burst a PDF into separate files for each page
+	"""
+	os.system("pdftk %s burst" % pdfName)
+
+
+def PDFtoPS(infile, outfile):
+	"""
+	Convert a PDF to a PS file
+	"""
+	os.system("pdftops %s %s" % (infile, outfile))
+
+
[…]

### 2017-01-28 Thesis/latex/summary.tex
Summary - beginning

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Summary of results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This chapter provides an overview of the results achieved in the course of research work. 
Detailed

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Whitham equation as a model for surface water waves}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Whitham equation was proposed as an alternate model equation for the simplified description of unidirectional 
wave motion at the surface of an inviscid fluid. As the Whitham equation incorporates the full
linear dispersion relation of the water wave problem, it is thought to provide a more faithful description
of shorter waves of small amplitude than traditional long wave models such as the KdV equation.
In this work, we identify a scaling regime in which the Whitham equation can be derived from the
Hamiltonian theory of surface water waves. A Hamiltonian system of Whitham type allowing for two-
way wave propagation is also derived. The Whitham equation is integrated numerically, and it is shown
that the equation gives a close approximation of inviscid free surface dynamics as described by the
Euler equations. The performance of the Whitham equation as a model for free surface dynamics is also
compared to different free surface models: the KdV equation, the BBM equation, and the Padé (2,2) model.
It is found that in a wide parameter range of amplitudes and wavelengths, the Whitham equation performs
on par with or better than the three considered models.

In its simplest form, the water-wave problem concerns the
flow of an incompressible inviscid fluid with a free surface over
a horizontal impenetrable bed. In this situation, the fluid flow
is described by the Euler equations with appropriate boundary
conditions, and the dynamics of the free surface are of particular
interest in the solution of this problem.
There are a number of model equations which allow the
approximate description of the evolution of the free surface
without having to provide a complete solution of the fluid flow
below the surface. In the present contribution, interest is focused

on the derivation and evaluation of a nonlocal water-wave model
known as the Whitham equation. The equation is written as

\begin{equation}
\eta_t + \frac{3}{2} \frac{c_0}{h_0} \eta \eta_x + K_{h_0} \ast \eta_x = 0, 
\label{whitham-equation}
\end{equation}

where the convolution kernel $K_{h_0}$ is given in terms of the Fourier
transform by

\begin{equation}
\mathcal{F}K_{h_0} (\xi) =  \sqrt{\frac{g \tanh(h_0\xi)}{\xi}}.
\label{whitham-kernel}
\end{equation}

$g$ is the gravitational acceleration, $h_0$ is the undisturbed depth of
the fluid, and $c_0 = \sqrt{gh_0}$ is the corresponding long-wave speed.
The convolution can be thought of as a Fourier multiplier operator,
and \eqref{whitham-kernel} represents the Fourier symbol of the operator.
The Whitham equation was proposed by Whitham \cite{Whitham1967} as an
alternative to the well known Korteweg–de Vries (KdV) equation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Whitham equation with surface tension}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A numerical study of nonlinear dispersive wave models with SpecTraVVave}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Explicit solutions for a long-wave model with constant vorticity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-01-29 Thesis/latex/summary.tex
Summary: paper A + some bibliography

@@
 \end{equation}
-
 where the convolution kernel $K_{h_0}$ is given in terms of the Fourier
@@
 \end{equation}
-
 $g$ is the gravitational acceleration, $h_0$ is the undisturbed depth of
@@
 alternative to the well known Korteweg–de Vries (KdV) equation
+
+\begin{equation}
+\eta_t + c_0 \eta_x + \frac{3}{2} \frac{c_0}{h_0} \eta \eta_x + \frac{1}{6} c_0 h_0^2 \eta_{xxx} = 0.
+\label{kdv-equation}
+\end{equation}
+
+The validity of the KdV equation as a model for surface water waves
+can be described as follows. Suppose a wave field with a prominent
+amplitude $a$ and characteristic wavelength $l$ is to be studied. The
+KdV equation is known to produce a good approximation of the
+evolution of the waves if the amplitude of the waves is small and
+the wavelength is large when compared to the undisturbed depth,
+and if in addition, the two non-dimensional quantities $a / h_0$ and
+$h_0^2 / l^2$ are of similar size. The latter requirement can be written in
+terms of the Stokes number as
+
+\begin{equation}
+\mathcal{S} = \frac{a l^2}{h_0^3} \sim 1. 
+\label{stockes-number}
+\end{equation}
+While the KdV equation is a good model for surface waves if $\mathcal{S} \sim 1$, 
+one notorious problem with the KdV equation is that it does
+not model accurately the dynamics of shorter waves. Recognizing
+this shortcoming of the KdV equation, Whitham proposed to use
+the same nonlinearity as the KdV equation, but couple it with a
+linear term which mimics the linear dispersion relation of the full
+water-wave problem. Thus, at least in theory, the Whitham
+equation can be expected to yield a description of the dynamics
+of shorter waves which is closer to the solutions of the more
+fundamental Euler equations which govern the flow.
+
+The Whitham equation has been studied from a number of
+vantage points during recent years. In particular, the existence of
+traveling and solitary waves has been established in \citep{Ehrnstrom2012, Ehrnstrom2009}.
+Well
+posedness of a similar equation was investigated in \citetext{\citealp[]{Lannes2013}-\citealp[]{Klein2015}}, and a
+model with variable depth has been studied numerically in \cite{Minzoni2013}.
+Moreover, it has been shown in \citep{Hur2015, Sanford2014} that periodic solutions of
+the Whitham equation feature modulational instability for short
+enough waves in a similar way as small-amplitude periodic wave
+solutions of the water-wave problem. However, even though the
+equation is routinely mentioned in texts on nonlinear waves \citep{Drazin89, Whitham74}, it appears that the performance of the Whitham equation in
+the description of surface water waves has not been investigated
+so far.
+The purpose of the present article is to give an asymptotic
+derivation of the Whitham equation as a model for surface water
+waves, and to confirm Whitham’s expectation that the equation is
+a fair model for the description of time-dependent surface water
+waves. For the purpose of the derivation, we introduce an expo-
+nential scaling regime in which the Whitham equation can be de-
+rived asymptotically from an approximate Hamiltonian principle
+for surface water waves. In order to motivate the use of this scaling,
+note that the KdV equation has the property that wide classes of
+initial data decompose into a number of solitary waves and small-
+amplitude dispersive residue \cite{Ablowitz81}. For the KdV equations, solitary-
+wave solutions are known in closed form, and are given by
+\begin{equation}
+\eta = \frac{a}{h_0}\mbox{sech}^2\left(\sqrt{\frac{3a}{4h_0^3}(x - ct)}\right)
+\label{kdv-soliton}
+\end{equation}
+for a certain wave celerity $c$. These waves clearly comply with
+the amplitude–wavelength relation $a / h_0 \sim h_0^2 / l^2$ which was
+mentioned above. It appears that the Whitham equation – as indeed do many other nonlinear dispersive equations – also has
+the property that broad classes of initial data rapidly decompose
+into ordered trains of solitary waves (see Fig. 1). Quantifying the
+amplitude–wavelength relation for these solitary waves yields an
+asymptotic regime which is expected to be relevant to the validity
+of the Whitham equation as a water wave model.
 

### 2017-01-30 Thesis/latex/summary.tex
thesis.bib got updates.

@@
 Euler equations. The performance of the Whitham equation as a model for free surface dynamics is also
-compared to different free surface models: the KdV equation, the BBM equation, and the Padé (2,2) model.
+compared to different free surface models: the KdV equation, the BBM equation, and the Pad\'e (2,2) model.
 It is found that in a wide parameter range of amplitudes and wavelengths, the Whitham equation performs
@@
 
+As the curve fit in the right panel of Fig. 1 shows, the relationship
+between wavelength and amplitude of the Whitham solitary
+waves can be approximately described by the relation $a/h_0 \sim e^{\kappa(l/h_0)^{\nu}}$ for certain values of $\kappa$ and $\nu$. Since the Whitham solitary
+waves are not known in exact form, the values of $\kappa$ and $\nu$ have to be
+found numerically. Then one may define a Whitham scaling regime
+\begin{equation}
+\mathcal{W}(\kappa, \nu) = \frac{a}{h_0}e^{\kappa(l/h_0)^{\nu}} \sim 1,
+\label{whitham-regime}
+\end{equation}
+and it will be shown in Sections 2 and 3 that this scaling can be used
+advantageously in the derivation of the Whitham equation. The
+derivation proceeds by examining the Hamiltonian formulation of
+the water-wave problem due to Zakharov, Craig and Sulem [13,14],
+and by restricting to wave motion which is predominantly in the
+direction of increasing values of x. The approach is similar to the
+method of [15], but relies on the new relation (5).
+First, in Section 2, a Whitham system is derived which allows for
+two-way propagation of waves. The Whitham equation is found in
+Section 3. Finally, in Section 4, a comparison of modeling properties
+of the KdV and Whitham equations is given. The comparison also
+includes the regularized long-wave equation
+\begin{equation}
+\eta_t + c_0 \eta_x + \frac{3}{2} \frac{c_0}{h_0} \eta \eta_x - \frac{1}{6} h_0^2 \eta_{xxt} = 0,
+\label{regularized-long-wave}
+\end{equation}
+which was put forward in [16] and studied in depth in [17], and
+which is also known as the BBM or PBBM equation. The linearized
+dispersion relation of this equation is not an exact match to the
+dispersion relation of the full water-wave problem, but it is much
+closer than the KdV equation, and it might also be expected that
+this equation may be able to model shorter waves more success-
+fully than the KdV equation. In order to obtain an even better match
+of the linear dispersion relation, one may make use of Pad\'e  expansions. In the context of simplified evolution equations, this
+approach was pioneered in [18]. For uni-directional models, this
+approach was advocated in [19], and in particular, the equation
+based on the Pad\'e (2,2) approximation was studied in depth. In dimensional variables, this model takes the form
+\begin{equation}
+\eta_t + c_0 \eta_x + \frac{3}{2} \frac{c_0}{h_0} \eta \eta_x
+- \frac{3}{20} c_0 h_0^2 \eta_{xxx} - \frac{19}{60} h_0^2 \eta_{xxt} = 0.
+\label{pade-2-2}
+\end{equation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-01-30 Babenko/SpecTraVVave/setup.py

@@ -1,0 +1,24 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from distutils.core import setup
+
+setup(
+	name = 'travwave',
+	description = 'Compute traveling waves',
+	author='Daulet Moldabayev, Henrik Kalisch, Olivier Verdier',
+	license = 'MIT',
+	keywords = ['Math'],
+	
+	packages=['travwave', 'travwave.equations', 'travwave.boundary', 'travwave.dynamic'],
+	classifiers = [
+	'Development Status :: 3 - Alpha',
+	'Environment :: Console',
+	'Intended Audience :: Science/Research',
+	'License :: OSI Approved :: MIT License',
+	'Operating System :: OS Independent',
+	'Programming Language :: Python',
+	'Topic :: Scientific/Engineering :: Mathematics',
+	],
+	zip_safe = False,
+	)

### 2017-01-30 Babenko/SpecTraVVave/test/test_diagram.py

@@ -1,0 +1,81 @@
+from __future__ import division
+
+import unittest
+
+from travwave.diagram import BifurcationDiagram
+import travwave.equations as teq
+import travwave.boundary as tbc
+
+from travwave.discretization import Discretization
+from travwave.solver import Solver
+from travwave.navigation import Navigator
+
+
+import numpy.testing as npt
+
+class TestRefine(unittest.TestCase):
+    def setUp(self):
+        length = 5
+        equation = teq.kdv.KDV(length)
+        boundary_cond = tbc.Const()
+        self.bd = bd = BifurcationDiagram(equation, boundary_cond)
+        bd.initialize()
+        bd.navigation.run(10)
+
+    def test_refine(self):
+        new_size = 500
+        n,v,p = self.bd.navigation.refine_at(new_size)
+        self.assertEqual(len(n), new_size)
+
+    def test_plot_diagram(self):
+        self.bd.plot_diagram()
+
+    def test_plot_solutions(self):
+        self.bd.plot_solutions()
+
+class TestGeneral(unittest.TestCase):
+    """
+    Only tests that the problem can be manually set up and run without raising any exception.
+    No tests are actually performed.
+    """
+    def get_size(self):
+        return 50
+
+    def get_length(self):
+        return 100
+
+    def get_equation_class(self):
+        return teq.kdv.KDV
+
+    def get_boundary(self):
+        return tbc.Minimum(0)
+
+    def get_nbsteps(self):
+        return 10
+
+    def test_manual_initialization(self):
+        size = self.get_size()
+        length = self.get_length()
+        self.equation = self.get_equation_class()(length)
+        self.boundary = self.get_boundary()
+
+        self.diagram = BifurcationDiagram(self.equation, self.boundary, size, size)
+
+        self.discretization = Discretization(self.equation, size)
+        solver = Solver(self.discretization, self.boundary)
+        nb_steps = self.get_nbsteps()
+        nav = Navigator(solver.solve, size=size)
+        initial_velocity = self.discretization.bifurcation_velocity()
+        p1 = (initial_velocity, 0)
+        epsilon = .1/nb_steps
+        p0 = (initial_velocity, -epsilon)
+        initial_guess = self.discretization.compute_initial_guess(epsilon/10)
+        nav.initialize(initial_guess, p1, p0)
+        nav.run(1)
+        self.nav = nav
+        store = nav[-1]
+        self.B = store['integration constant']
[…]

### 2017-01-30 Babenko/SpecTraVVave/test/test_discretization.py

@@ -1,0 +1,19 @@
+import unittest
+import numpy as np
+import numpy.testing as npt
+
+from travwave.equations.kdv import KDV
+from travwave.discretization import DiscretizationOperator
+
+class TestLinear(unittest.TestCase):
+    def test_compare(self):
+        size = 64
+        d = DiscretizationOperator(equation=KDV(np.pi), size=size)
+        op = d.compute_linear_operator()
+        u = np.random.randn(size)
+        # u_ = # same with an initial guess
+        npt.assert_allclose(np.dot(op, u), d.apply_operator(u))
+
+        
+
+

### 2017-01-30 Babenko/SpecTraVVave/test/test_dynamic.py

@@ -1,0 +1,38 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+import unittest
+import numpy.testing as npt
+
+from travwave.equations import *
+from travwave.boundary import *
+from travwave.dynamic import *
+from travwave.diagram import *
+
+
+class TestDynamic(unittest.TestCase):
+    def test_dynamid(self):
+        length = 3*np.pi
+        equation = kdv.KDV(length)
+        boundary_cond = Const()
+        bd = BifurcationDiagram(equation, boundary_cond)
+        bd.initialize()
+        bd.navigation.run(10)
+
+        u = bd.navigation[-1]['solution']
+        velocity = bd.navigation[-1]['current'][bd.navigation.velocity_]
+        dyn = self.get_dynamic_class()(equation, u, velocity)
+        uu = dyn.mirror()
+        t_wave = dyn.evolution(solution = uu, nb_steps=int(1e2), periods = 1)
+
+        error = t_wave - uu
+        print(max(abs(error)))
+        npt.assert_allclose(t_wave, uu, atol=1e-3, err_msg="Wave is equal to itself after travelling one period")
+
+    def get_dynamic_class(self):
+        return Trapezoidal_rule
+
+class TestDefrutos(TestDynamic):
+    def get_dynamic_class(self):
+        return DeFrutos_SanzSerna

### 2017-01-30 Babenko/SpecTraVVave/test/test_equation.py

@@ -1,0 +1,99 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+import unittest
+import numpy.testing as npt
+import numpy as np
+
+from travwave.equations import kdv, whitham, benjamin, kawahara
+from travwave.discretization import DiscretizationOperator
+
+class HarnessEquation(object):
+    def setUp(self):
+        self.equation = self.get_equation()
+        self.discretization = DiscretizationOperator(self.equation, 8)
+
+    def test_flux(self):
+        """
+        Test that flux_prime works
+        """
+        u = .5
+        eps = 1e-5
+        expected = (self.equation.flux(u+eps) - self.equation.flux(u))/eps
+        computed = self.equation.flux_prime(u)
+        npt.assert_allclose(computed, expected, rtol=1e-4)
+
+    def test_degree(self):
+        """
+        The degree is correct.
+        """
+        A = 1e10
+        degree = np.log(self.equation.flux(A)/self.equation.flux(1))/np.log(A)
+        npt.assert_allclose(self.equation.degree(), degree)
+
+    def test_linop(self):
+        """
+        The numba implementation of the linear operator gives the same results as the previous one.
+        """
+        weights = self.discretization.get_weights()
+        f = lambda x: np.cos(np.pi/self.equation.length*x)
+        x = self.discretization.get_nodes()
+        tensor = np.dstack([wk*(f(k*x) * f(k*x).reshape(-1,1)) for k, wk in enumerate(weights)])
+        expected = np.sum(tensor, axis=2)
+        computed = self.discretization.compute_linear_operator()
+        npt.assert_allclose(computed, expected)
+
+    def test_residual(self):
+        """
+        The discretization residual is equal to the linear operator plus the flux.
+        """
+        parameter = (2.,3.)
+        u = np.random.rand(self.discretization.size)
+        expected = np.dot(self.discretization.compute_shifted_operator(self.discretization.size, parameter), u) + self.equation.flux(u)
+        computed = self.discretization.residual(u, parameter, 0)
+        npt.assert_allclose(computed, expected)
+
+class TestKDV(HarnessEquation, unittest.TestCase):
+    def get_equation(self):
+        return kdv.KDV(1)
+
+class TestKDV3(HarnessEquation, unittest.TestCase):
+    def get_equation(self):
[…]

### 2017-01-30 Babenko/SpecTraVVave/test/test_functional.py

@@ -1,0 +1,39 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+import unittest
+
+from travwave.navigation import *
+from travwave.solver import *
+from travwave.equations import *
+from travwave.discretization import Discretization
+from travwave.boundary import *
+
+import numpy as np
+
+class TestGeneral(unittest.TestCase):
+    """
+    Only tests that the problem can be set up and run without raising any exception.
+    No tests are actually performed.
+    """
+    def test_run(self):
+        size = 10
+        length = np.pi
+        equation = kdv.KDV(length)
+        discretization = Discretization(equation, size)
+        boundary = Mean()
+        solver = Solver(discretization, boundary)
+        nav = Navigator(solver.solve)
+        initial_velocity = discretization.bifurcation_velocity()
+        p1 = (initial_velocity, 0)
+        epsilon = .1
+        initial_guess = discretization.compute_initial_guess(epsilon)
+        p0 = (initial_velocity, -epsilon)
+        nav.initialize(initial_guess, p1, p0)
+        nav.run(1)
+        print(nav[1])
+
+if __name__ == '__main__':
+    unittest.main()
+#

### 2017-01-30 Babenko/SpecTraVVave/test/test_kdv.py

@@ -1,0 +1,148 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+import unittest
+
+import numpy as np
+
+from travwave.navigation import Navigator
+from travwave.solver import Solver
+from travwave.equations import kdv
+from travwave.discretization import Discretization
+from travwave.boundary import Minimum, Mean, Const
+
+import numpy.testing as npt
+
+import matplotlib.pyplot as plt
+
+def sech(x):
+    return 1/np.cosh(x)
+
+def soliton(amplitude, x):
+    s = sech(np.sqrt(3*amplitude)/2*x)
+    return amplitude*s*s
+
+def identity(xs, us):
+    return us[1:-1]
+
+def derivative(xs, us):
+    du = us[2:] - us[:-2]
+    dx = xs[2:] - xs[:-2]
+    return du/dx
+
+def derivative2(xs, us):
+    ddu = us[2:] - 2*us[1:-1] + us[:-2]
+    dx = xs[1]-xs[0]
+    return ddu/(dx*dx)
+
+class TestKDV(unittest.TestCase):
+
+    def get_length(self):
+        return 10
+
+    def get_nbsteps(self):
+        return 1
+
+    def get_size(self):
+        return 128
+
+    def setUp(self):
+        size = self.get_size()
+        length = self.get_length()
+        self.equation = self.get_equation_class()(length)
+        self.discretization = Discretization(self.equation, size)
+        self.boundary = self.get_boundary()
+        solver = Solver(self.discretization, self.boundary)
+        nb_steps = self.get_nbsteps()
+        nav = Navigator(solver.solve, size=size)
+        initial_velocity = self.discretization.bifurcation_velocity()
+        p1 = (initial_velocity, 0)
+        epsilon = .1/nb_steps
+        p0 = (initial_velocity, -epsilon)
+        initial_guess = self.discretization.compute_initial_guess(epsilon/10)
+        nav.initialize(initial_guess, p1, p0)
+        nav.run(1)
+        self.nav = nav
+        store = nav[-1]
+        self.B = store['integration constant']
+        self.c = store['current'][nav.velocity_]
+        self.amplitude = store['current'][nav.amplitude_]
+        self.xs = self.discretization.get_nodes()
+        self.computed = store['solution']
+
+    def get_residual_tolerance(self):
+        return 1e-6
+
+    def test_residual(self):
+        """
+        Discretized residual is below tolerance when applied to the computed solution.
+        """
[…]

### 2017-01-30 Babenko/SpecTraVVave/test/test_navigation.py

@@ -1,0 +1,67 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+import unittest
+import numpy.testing as npt
+import numpy as np
+
+import itertools
+
+from travwave.navigation import *
+
+def dummy_solver(x, pstar, direction):
+    return x + 1., 0, pstar
+
+
+class TestDummyNavigation(unittest.TestCase):
+    def test_ortho_direction(self):
+        p0 = (0,0)
+        p1 = (1,0)
+        dp = (p1[0]-p0[0],p1[1]-p0[1])
+        computed = ortho_direction(p0, p1, step=1)
+        expected_pstar = (2, 0)
+        computed_pstar = computed[0]
+        self.assertIsInstance(computed_pstar, tuple, msg="stored parameter is a tuple")
+        npt.assert_allclose(computed_pstar, expected_pstar, err_msg="computed parameters are right")
+        expected_direction = (0,1)
+        computed_direction = computed[1]
+        self.assertIsInstance(computed_direction, tuple, msg="direction is a tuple")
+        self.assertEqual(computed_direction[0]*dp[0] + computed_direction[1]*dp[1], 0, msg="direction is orthogonal to the line joining the two previous parameter points")
+
+    def setUp(self):
+        low_size = 24
+        self.nav = Navigator(dummy_solver, size=low_size)
+        init_size = 50
+        x0 = np.ones(init_size)
+        self.nav.initialize(x0, (1.,0), (0.,0))
+
+    def test_refine(self):
+        sol, v, p_ = self.nav.refine(resampling=20, sol=np.ones(5), p=(1,2), direction=(0,1))
+        self.assertEqual(len(sol), 20)
+
+    def test_nav(self):
+        N = 20
+        nav = self.nav
+        nav.run(N)
+        points = [a['solution'] for a in nav]
+        self.assertEqual(len(nav), len(points), msg="Possible to compute the length of a Navigation object")
+        self.assertEqual(len(nav), N+1) # Total size is number of steps plus one
+        self.assertEqual(len(nav[0]['solution']), nav.size) # Right size of the initialized navigation point
+        ## npt.assert_allclose(np.array(points), np.arange(N+1) + x0)
+        for i in range(N):
+            p2 = nav[i]['current']
+            p1_ = nav[i+1]['previous']
+            if len(nav[i]['solution']) == len(nav[i+1]['solution']):
+                self.assertEqual(p2, p1_) # parameter is properly passed at the next stage
+
[…]

### 2017-01-30 Babenko/SpecTraVVave/test/test_solver.py

@@ -1,0 +1,20 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+import unittest
+import numpy.testing as npt
+
+from travwave.solver import *
+
+@unittest.skip("The Jacobian is not implemented")
+class TestJacobianExtension(unittest.TestCase):
+    def test_extension(self):
+        jacobian = np.zeros((3,3))
+        N = len(jacobian)
+        ortho = (1.,2.)
+        extended = compute_extended_jacobian(jacobian, ortho)
+        self.assertEqual(extended.shape, (N+2,N+2))
+        last_rows = extended[-2:]
+        ## last_columns = extended[:, -2:]
+        npt.assert_allclose(last_rows[:, -2:], np.array([[-1, 0], [ortho[1], ortho[0]]]))

### 2017-01-30 Babenko/SpecTraVVave/travwave/__init__.py

@@ -1,0 +1,5 @@
+from . import equations
+from . import boundary
+from . import dynamic
+
+from .diagram import BifurcationDiagram

### 2017-01-30 Babenko/SpecTraVVave/travwave/diagram.py

@@ -1,0 +1,45 @@
+from __future__ import division
+
+from travwave import navigation, solver, discretization
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+class BifurcationDiagram(object):
+
+    def __init__(self, equation, boundary_condition, size=32, init_size=256):
+        self.discretization = discretization.Discretization(equation, init_size)
+        solve = solver.Solver(self.discretization, boundary=boundary_condition)
+        self.navigation = navigation.Navigator(solve.solve, size=size)
+
+    def initialize(self, amplitude=0.01, step=0.005):
+        initial_guess = self.discretization.compute_initial_guess(amplitude)
+        initial_velocity = self.discretization.bifurcation_velocity()
+        p1 = (initial_velocity, 0)
+        p0 = (initial_velocity, -step)
+        self.navigation.initialize(initial_guess, p1, p0)
+
+    def plot_data(self): # prints out all (velocity, amplitude) vectors for computed solutions
+        parameters = [result['current'] for result in self.navigation]
+        aparameters = np.array(parameters)
+        return aparameters.T
+
+    def plot_diagram(self):
+        aparameters = self.plot_data()
+        plt.plot(aparameters[0], aparameters[1], '.--')
+        plt.xlabel('Wavespeed')
+        plt.ylabel('Waveheight')
+
+    def plot_solution(self, solution):
+        size = len(solution)
+        self.discretization.size = size
+        nodes = self.discretization.get_nodes()
+        plt.plot(nodes, solution)
+        plt.xlabel('x')
+        plt.ylabel('Surface Elevation')
+
+    def plot_solutions(self, index = [-1]):
+        counter = np.arange(np.size(index))
+        for i in counter:
+            solution = self.navigation[index[i]]['solution']
+            self.plot_solution(solution)

### 2017-01-30 Babenko/SpecTraVVave/travwave/discretization.py

@@ -1,0 +1,106 @@
+from __future__ import division
+
+import numpy as np
+import numba
+
+import scipy.fftpack
+
+def get_nodes(size, length=1.): # returns vector x \in R^N, 
+    return length*(np.linspace(0, 1, size, endpoint=False) + 1/2/size)
+
+def resample(wave, new_size):   # this function is used for navigation with different grid numbers
+                                # takes in wave with fewer points and returns that wave with more points
+    size = len(wave)
+    new_nodes = get_nodes(new_size)
+    resampled = np.interp(new_nodes, get_nodes(size), wave)
+    return resampled
+
+
+class Discretization(object):
+    def __init__(self, equation, size): # 'equation' has class Equation (e.g. KDV)
+                                        # 'size' is the number of discrete points, i.e. N
+        self.equation = equation
+        self.size = size
+
+    def residual(self, u, parameters, integrconst): # parameters \in R^2, parameters = (c,a)
+                                                    # u \in R^{size}, u = wave profile, i.e. u = phi
+                                                    # integrconst \in R, integrconst = B
+        residual = self.apply_operator(u) - parameters[0]*u + self.equation.flux(u) - integrconst
+        return  residual
+
+    def frequencies(self): # returns a vector of frequencies, k \in R^N
+        return np.pi/self.equation.length*np.arange(self.size, dtype=float)
+
+    def image(self): # returns a vector of operator's kernel applied to frequencies
+                     # \alpha(k) \in R^N
+        return self.equation.compute_kernel(self.frequencies())
+
+    def bifurcation_velocity(self): # computes velocity approximation c_0
+        return self.image()[1] # check this
+
+    def get_nodes(self): # returns a vector with discrete (grid) points x_n
+        nodes = get_nodes(self.size, self.equation.length)
+        return nodes
+
+    def compute_initial_guess(self, amplitude): # returns initial guess for u as a cosine function
+                                                # amplitude is a small positive number 
+        cosine = np.cos(get_nodes(self.size, np.pi))
+        init_guess = amplitude*cosine
+        return init_guess
+    """
+    def get_weights(self):  # this methods is not used now.
+        image = self.image()
+        weights = image*2/(len(image))
+        weights[0] /= 2
+        return weights
+    """    
[…]

### 2017-01-30 Babenko/SpecTraVVave/travwave/dynamic/trapezoidal.py

@@ -1,0 +1,88 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+import numpy as np
+from scipy.fftpack import fft, ifft, dct
+
+class Trapezoidal_rule(object):
+    """
+    Dynamic integrator based on trapezoidal rule, 2nd order precision in dt (Li, Sattinger, 1998).
+    """
+    def __init__(self, equation, wave, velocity):
+        self.equation = equation
+        self.u = wave
+        self.velocity = velocity
+
+    def shift_frequencies(self, NN):
+        """
+        Compute the Fourier multiplier of the operator complying with the fft conventions
+        """
+        scale = self.equation.length/np.pi
+        # todo: use fft shift and freq instead
+        centered_frequencies = 1/scale * np.concatenate((np.arange(1, NN/2+1, 1), np.arange(1-NN/2, 0, 1)))
+        kerne = 1j * centered_frequencies * self.equation.compute_kernel(centered_frequencies)
+        kernel = np.concatenate(([0], kerne))
+        kernel[NN//2] = 0
+
+        shifted_frequencies = 1/scale * np.concatenate((np.arange(0, NN/2, 1), [0], np.arange(1-NN/2, 0, 1)))
+        return kernel, shifted_frequencies
+
+    def mirror(self):
+        """
+        Mirrors the half-period profile to a full period profile.
+        """
+        u = self.u
+        uu = np.hstack([u[::-1],u])
+        return uu
+
+    def evolution(self, solution, nb_steps=1000, periods = 1):
+        """
+        The main body of the integrator's code. Takes full wave profile as input. Returns the result of integration
+        of the given equation with input as the initial value. Integrates over the given number of time periods
+        with given time-step dt.
+        """
+        u = solution
+        NN = len(u)
+
+        kernel, shifted_frequencies = self.shift_frequencies(NN)
+
+        T = 2 * self.equation.length / self.velocity
+        dt = periods * T / nb_steps
+
+        m = 0.5 * dt * kernel
+        d1 = (1-m)/(1+m)
+        d1[NN/2] = 0
+        d2 = -0.5 * 1j * dt * shifted_frequencies/(1+m)
+
+        eps = 1e-10
+
+        for i in range(nb_steps):
+            fftu = fft(u)
+            fftuu = fft(self.equation.flux(u))
+
+            z1 = d1*fftu + d2*fftuu
[…]

### 2017-01-30 Babenko/SpecTraVVave/travwave/equations/__init__.py

@@ -1,0 +1,4 @@
+from .base import Equation
+from . import kdv
+from . import whitham
+from . import benjamin

### 2017-01-30 Babenko/SpecTraVVave/travwave/equations/base.py

@@ -1,0 +1,13 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+class Equation(object):
+    def __init__(self, length):
+        self.length = length
+
+    def compute_kernel(self, k):
+        raise NotImplementedError()
+
+    def flux(self, u):
+        raise NotImplementedError()

### 2017-01-30 Babenko/SpecTraVVave/travwave/equations/benjamin.py

@@ -1,0 +1,33 @@
+from __future__ import division
+
+from .base import Equation
+import numpy as np
+
+class Benjamin_Ono(Equation):
+    """
+    The equation is :     -c*u + u + 1/2*u^2 + H(u_x)=0
+    """
+    def degree(self):
+        return 2
+
+    def compute_kernel(self, k):        
+        return  1.0 - np.abs(k)
+            
+    def flux(self, u):
+        return 0.5*u*u
+
+    def flux_prime(self, u):
+        return u
+
+class modified_Benjamin_Ono(Benjamin_Ono):
+    """
+    The equation is :     -c*u + u + 1/3*u^3 + H(u_x)=0
+    """
+    def degree(self):
+        return 3
+
+    def flux(self, u):
+        return 1/3*u*u*u
+
+    def flux_prime(self, u):
+        return u*u

### 2017-01-30 Babenko/SpecTraVVave/travwave/equations/kawahara.py

@@ -1,0 +1,23 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+from .base import Equation
+
+class Kawahara (Equation):
+    """
+    The equation is :     -c*u + 3/4u^2 + (u - 1/2*kp*u" + 1/90*u'''')=0
+    """
+    def degree(self):
+        return 2
+
+    def compute_kernel(self, k):
+        kp = 1
+        return 1.0+0.5*kp*k**2 + 1.0/90*k**4
+            
+    def flux(self, u):
+        return 0.75*u**2  
+
+    def flux_prime(self, u):
+        return 1.5*u
+

### 2017-01-30 Babenko/SpecTraVVave/travwave/equations/whitham.py

@@ -1,0 +1,80 @@
+#!/usr/bin/env python
+# coding: utf-8
+from __future__ import division
+
+from .base import Equation
+import numpy as np
+
+class Whitham(Equation):
+    def degree(self):
+        return 2    
+
+    def compute_kernel(self,k):
+        whitham = np.zeros(len(k))
+        for i in range(len(k)):        
+            if k[i] == 0:          #It is incorrect))
+                whitham[i] = 1
+            else:    
+                whitham[i]  = np.sqrt(np.tanh(k[i])/k[i])        
+        return whitham
+        
+    def flux(self, u):
+        return 0.75*u*u  
+
+    def flux_prime(self, u):
+        return 1.5*u
+
+class Whitham_scaled(Whitham):
+    def compute_kernel(self,k):
+        scale = self.length/np.pi
+        if k[0] == 0:
+            k1 = k[1:]
+            whitham = np.sqrt(scale) * np.concatenate( ([1], np.sqrt( np.tanh( k1/scale )/k1 )))
+        else:    
+            whitham  = np.sqrt(scale) * np.sqrt(1./k*np.tanh(1/scale * k))
+       
+        return whitham    
+
+class Whitham3(Whitham):
+    def degree(self):
+        return 3  
+        
+    def flux(self, u):
+        return 0.5*u**3  
+
+    def flux_prime(self, u):
+        return 1.5*u**2
+        
+class Whitham5(Whitham):
+    def degree(self):
+        return 5
+          
+    def flux(self, u):
+        return 0.5*u**5  
+
+    def flux_prime(self, u):
+        return 2.5*u**4
+        
+class Whithamsqrt (Whitham):
+    def degree(self):
+        return 1.5
+ 
+    def flux(self, u):
+        return 2*np.power(u+1, 1.5) - 3*u - 2
+
+    def flux_prime(self, u):
+        return 3*(u+1)**(0.5)-3
+    
+class WhithamIce (Whitham):
+     def compute_kernel(self,k):
+        kappa = 0.1
+        whitham = np.zeros(len(k))
+        for i in range(len(k)):        
+            if k[i] == 0:          #It is incorrect))
+                whitham[i] = 1
+            else:    
+                whitham[i]  = np.sqrt(np.tanh(k[i])/k[i] * ( 1+kappa* np.power(k[i],4) ) )        
+        return whitham
+ 
+   
+

### 2017-01-30 Babenko/SpecTraVVave/travwave/navigation.py

@@ -1,0 +1,106 @@
+from __future__ import division
+
+from .discretization import resample
+
+def ortho_direction(p1, p2, step):   # takes in bifurcation points p1, p2 and step_size
+                                     # computes point 'p3' and orthogonal direction vector 'd'
+    """
+    Returns pstar such that
+        p3 = p2 + step*(p2-p1)
+    """
+    dp = (p2[0] - p1[0], p2[1] - p1[1])
+    # orthogonal direction
+    direction = (-dp[1], dp[0])
+
+    p3 = (p2[0] + dp[0]*step, p2[1] + dp[1]*step)
+    return p3, direction
+
+class Navigator(object):
+    """
+    Runs the Newton iterator and stores the computed solutions.
+    """
+
+    def __init__(self, solve, size=32):  # 'solve' is the 'solve'-method of object of class Solver
+        """
+        solve: solve function
+        size: size for the navigation
+        """
+        self.solve = solve   # Solver object
+        self.size = size     # number of grid points that is inherent from Discretization object
+
+    # the indices for velocity and amplitude
+    velocity_, amplitude_ = (0, 1)
+
+    def __getitem__(self, index):    # internal routine for accessing stored solutions by index number
+        return self._stored_values[index]
+
+    def __len__(self):    # returns the number of stored solutions
+        return len(self._stored_values)
+
+    def initialize (self, current, p, p0): # 'current' = initial guess for wave
+                                           # 'p' = bifurcation point (c0, a=0), 'p0' = (c0, a = -epsilon)
+                                           # 'epsilon' is the step_size to move up on bifurcation curve
+        """
+        Creates a list for solutions and stores the first solution (initial guess).
+        """
+        self._stored_values = []
+        variables = [0]     # integration constant B
+        self._stored_values.append({'solution': resample(current, self.size), 'integration constant': variables, 'current': p, 'previous': p0 }) # the first stored solution is the set of initial guess parameters, i.e. (wave, B, p, p0)
+
+    def compute_direction(self, p1, p2, step):
+        """
+        Strategy for a new parameter direction search.
+        """
+        return ortho_direction(p1, p2, step)
+
+    def run(self, N):
+        """
+        Iterates the solver N times, navigating over the bifurcation branch and storing found solutions.
+        """
+        for i in range(N):
+            self.step() # runs Newton iterator one time
+
+    def prepare_step(self, step, p2, p1): # computes p3 and direction for new iteration of Newton solver
+        """
+        Return the necessary variables to run the solver.
+        """
[…]

### 2017-01-30 Babenko/SpecTraVVave/travwave/solver.py

@@ -1,0 +1,59 @@
+from __future__ import division
+
+import numpy as np
+import newton
+
+def compute_parameter(parameters, direction, extent):
+    """
+    Computes gamma_0 + theta * direction.
+    """
+    return (parameters[0] + extent*direction[0],
+            parameters[1] + extent*direction[1])
+
+class Solver(object):
+    def __init__(self, discretization, boundary):
+        self.discretization = discretization
+        self.boundary = boundary
+
+    def construct(self, wave, variables, extent):
+        """
+        Attaches the solution wave and extent parameter together for further computation.
+        """
+        return np.hstack([wave, variables, extent])
+
+    def destruct(self, vector):
+        """
+        Separates the solution wave and the extent parameter.
+        """
+        n = self.boundary.variables_num()
+        wave = vector[:-(1+n)]
+        variables = vector[-(1+n):-1]
+        extent = vector[-1]
+        return wave, variables, extent
+
+    def solve(self, guess_wave, parameter_anchor, direction):
+        """
+        Runs a Newton solver on a system of nonlinear equations once. Takes the residual(vector) as the system to solve.
+        """
+        size = len(guess_wave)
+        self.discretization.size = size
+
+        def residual(vector):
+            """
+            Contructs a system of nonlinear equations. First part, main_residual, is from given wave equation;
+            second part, boundary_residual, comes from the chosen boundary conditions.
+            """
+            wave, variables, extent = self.destruct(vector)
+            parameter = compute_parameter(parameter_anchor, direction, extent)
+            boundary_residual = self.boundary.enforce(wave, variables, parameter)
+            amplitude_residual = np.array([parameter[1] - wave[0] + wave[-1]])
+            main_residual = self.discretization.residual(wave, parameter, variables)
+            return np.hstack([main_residual, boundary_residual, amplitude_residual])
+
+        guess = self.construct(guess_wave, variables = np.zeros(self.boundary.variables_num()), extent = 0)
+        nsolver = newton.MultipleSolver(residual)
+        computed = nsolver.run(guess)
+        wave, variables, extent = self.destruct(computed)
+        new_parameter = compute_parameter(parameter_anchor, direction, extent)
+
+        return wave, variables, new_parameter

### 2017-01-30 Thesis/latex/paperA.tex
paperA - continued.

--- 2017-01-30 Thesis/latex/summary.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Summary of results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This chapter provides an overview of the results achieved in the course of research work. 
-Detailed
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\subsection{Introduction.}
 The Whitham equation was proposed as an alternate model equation for the simplified description of unidirectional 
@@
 derivation proceeds by examining the Hamiltonian formulation of
-the water-wave problem due to Zakharov, Craig and Sulem [13,14],
+the water-wave problem due to Zakharov, Craig and Sulem \citep{Craig1993, Zakharov1968},
 and by restricting to wave motion which is predominantly in the
 direction of increasing values of x. The approach is similar to the
-method of [15], but relies on the new relation (5).
+method of \cite{Craig1994}, but relies on the new relation (5).
 First, in Section 2, a Whitham system is derived which allows for
@@
 \end{equation}
-which was put forward in [16] and studied in depth in [17], and
+which was put forward in \cite{Peregrine1966} and studied in depth in \cite{Benjamin1972}, and
 which is also known as the BBM or PBBM equation. The linearized
@@
 of the linear dispersion relation, one may make use of Pad\'e  expansions. In the context of simplified evolution equations, this
-approach was pioneered in [18]. For uni-directional models, this
-approach was advocated in [19], and in particular, the equation
+approach was pioneered in \cite{Witting1984}. For uni-directional models, this
+approach was advocated in \cite{Fetecau2005}, and in particular, the equation
 based on the Pad\'e (2,2) approximation was studied in depth. In dimensional variables, this model takes the form
@@
 \end{equation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Whitham equation with surface tension}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The dispersion relations for the KdV, BBM and Pad\'e (2,2) models are respectively
+\begin{align*}
+c(k) & = c_0 - \frac{1}{6} c_0 h_0^2 k^2 \qquad \mbox{(KdV)}, \\
+c(k) & = c_0 \frac{1}{1 + \frac{1}{6} h_0^2 k^2} \qquad ~\mbox{(BBM)}, \\
+c(k) & = c_0 \frac{1 + \frac{3}{20} h_0^2 k^2}{1 + \frac{19}{60} h_0^2 k^2} \qquad \mbox{(Pad\'e (2,2))}. \\
+\label{dispersive-relations}
+\end{align*}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A numerical study of nonlinear dispersive wave models with SpecTraVVave}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+These approximate dispersion relations are compared to the full
+dispersion relation in Fig. 2. It appears clearly that the Pad\'e (2,2)
+approximation remains much closer to the full dispersion relation than the dispersion relations based on either the linear KdV
+or linear BBM equations. As will be seen in Section 4, solutions of
+both the Whitham and Pad\'e (2,2) equations give closer approximations to solutions of the full Euler equations than either the KdV or
+BBM equations in most cases investigated. However, the Whitham
+equation still keeps a slight edge over the Pad\'e model.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Explicit solutions for a long-wave model with constant vorticity}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Derivation of evolution systems of Whitham type.}
+
+The surface water-wave problem is generally described by the
+Euler equations with slip conditions at the bottom, and kinematic
+and dynamic boundary conditions at the free surface. Assuming
+weak transverse effects, the unknowns are the surface elevation
+$\eta ( x , t )$, the horizontal and vertical fluid velocities $u_1 ( x , z , t )$
+and $u_2 ( x , z , t )$, respectively, and the pressure $P_( x , z , t )$. If the
+assumption of irrotational flow is made, then a velocity potential
+$\phi( x , z , t )$ can be used. In order to nondimensionalize the problem,
+the undisturbed depth $h_0$ is taken as a unit of distance, and the
+parameter $\sqrt{h_0 / g}$ as a unit of time. For the remainder of this article,
+all variables appearing in the water-wave problem are considered
+as being non-dimensional. The problem is posed on a domain
+$\lbrace( x , z )^T \in R^2 ~|~ 1 < z < \eta( x , t )\rbrace$ which extends to infinity in the
+positive and negative $x$-direction. Due to the incompressibility of
+the fluid, the potential then satisfies Laplace’s equation in this
+domain. The fact that the fluid cannot penetrate the bottom is
+expressed by a homogeneous Neumann boundary condition at the
+flat bottom. Thus we have
+%
+\begin{align*}
+	\phi_{xx} + \phi_{zz} = 0 ~ &\mbox{in} ~ -1< z < \eta(x,t) \\
+	\phi_{zz} = 0 ~ &\mbox{on} ~ z = -1.
+\end{align*}
+%
+The pressure is eliminated with the help of the Bernoulli equation,
+and the free-surface boundary conditions are formulated in terms
+of the potential and the surface excursion by
+%
+\begin{equation*}
+	\left.
+		\begin{array}{rc}
+			\eta_t+\phi_x\eta_x-\phi_z
+			& =0,
+			\\
+			\phi_t+\frac{1}{2} \left( \phi^2_x+\phi^2_z \right) + \eta
+			& = 0, 
+		\end{array}
+	\right\}
+	\mbox{on} \ z=\eta(x,t).
+\end{equation*}
+%
+The total energy of the system is given by the sum of kinetic energy and potential energy, and normalized such that the potential
+energy is zero when no wave motion is present at the surface. Accordingly the Hamiltonian function for this problem is
+%
+\begin{equation}
+	H = \int _{\mathbb{R}} \int_0^\eta z \, dz dx +
+	\int _{\mathbb{R}} \int_{-1}^\eta \frac{1}{2} |\nabla \phi|^2 \, dz dx.
+\end{equation}
+%
+Defining the trace of the potential at the free surface as $\Phi( x , t ) = \phi( x , \eta( x , t ), t )$, one may integrate in $z$ in the first integral and use
+the divergence theorem on the second integral in order to arrive at
+the formulation
+%
+\begin{equation}
+	H  = \frac{1}{2} \int_{\mathbb{R}} \left[ \eta^2 + \Phi G(\eta) \Phi \right] \, dx.
+\end{equation}
+%
+This is the Hamiltonian formulation of the water wave problem as
+found in \citep{Craig1993, Petrov1964, Zakharov1968}, and written in terms of the Dirichlet–Neumann
+operator $G (\eta)$ . As shown in \cite{Nicholls2001}, the Dirichlet–Neumann operator
+can be expanded in a series of the form
+%
+\[
+	G(\eta)\Phi = \sum_{j=0}^\infty G_j(\eta) \Phi
+	.
+\]
+%
+In order to proceed, we need to understand the first few terms in
+this series. As shown in \citep{Craig1994, Craig1993}, the first two terms in this series
+can be written with the help of the operator $D = -i \partial_x$ as
+%
+\[
+	G_0(\eta)= D\tanh(D)
+	, \qquad 
+	G_1(\eta)= D\eta D - D\tanh(D) \eta D\tanh(D)
+	. 
+\]
+%
+Note that it can be shown that the terms $G_j (\eta)$ for $j \geq 2$ are of
+quadratic or higher-order in $\eta$, and will therefore not be needed in
+the current analysis.
+
+It will be convenient for the present purpose to formulate the
+Hamiltonian in terms of the dependent variable $u = \Phi_x$. To this
+end, we define the operator $\mathcal{K} (\eta)$ by
+%
+\begin{equation*}
+	G(\eta) = D \mathcal{K}(\eta) D
+	.
+\end{equation*}
+%
+As was the case with $G (\eta)$, the operator $\mathcal{K}(\eta)$ can be expanded in
+a Taylor series around zero as
+%
+\[
+	\mathcal{K}(\eta) = \sum_{j=0}^\infty \mathcal{K}_j(\eta)
+	, \qquad
+	\mathcal{K}_j(\eta) = D^{-1}G_j(\eta)D^{-1}
+	.
+\]
+%
+In particular, note that $\mathcal{K}_0 = \frac{\tanh D}{D}$.
+In non-dimensional variables, we write the operator with the integral kernel $K_{h_0}$ as $K = \sqrt{\frac{\tanh D}{D}}$, so that we have $\mathcal{K}_0 = K^2$.
+The Hamiltonian is then expressed as
+%
+\begin{equation}
+	H  = \frac{1}{2} \int_{\mathbb{R}} \left[ \eta^2 + u \mathcal{K}(\eta) u \right] \, dx.
+\label{scaled-Hamiltonian}
+\end{equation}
+%
+The water-wave problem can then be written as a Hamiltonian
+system using the variational derivatives of $H$ and posing the
+Hamiltonian equations
+%
+\begin{equation}
+	\eta_t = - \partial_x \frac{\delta H}{\delta u}, \qquad u_t = -\partial_x \frac{\delta H}{\delta \eta}.
+\end{equation}
+%
+This system is not in canonical form as the associated structure
+map $J_{\eta, u}$ is symmetric:
+%
+\begin{equation}
+	J_{\eta, u} =
+	\begin{pmatrix}
+		0 & -\partial_x
+		\\
+		-\partial_x & 0
+	\end{pmatrix}.
+\end{equation}
+%
+We now proceed to derive a system of equations which is similar to
+the Whitham equation \eqref{whitham-equation}, but allows bi-directional wave propagation. This system will be a stepping stone on the way to a derivation
+of \eqref{whitham-equation}, but may also be of independent interest. Consider a wave-
+field having a characteristic wavelength $l$ and a characteristic amplitude $a$. Taking into account the nondimensionalization, the two
+scalar parameters $\lambda = l / h_0$ and $ \alpha = a / h_0$ appear. In order to introduce the long-wave and small amplitude approximation into the
+non-dimensional problem, we use the scaling $\tilde{x}= \frac{1}{\lambda} x$, and $ \eta = \alpha \tilde{\eta}$.
+This induces the transformation $\tilde{D} = \lambda D = -\lambda i \partial_x $. If the energy
+is nondimensionalized in accord with the nondimensionalization
+mentioned earlier, then the natural scaling for the Hamiltonian is
+$ \tilde{H}= \alpha^2 H$. In addition, the unknown $u$ is scaled as $u = \alpha \tilde{u}$. The
+scaled Hamiltonian \eqref{scaled-Hamiltonian} is then written as

### 2017-01-30 Thesis/latex/summary.tex
paperA - continued.

--- 2017-01-28 Thesis/latex/summary.tex
@@
 Detailed 
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Whitham equation as a model for surface water waves}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The Whitham equation was proposed as an alternate model equation for the simplified description of unidirectional 
-wave motion at the surface of an inviscid fluid. As the Whitham equation incorporates the full
-linear dispersion relation of the water wave problem, it is thought to provide a more faithful description
-of shorter waves of small amplitude than traditional long wave models such as the KdV equation.
-In this work, we identify a scaling regime in which the Whitham equation can be derived from the
-Hamiltonian theory of surface water waves. A Hamiltonian system of Whitham type allowing for two-
-way wave propagation is also derived. The Whitham equation is integrated numerically, and it is shown
-that the equation gives a close approximation of inviscid free surface dynamics as described by the
-Euler equations. The performance of the Whitham equation as a model for free surface dynamics is also
-compared to different free surface models: the KdV equation, the BBM equation, and the Padé (2,2) model.
-It is found that in a wide parameter range of amplitudes and wavelengths, the Whitham equation performs
-on par with or better than the three considered models.
-
-In its simplest form, the water-wave problem concerns the
-flow of an incompressible inviscid fluid with a free surface over
-a horizontal impenetrable bed. In this situation, the fluid flow
-is described by the Euler equations with appropriate boundary
-conditions, and the dynamics of the free surface are of particular
-interest in the solution of this problem.
-There are a number of model equations which allow the
-approximate description of the evolution of the free surface
-without having to provide a complete solution of the fluid flow
-below the surface. In the present contribution, interest is focused
-
-on the derivation and evaluation of a nonlocal water-wave model
-known as the Whitham equation. The equation is written as
-
-\begin{equation}
-\eta_t + \frac{3}{2} \frac{c_0}{h_0} \eta \eta_x + K_{h_0} \ast \eta_x = 0, 
-\label{whitham-equation}
-\end{equation}
-
-where the convolution kernel $K_{h_0}$ is given in terms of the Fourier
-transform by
-
-\begin{equation}
-\mathcal{F}K_{h_0} (\xi) =  \sqrt{\frac{g \tanh(h_0\xi)}{\xi}}.
-\label{whitham-kernel}
-\end{equation}
-
-$g$ is the gravitational acceleration, $h_0$ is the undisturbed depth of
-the fluid, and $c_0 = \sqrt{gh_0}$ is the corresponding long-wave speed.
-The convolution can be thought of as a Fourier multiplier operator,
-and \eqref{whitham-kernel} represents the Fourier symbol of the operator.
-The Whitham equation was proposed by Whitham \cite{Whitham1967} as an
-alternative to the well known Korteweg–de Vries (KdV) equation
+\input{paperA}
 

### 2017-02-03 Thesis/latex/paperA.tex
TBD: paperB, paperC, bibliography

@@
 interest in the solution of this problem.
+
 There are a number of model equations which allow the
@@
 below the surface. In the present contribution, interest is focused
-
 on the derivation and evaluation of a nonlocal water-wave model
 known as the Whitham equation. The equation is written as
-
+%
 \begin{equation}
@@
 \end{equation}
+%
 where the convolution kernel $K_{h_0}$ is given in terms of the Fourier
 transform by
-
+%
 \begin{equation}
@@
 \end{equation}
+%
 $g$ is the gravitational acceleration, $h_0$ is the undisturbed depth of
@@
 and \eqref{whitham-kernel} represents the Fourier symbol of the operator.
+
 The Whitham equation was proposed by Whitham \cite{Whitham1967} as an
 alternative to the well known Korteweg–de Vries (KdV) equation
-
+%
 \begin{equation}
@@
 \end{equation}
-
+%
 The validity of the KdV equation as a model for surface water waves
@@
 terms of the Stokes number as
-
+%
 \begin{equation}
@@
 \end{equation}
+%
 While the KdV equation is a good model for surface waves if $\mathcal{S} \sim 1$, 
@@
 traveling and solitary waves has been established in \citep{Ehrnstrom2012, Ehrnstrom2009}.
-Well
-posedness of a similar equation was investigated in \citetext{\citealp[]{Lannes2013}-\citealp[]{Klein2015}}, and a
+Well posedness of a similar equation was investigated in \citetext{\citealp[]{Lannes2013}-\citealp[]{Klein2015}}, and a
 model with variable depth has been studied numerically in \cite{Minzoni2013}.
@@
 so far.
+
 The purpose of the present article is to give an asymptotic
@@
 a fair model for the description of time-dependent surface water
-waves. For the purpose of the derivation, we introduce an expo-
-nential scaling regime in which the Whitham equation can be de-
-rived asymptotically from an approximate Hamiltonian principle
+waves. For the purpose of the derivation, we introduce an exponential scaling regime in which the Whitham equation can be derived asymptotically from an approximate Hamiltonian principle
 for surface water waves. In order to motivate the use of this scaling,
@@
 wave solutions are known in closed form, and are given by
+%
 \begin{equation}
@@
 \end{equation}
+%
 for a certain wave celerity $c$. These waves clearly comply with
@@
 found numerically. Then one may define a Whitham scaling regime
+%
 \begin{equation}
@@
 \end{equation}
+%
 and it will be shown in Sections 2 and 3 that this scaling can be used
@@
 direction of increasing values of x. The approach is similar to the
-method of \cite{Craig1994}, but relies on the new relation (5).
+method of \cite{Craig1994}, but relies on the new relation \eqref{whitham-regime}.
+
 First, in Section 2, a Whitham system is derived which allows for
@@
 includes the regularized long-wave equation
+%
 \begin{equation}
@@
 \end{equation}
+%
 which was put forward in \cite{Peregrine1966} and studied in depth in \cite{Benjamin1972}, and
@@
 based on the Pad\'e (2,2) approximation was studied in depth. In dimensional variables, this model takes the form
+%
 \begin{equation}
@@
 \end{equation}
+%
 The dispersion relations for the KdV, BBM and Pad\'e (2,2) models are respectively
+%
 \begin{align*}
@@
 	c(k) & = c_0 \frac{1}{1 + \frac{1}{6} h_0^2 k^2} \qquad ~\mbox{(BBM)}, \\
-c(k) & = c_0 \frac{1 + \frac{3}{20} h_0^2 k^2}{1 + \frac{19}{60} h_0^2 k^2} \qquad \mbox{(Pad\'e (2,2))}. \\
+	c(k) & = c_0 \frac{1 + \frac{3}{20} h_0^2 k^2}{1 + \frac{19}{60} h_0^2 k^2} \qquad 			\mbox{(Pad\'e (2,2))}.
 	\label{dispersive-relations}
 \end{align*}
-
+%
 These approximate dispersion relations are compared to the full
@@
 scaled Hamiltonian \eqref{scaled-Hamiltonian} is then written as
+%
+\begin{multline*}
+	\tilde{H} = \frac{1}{2} \int_{\R} \teta^2 \, dx
+ + \frac{1}{2} \int_{\R} \tilde{u} \left[ 1 - \Sfrac{1}{3} \lambda^{-2} \tilde{D}^2 + \cdots \right] \tilde{u} \, dx  
+ + \frac{\alpha}{2} \int_{\R}  \teta  \tilde{u}^2 \, dx
+\\
+ - \frac{\alpha}{2} \int_{\R}  \tilde{u} \left[ \lambda^{-1} \tilde{D} - \Sfrac{1}{3} \lambda^{-3}\tilde{D}^3 +  \cdots \right] \teta 
+   \left[ \lambda^{-1} \tilde{D} - \Sfrac{1}{3} \lambda^{-3}\tilde{D}^3 + \cdots \right] \tilde{u} \, dx.
+\end{multline*}
+%
+Let us now introduce the small parameter $\mu= \frac{1}{\lambda}$,
+and assume for simplicity that $\alpha = e^{- \kappa / \mu^\nu}$, which corresponds to the
+case where $\mathcal{W}(\kappa,\nu) = 1$. 
+Then the Hamiltonian can be written in the following form:
+%
+\begin{multline*}
+\tH = \ohf \int_{\R} \teta^2 \, dx
+        + \ohf \int_{\R} \tu \left[ 1 - \Sfrac{1}{3} \mu^{2} \tD^2 + \cdots \right] \tu \, dx  
++ \frac{e^{-\kappa/\mu^\nu}}{2} \int_{\R}  \teta  \tu^2 \, dx
+\\
+- \frac{e^{-\kappa/\mu^\nu}}{2} \int_{\R}  \tu \left[ \mu \tD - \Sfrac{1}{3} \mu^{3}\tD^3 +  \dots\right]
+                      \teta \left[ \mu \tD - \Sfrac{1}{3} \mu^{3}\tD^3 + \cdots \right] \tu \, dx.
+\end{multline*}
+%
+Disregarding terms of order $\scrO(\mu^2 e^{-\kappa/\mu^\nu})$, but not of order
+$\scrO(e^{\sca})$ yields the expansion
+\begin{equation}
+\tH = \ohf \int_{\R} \teta^2 \, dx 
+    + \ohf \int_{\R} \tu \left[ 1 - \Sfrac{1}{3} \mu^{2} \tD^2 + \dots \right] \tu \, dx  
+    + \frac{e^{\sca}}{2} \int_{\R}  \teta  \tu^2 \, dx.   
+\end{equation}
+%
+Note that by taking $\mu$ small enough, an arbitrary number of terms of algebraic order 
+in $\mu$ may be kept in the asymptotic series, 
+so that the truncated version of the Hamiltonian in dimensional 
+variables may be written as
+%
+\begin{equation}\label{Hamiltonian-eta-u-truncated}
+	H   = \ohf \int_\R \left[ \eta^2 + u \mathcal{K}_0^N (\eta) u + u \eta u \right] \, dx dz.
+\end{equation}
+%
+However, the difference between $\mathcal{K}_0$ and $\mathcal{K}_0^N$
+is below the order of approximation, so that it is possible
+to formally define the truncated Hamiltonian with
+$\mathcal{K}_0$ instead of $\mathcal{K}_0^N$.
+%
+Hence, the Whitham system is obtained from the 
+Hamiltonian \eqref{Hamiltonian-eta-u-truncated} as follows:
+%
+\begin{align}
+	\label{sys1}
+	\eta_t = - \partial_x \frac{\delta H}{\delta u} 
+       &= -  \mathcal{K}_0 u_x - (\eta u)_x, \\
+	\label{sys2}
+	u_t = -\partial_x \frac{\delta H}{\delta \eta}
+       &= - \eta_x - u u_x.
+\end{align}
+%
+One may also derive a higher-order equation by keeping terms of order 
+$\scrO(\mu^2 e^{\sca})$, but discarding terms of order $\scrO(\mu^4 e^{\sca})$.
+In this case we find the system
+%
+\begin{align*}
+	\eta_t & = -  \mathcal{K}_0 u_x - (\eta u)_x - (\eta u_x)_{xx}, \\
+	u_t    & = - \eta_x - u u_x + u_x u_{xx}.
+\end{align*}
+%
+% *************************************************************************************
+\subsection{Derivation of evolution equations of Whitham type.}
+% *************************************************************************************
+%
+In order to derive the Whitham equation for uni-directional wave propagation,
+it is important to understand how solutions of the Whitham system \eqref{sys1}-\eqref{sys2}
+can be restricted to either left or right-going waves.
+As it will turn out, if  $\eta$ and $u$ are such that $\eta = K u$, 
+then this pair of functions represents a solution
+of \eqref{sys1}-\eqref{sys2} which is propagating to the right. 
+%
+% *************************************************************************************
+Indeed, let us analyze the relation between $\eta$ and $u$ in the linearized Whitham system
+\begin{align}
+\label{linsys1}
+\eta_t & = -  \mathcal{K}_0 u_x, \\
+\label{linsys2}
+u_t    & = - \eta_x.
+\end{align}
+%
+Considering a solution of the system \eqref{linsys1}-\eqref{linsys2} in the form
+\begin{equation}
+\eta(x,t) = A e^{(i\xi x - i \omega t)}, \qquad u(x,t) = B e^{(i\xi x - i \omega t)}.
+\end{equation}
+%
+\noindent 
+gives rise to the matrix equation
+\begin{equation}
+\begin{pmatrix}
+-\omega & \frac{\tanh{\xi}}{\xi} \xi \\
+\xi  & -\omega
+\end{pmatrix}
+\begin{pmatrix}
+A \\
+B
+\end{pmatrix} =
+\begin{pmatrix}
+0 \\
+0
+\end{pmatrix}.
+\label{matrix-equation}
+\end{equation}
+%
+If existence of a nontrivial solution of this system is to be guaranteed,
+the determinant of the matrix has to be zero, so that we have
+$\omega^2 - \frac{\tanh{\xi}}{\xi} \xi^2 = 0$. Defining the phase speed
+as $c = \omega / \xi$, we obtain the dispersion relation
+\begin{equation}
+c = \pm \sqrt{{\textstyle \frac{\tanh{\xi}}{\xi}}}.
+\end{equation}
+%
+%
+The choice of $c > 0$ corresponds to right-going wave solutions of the system 
+\eqref{linsys1}-\eqref{linsys2},
+and the relation between $\eta$ and $u$ can be deduced from \eqref{linsys2}.
+%
+%
+Accordingly, it is expedient to separate solutions
+into a right-going part $r$ and a left-going part $s$
+which are defined by
+\begin{align*}
+r = \ohf (\eta + K  u), \qquad
+s = \ohf (\eta - K  u). 
+\end{align*}
+According to the transformation theory detailed in \cite{Craig2005}, 
+if the unknowns $r$ and $s$ are used instead of $\eta$ and $u$,
+the structure map changes to 
+\begin{equation}
+J_{r,s} = \left( \frac{\partial F}{\partial (\eta, u)} \right) J_{\eta, u}  
+\left( \frac{\partial F}{\partial (\eta, u)} \right)^T
+= 
+\begin{pmatrix}
+-\ohf \partial_x K & 0\\
+0 & \ohf \partial_x K
+\end{pmatrix}.
+\end{equation}
+%
+\noindent 
+We now use the same scaling for both dependent and independent variables as before.
+Thus we have $r = \alpha\tr$ and $s = \alpha\ts$. 
+%
+%
+%
+% *************************************************************************************
+%
+The Hamiltonian is written in terms of $\tr$ and $\ts$ as
+\begin{multline*}
+\tH = \ohf \int_{\R} (\tr+\ts)^2 \, dx
+\\
+ + \ohf \int_{\R} \tK^{-1}(\tr-\ts) \left[ 1 - \Sfrac{1}{3} \mu^2 \tD^2 + \cdots \right] \tK^{-1}(\tr-\ts) \, dx
+%\\
+ + \frac{\alpha}{2} \int_{\R}  (\tr+\ts) \left(\tK^{-1}(\tr-\ts)\right)^2 \, dx
+\\
+ - \frac{\alpha}{2} \int_{\R}  \tK^{-1}(\tr-\ts) \left[ \mu \tD - \Sfrac{1}{3} \mu^3\tD^3 +  \cdots \right] (\tr+\ts)
+   \left[ \mu \tD - \Sfrac{1}{3} \mu^3\tD^3 + \cdots \right] \tK^{-1}(\tr-\ts) \, dx. 
+\end{multline*}
+%
+%
+%
+% *************************************************************************************
+%
+%
+Following the transformation rules, the structure map transforms to
+$J_{\tr,\ts} = 1/\ta^2 J_{r,s}$. 
+In addition, the time scaling $t = \lambda \ttt$ is employed. 
+Since the focus is on right-going solutions, the equation to be considered is
+\begin{equation}
+\lambda \tr_{\ttt} 
+= -\frac{1}{2\ta^2}\lambda\partial_{\tx} \tK \left[ \frac{\delta \big( \ta^2 \tH \big)}{\delta \tr} \right].
+\end{equation}
+So far, this equation is exact. If we now assume that 
+%the waves are predominantly right-going, and 
+$s$ is of the order of
+$\scrO(\mu^2 e^{\sca})$, then the equation for $\tr$ is
+\begin{multline*}
+\tr_{\ttt} = -\frac{1}{2}\partial_{\tx} \big[ 1 - \Sfrac{1}{6} \mu^2 \tD^2 + \cdots \big] 
+           \bigg\{ 2 \tr + \frac{\alpha}{2} \left( \big[1 + \Sfrac{1}{6} \mu^2 \tD^2 + \cdots \big] 
+           \tr \right)^2 
+%\right.
+\\
++ \alpha \big[ 1 + \Sfrac{1}{6} \mu^2 \tD^2 + \cdots \big] 
+   \left( \tr \, \big[1 + \Sfrac{1}{6} \mu^2 \tD^2 + \cdots \big]  \tr  \right)  
+- \frac{\alpha}{2} \left( \big[ \mu \tD - \Sfrac{1}{3} \mu^3 \tD^3 + \cdots \big] 
+    \big[ 1 + \Sfrac{1}{6} \mu^2 \tD^2 + \cdots \big]  \tr  \right)^2
+\\
+%\left. 
+- \alpha \big[ \mu \tD - \Sfrac{1}{3} \mu^3 \tD^3 + \cdots \big] \big[1 + \Sfrac{1}{6} \mu^2 \tD^2 + \cdots \big] 
+ \left( \tr \big[ \mu \tD - \Sfrac{1}{3} \mu^3 \tD^3 + \cdots \big] [1 + \Sfrac{1}{6} \mu^2 \tD^2 
+    + \cdots \big] \tr \right) \bigg\} + \scrO(\alpha \mu^2).
+\end{multline*}
+As in the case of the Whitham system, we use $\ta = \scrO(e^{- \kappa / \mu^\nu})$,
+and disregard terms of order $\scrO(\mu^2 e^{\sca})$, but not of order $\scrO(e^{\sca})$.
+This procedure yields the Whitham equation \eqref{whitham-equation} 
+which is written in nondimensional variables as
+\begin{equation*}
+r_t = -K r_x - \frac{3}{2} r r_x.
+\end{equation*}
+As was the case for the system found in the previous section, it is also 
+possible here to include terms of order $\scrO(\mu^2 e^{-\kappa/\mu^\nu})$, resulting
+in the higher-order equation
+\begin{equation*}
+r_t = -K r_x - \frac{3}{2} r r_x - \frac{13}{12} r_x r_{xx} - \frac{5}{12} r r_{xxx}.
+\end{equation*}
+%
+%
+% *************************************************************************************
+\subsection{Numerical results.}
+% *************************************************************************************
+%
+In this section, the performance of the Whitham
+equation as a model for surface water waves 
+is compared to the KdV equation \eqref{kdv-equation}, the BBM equation \eqref{regularized-long-wave},
+and the Pad\'e (2,2) equation \eqref{pade-2-2}.
+For this purpose initial data are imposed, the Whitham, KdV, BBM, and Pad\'e equations are solved numerically,
+and the solutions are compared to numerical solutions of the full Euler equations
+with free-surface boundary conditions. We continue to work in normalized
+variables, such as stated in the beginning of Section 2.
+
+The numerical treatment of the three model equations is 
+by a standard pseudo-spectral scheme,
+such as explained in \cite{Ehrnstrom2013,Fornberg1978} for example. For the time stepping,
+an efficient fourth-order implicit method developed in \cite{DeFrutos1992} is used. 
+The numerical treatment of the free-surface problem for the Euler equations
+is based on a conformal mapping of the fluid domain into a rectangle.
+In the time-dependent case, this method has roots in the work of Ovsyannikov \cite{Ovsyannikov1974}, 
+and was later used in \cite{Dyachenko1996} and \cite{Li2004}. 
+The particular method used for the numerical experiments reported here 
+is a pseudo-spectral scheme which is detailed in \cite{Mitsotakis2014}.
+
+Initial conditions for the Euler equations are chosen in such a way that 
+the solutions are expected to be right moving. This is achieved by
+posing an initial surface disturbance $\eta_0(x)$ together with the
+trace of the potential $\Phi(x) = \int_0^x \eta_0(x')\, dx'.$
+In order to normalize the data, we choose $\eta_0(x)$ in such a way that
+the average of $\eta_0(x)$ over the computational domain is zero.
+The experiments are performed %with two different initial functions and 
+with several different amplitudes $\alpha$ and wavelengths $\lambda$
+(for the purpose of this section, we define the 
+wavelength $\lambda$ as the distance between 
+the two points $x_1$ and $x_2$ at which $\eta_0(x_1) = \eta_0(x_2) = \alpha/2$).
+Both positive and negative initial disturbances are considered.
+While disturbances with positive main part have been studied widely, 
+an initial wave of depression is somewhat more exotic, but nevertheless important,
+as shown for instance in \cite{Hammack1974}.
+A summary of the experiments' settings is given in Table 1.
+Experiments run with an initial wave of elevation are
+labeled as {\it positive},
+and experiments run with an initial wave of depression are
+labeled as {\it negative}.
+The domain for the computations is $-L\leq x \leq L$, with $L = 50$. 
+The function initial data in the {\it positive} cases is given by
+\begin{equation}
+\eta_0(x) =  \alpha \sech^2(f(\lambda)x)-C, \label{test_1}
+\end{equation}
+where 
+\begin{equation*}
+f(\lambda) = \frac{2}{\lambda}\log\left({\textstyle \frac{1+\sqrt{1/2}}{\sqrt{1/2}}}\right), 
+%\end{equation}
+\ \mbox{ and } \
+%\begin{equation}
+%\label{C_const},
+C = \frac{1}{L}\frac{\alpha}{f(\lambda)}\tanh \left({\textstyle \frac{L}{f(\lambda)}}\right).
+\end{equation*}
+and $C$ and $f(\lambda)$ are chosen so that 
+$\int_{-L}^{L}\eta_0(x)dx = 0$, and
+the wavelength $\lambda$ is the distance between the 
+two points $x_1$ and $x_2$ at which $\eta_0(x_1) = \eta_0(x_2) = a/2$.
+The velocity potential in this case is given by
+\begin{equation}
+\label{velocity1}
+\Phi(x) = \frac{\alpha}{f(\lambda)}\tanh(f(\lambda)x)-Cx. 
+\end{equation}
+In the {\it negative} case, the initial data are given by
+\begin{equation*}
+%\label{test_2}
+\eta_0(x) = - \alpha \sech^2(f(\lambda)x) + C. 
+\end{equation*}
+The definitions for $f(\lambda)$ and $C$ are the same, and the velocity potential is
+\begin{equation*}
+%\label{velocity2}
+\Phi(x) = -\frac{\alpha}{f(\lambda)}\tanh(f(\lambda)x)+Cx. 
+\end{equation*}
+%
+\begin{table}
+\begin{center}
+	\begin{tabular}{| c | c | c | c |}
+	\hline
+	Experiment & Stokes number & $\alpha$ & $\lambda$ \\
+	\hline
+	A	& 0.2  &  	 0.1 &	$\sqrt{2}$ 	\\	
+	B	& 0.2  & 	 0.2 &	1 			\\
+	C	& 1 	   & 	 0.1 &	$\sqrt{10}$	\\
+	D	& 1    & 	 0.2 &	$\sqrt{5}$	\\	
+	E	& 5    & 	 0.1 &	$\sqrt{50}$ 	\\
+	F	& 5    & 	 0.2 &	5			\\
+	\hline
+	\end{tabular}
+\end{center}
+\caption{Summary of the Stokes number, nondimensional amplitude and
+nondimensional wavelength of the initial data used in the numerical
+experiments.}
+\end{table}
+%
+In figures Figs. 3 and 4, the time evolution of a wave with an initial narrow peak
+and one with an initial narrow depression at the center is shown. 
+%
+The amplitude is $\alpha=0.2$, and the wavelength is $\lambda = \sqrt{5}$. 
+In Fig. 3 , the time evolution
+according to the Euler, Whitham, KdV and BBM equations are shown,
+and in Fig. 4, the time evolution
+according to the Euler, Whitham, and Pad\'e (2,2) equations are shown.
+
+It appears in Fig. 3 
+that the KdV equation produces a significant number of spurious oscillations,
+the BBM equation also produces a fair number of spurious oscillations, and
+the Whitham equation produces fewer small oscillations than Euler equations.
+Moreover, while the highest peak in the upper panels in Fig. 3 is
+underpredicted by the KdV and BBM equation, the Whitham equation
+produces peaks which are slightly too high.
+In the case of an initial depression, the Whitham equation also
+produces some peaks which are too high, but on the other hand,
+the KdV and the BBM equations introduce phase errors in the
+main oscillations. 
+As is visible in Fig. 4,
+the Pad\'e (2,2) equation reproduces the leading wave fairly accurately,
+but overpredicts the trailing oscillations in the case of a positive disturbance,
+and underpredicts the trailing oscillations in the case of a negative initial disturbance.
+Nevertheless, since the Pad\'e (2,2) does not introduce a phase error, the overall performance
+is better than that of the KdV and BBM equations.
+
+In order to compare the performance of the four approximate equations
+in a more quantitative manner, the discrepancies between solutions of the model equations
+and the prediction due to solving the Euler equations are measured in an integral norm.
+In the center right panels of Figs. 5 and 6, 
+the computations from Figs. 3 and 4 are
+summarized by plotting the normalized $L^2$-error between the KdV, BBM, Pad\'e and Whitham,
+respectively, and the Euler solutions as a function of non-dimensional time.
+Using this quantitative measure of comparison, it appears that the Whitham
+equation gives the best overall rendition of the free surface dynamics
+predicted by the Euler equations.
+
+In the center left panels of Figs. 5 and 6, 
+a similar computation with $\mathcal{S} = 1$, but smaller amplitude is analyzed. 
+Also in these cases, it appears that the Whitham
+equation gives a good approximation to the corresponding Euler solutions, and in
+particular, a much better approximation than either the KdV or the BBM equation.
+The Pad\'e equation also does better than both KdV and BBM equations, but not better
+than the Whitham equation.
+
+Figs. 5 and 6 show comparisons in several other cases of both positive
+and negative initial amplitude, 
+and Stokes numbers $\mathcal{S} = 0.2$, $\mathcal{S} = 1$ and $\mathcal{S} = 5$.
+In most cases, the normalized $L^2$-error between the Whitham and Euler solutions is similar
+or smaller than the errors between the Euler solutions and the other three model equations.
+However, the Pad\'e equation generally outperforms both the KdV and the BBM equation by some measure.
+
+The only case in this study in which the KdV, BBM and Pad\'e equations outperform
+the Whitham equation is in the case of very long positive initial disturbances.
+The case when $\mathcal{S} = 5$ is shown in the lower panels of Fig.5. 
+However, even in this case, the Whitham equation yields approximations of the
+Euler solutions which are similar or better than in the case of smaller wavelengths.
+In addition, in the case of negative initial data, the performance
+of the Whitham equation is on par with the KdV, BBM and Pad\'e equations 
+in the case when $\mathcal{S} = 5$ (lower panels of Fig.5).
+
+%
+%
+% *************************************************************************************
+\subsection{Conclusion.}
+% *************************************************************************************
+%
+In this article, the Whitham equation \eqref{whitham-equation}
+has been studied as an approximate model equation for wave motion
+at the surface of a perfect fluid. Numerical integration of the
+equation suggests that broad classes of initial data decompose
+into individual solitary waves. The wavelength-amplitude ratio
+of these approximate solitary waves has been studied, and it 
+was found that this ratio can be described by an exponential
+relation of the form
+$\frac{a}{h_0} \sim e^{-\kappa(l/h_0)^\nu}$.
+Using this scaling in the Hamiltonian formulation of the water-wave problem,
+a system of evolution equations has been derived which contains the exact
+dispersion relation of the water-wave problem in its linear part.
+Restricting to one-way propagation, the Whitham equation emerged
+as a model which combines the usual quadratic nonlinearity 
+with one branch of the exact dispersion relation of the water-wave problem.
+The performance of the Whitham equation in the approximation of
+solutions of the Euler equations free-surface boundary conditions
+was analyzed, and compared to the performance of the KdV and BBM equations,
+and to the Pad\'e (2,2) model.
+It was found that the Whitham equation gives a more faithful representation 
+of the solutions of the Euler equations than either the KdV or the BBM equations, 
+except in the case of very long waves with positive main part. In this last case,
+the KdV and BBM equations have the upper hand over the Whitham equation.
+The Pad\'e (2,2) model also outperforms the KdV and BBM equations,
+but does not quite as well as the Whitham equation for shorter waves and negative
+disturbances. However, in the case of very long waves with positive main part,
+the Pad\'e model stays on par with the the KdV and BBM equations.

### 2017-02-03 Thesis/latex/paperB.tex
TBD: paperB, paperC, bibliography

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Whitham equation with surface tension}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Introduction.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\setcounter{equation}{0}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In its simplest form, the water-wave problem concerns
the flow of an incompressible inviscid
fluid with a free surface over a horizontal impenetrable bed.
In this situation, the fluid flow 
is described by the Euler equations with appropriate boundary conditions,
and the dynamics of the free surface are of particular interest
in the solutions of this problem.

There are a number of model equations which
allow the approximate description of the evolution of the free surface
without having to provide a complete solution of the fluid flow
below the surface.
%
In the present contribution, interest is focused on the derivation and evaluation of
a non-local water-wave model known as the Whitham equation.
We extend here the results of the work \cite{Moldabayev2015}
to the case where surface tension is taken into account.
The model equation under study is written as
%
\begin{equation}
\label{dimWhitham}
%	\eta_t+\frac{3}{2}\frac{c_0}{h_0}\eta\eta_x+K_{h_0} \! \ast \eta_{x}=0
	\eta_t + W \eta_x + \frac 32 \eta \eta_x = 0
	,
\end{equation}
%
where the convolution kernel of the operator
\(
	W \eta_x = w(-i \partial _x) \eta_x
	= \big( {\mathcal F}^{-1} w \big) * \eta_x
\)	
is given in terms of the Fourier transform by
%
\begin{equation}
\label{symbol}
%	\mathcal{F} \big( K_{h_0} \big) =
%	{\textstyle{ \sqrt{\frac{g\tanh (h_0\xi)}{\xi}}}}
	w(\xi) = \sqrt{ ( 1 + \varkappa \xi^2 ) \Sfrac{ \tanh(\xi) }{ \xi }}.
\end{equation}
%
Here it is assumed that the variables are suitably normalized so that
the gravitational acceleration, the undisturbed depth of the fluid
and the density are all unity.
The surface fluid tension is included here by means of the
capillarity parameter $\varkappa$ which is the inverse of the Bond number.
The convolution can be thought of as a Fourier multiplier operator,
and \eqref{symbol} represents the Fourier symbol of the operator. 
It is also convenient from the analytical point of view to
regard the Whitham operator $W = w(-i \partial _x)$
as an integral with respect to the spectral measure
of the self-adjoint operator $-i \partial _x$ in $L^2( \mathbb R )$.
Thus after linearisation \eqref{dimWhitham} can be considered
as a Schr\"odinger equation with the self-adjoint operator
$-i \partial _x w(-i \partial _x)$.
Indeed, introducing operator $D = -i \partial _x$ one may
rewrite \eqref{dimWhitham} as
%
\[
	i \eta_t = D w(D) \eta + \frac 34 D \eta^2
	.
\]
%
From this point of view, for example, one may deduce straight away that for any
real valued solution $\eta \in C^1( \mathbb R, L^2( \mathbb R ) )$
of this equation the $L^2$-norm does not depend on time.

The Whitham equation was proposed by Whitham \cite{Wh1} as an alternative to the
well known Korteweg-de Vries (KdV) equation
%
\begin{equation}
\label{KdV}
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	- \frac 12 \left( \varkappa - \frac 13 \right) \eta_{xxx} = 0
	.
\end{equation}
%
Provided $\varkappa < 1/3$ one may rescale $x$ and $t$ by
$\sqrt{1 - 3 \varkappa}$ and arrive at the equation
%
\[
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	+ \frac 16 \eta_{xxx} = 0
	.
\]
%
Thus it is apparent that small capillary effect do not add anything new
to the KdV model.
However, when $\varkappa$ is near $1/3$ one cannot expect
that this model be applicable.
To describe surface waves in such a situation, one may 
use instead the fifth-order-model equation
%
\begin{equation}
\label{Kawahara}
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	- \frac 12 \left( \varkappa - \frac 13 \right) \eta_{xxx}
	+ \frac{1}{360} ( 19 - 30 \varkappa - 45 \varkappa^2 ) \eta_{xxxxx}
	= 0
	.
\end{equation}
%
In our numerical experiments we use $\varkappa = 1/3$, so that the equation
reduces to what is known as the Kawahara equation \citep{Biswas2009, Kawahara1972}, which 
has the following form:
%
\[
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	+ \frac{1}{90} \eta_{xxxxx}
	= 0
	.
\]
%

The validity of the KdV and Kawahara equations can be described in the terms
of the Stokes number
\(
	\mathcal S = \alpha \lambda^2
\)
where $\alpha = a/h_0$ and $\lambda/h_0$ represent a prominent amplitude and
a characteristic wavelength of the wave field respectively.
The KdV equation is known to be a good model for water waves if the amplitude of 
the waves is small and the wavelength is large when compared to the undisturbed depth,
and if in addition, the two non-dimensional quantities $\alpha$ and 
$1 / \lambda^2$ are of similar size which means $\mathcal S \sim 1$.
With the same requirement and $\varkappa$ near $1/3$ the Kawahara
equation gives better results for waves where capillarity is important.
Another alternative model to the KdV equation \eqref{KdV} known as the BBM equation
was put forward in \cite{Peregrine1966} and studied in depth in \cite{Benjamin1972}.
That one with the capillarity $\varkappa$ has the form
%
\begin{equation}
\label{BBM}
	\eta_t + \eta_x + \frac 32 \eta \eta_x
	+ \frac 12 \left( \varkappa - \frac 13 \right) \eta_{xxt} = 0
	.
\end{equation}
%
The linearized dispersion relation 
of this equation is not an exact match to the dispersion relation of the full water-wave problem,
but it is much closer than the KdV equation in the case when $\varkappa < 1/3$, 
and it might also be expected that this equation may be able to model shorter waves more
successfully than the KdV equation.
However, the domain of its applicability is $\mathcal S \sim 1$,
that coincides with the corresponding
restrictions of the KdV model \cite{Benjamin1972}.
One may also notice that \eqref{BBM} can also be scaled
to the equation without capillarity in the same way as \eqref{KdV}
providing capillarity $\varkappa$ is small.

Both KdV and Kawahara equations are generally believed to approximate
very long waves quite well,
but one notorious problem with these equations is that they do not model
accurately the dynamics of shorter waves.
Recognizing this shortcoming of the KdV equation, Whitham proposed to use
the same nonlinearity as the KdV equation, but coupled with a linear term
which mimics the linear dispersion relation of the full water-wave problem.
Thus, at least in theory, 
the Whitham equation can be expected to yield a 
description of the dynamics of shorter waves which is closer to
the governing Euler equations.
The Whitham equation \eqref{dimWhitham} has been studied from
a number of vantage points during recent years.
In particular, the existence of traveling and solitary waves
has been studied  \citep{Minzoni2013, Ehrnstrom2012, Ehrnstrom2009, Ehrnstrom2013}.
Well posedness of a similar equation
was investigated in \cite{Lannes2013}.
Moreover, it has been shown
in \citep{Hur2015, Hur2015104,Sanford2014}
that periodic solutions of equation \eqref{dimWhitham}
feature modulational instability 
for short enough waves in a similar way as small-amplitude periodic wave
solutions of the water-wave problem. 
The performance of the Whitham equation in the description of surface water
waves has been investigated in \cite{BKN} in the steady case without surface tension.
However, it appears that no study of the performance of the Whitham equation
in the presence of capillarity has been done.

In the present note, we give an asymptotic derivation of the Whitham equation
as a model for surface water waves, 
giving close consideration to the influence of the surface tension.
The derivation proceeds by examining the Hamiltonian
formulation of the water-wave problem due to Zhakarov, Craig and Sulem \cite{Zakharov1968, Craig1993}.
This approach is similar to the method of \cite{Craig1994}.
However, our consideration is not constrained heavily
by any particular scalar regime.
Firstly, a corresponding Whitham system is derived, and then the Whitham
equation is found by restricting the system to one-way propagation.
Secondly, we derive different models from the Whitham equation
and point out the corresponding domains of their applicability.

Finally, a numerical comparison of modeling properties
of the KdV, Kawahara and Whitham equations is given
with respect to the Euler system.

% *************************************************************************************
\subsection{Euler system and its Hamiltonian.}
% *************************************************************************************
%\setcounter{equation}{0}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The surface water-wave problem 
is generally described by the Euler equations with no-flow conditions at
the bottom, and kinematic and dynamic boundary conditions at
the free surface. Assuming weak transverse effects, the unknowns
are the surface elevation $\eta(x,t)$, the horizontal and vertical 
fluid velocities $u_1(x,z,t)$ and $u_2(x,z,t)$, respectively,
and the pressure $P(x,z,t)$.
If the assumption of irrotational flow is made, then
a velocity potential $\phi(x,z,t)$ can be used.
Taking the undisturbed depth $h_0 = 1$ as a unit of distance, 
and the parameter $\sqrt{h_0/g} = 1$ as a unit of time, 
the problem may be posed on a domain
$\left\{(x,z) \in \R^2 | -1 < z < \eta(x,t) \right\}$
which extends to infinity in the positive and negative $x$-direction.
Due to the incompressibility of the fluid, the potential then satisfies
the Laplace's equation in this domain.
The fact that the fluid cannot penetrate the bottom is expressed by
a homogeneous Neumann boundary condition at the flat bottom.
Thus we have
%
\begin{eqnarray*}
	\phi_{xx} + \phi_{zz} = 0 & \mbox{in} &  -1 < z < \eta(x,t)
	\\
	\phi_z = 0              & \mbox{on} &  z=-1.
\end{eqnarray*}
%
The pressure is eliminated with help
of the Bernoulli equation, and the free-surface boundary conditions are formulated
in terms of the potential $\varphi$ and the surface displacement $\eta$ by
%
\[
	\left.
		\begin{array}{rc}
			\eta_t+\phi_x\eta_x-\phi_z
			& =0,
			\\
			\phi_t+\frac{1}{2} \big( \phi^2_x+\phi^2_z \big) + \eta
			- \varkappa \eta_{xx} \big( 1 + \eta_x^2 \big)^{- 3/2}
			& = 0, 
		\end{array}
	\right\}
	\mbox{on} \ z=\eta(x,t).
\]
%
The first equation represents the definition of the fluid velocity
with respect to the Euler coordinates.
The second one is the Bernoulli equation with the capillary term.

The total energy of the system is given by the sum of the kinetic energy,
the potential energy and the surface tension energy,
and normalized in such a way that the total energy
is zero when no wave motion is present at the surface.
Accordingly the Hamiltonian function for this problem is
%
\[
	H = \int _{\mathbb R} \int_0^\eta z \, dz dx +
	\int _{\mathbb R} \int_{-1}^\eta \Sfrac{1}{2} |\nabla \phi|^2 \, dz dx
	+ \varkappa \int _{\mathbb R} \Sfrac{ \eta_x^2 }{ 1 + \sqrt{1 + \eta_x^2} } dx	
	.
\]
%
Defining the trace of the potential at the free surface as
$\Phi(x,t) = \phi(x,\eta(x,t),t)$,
one may integrate in $z$ in the first integral and use the divergence theorem 
on the second integral in order to arrive at the formulation
%
\begin{equation}
\label{Hamiltonian-eta-phi}
	H  = \int_\R \Big[ \Sfrac{1}{2} \eta^2 + \Sfrac{1}{2} \Phi G(\eta) \Phi
	+ \varkappa \Sfrac{ \eta_x^2 }{ 1 + \sqrt{1 + \eta_x^2} }	
	\Big] \, dx.
\end{equation}
%
This is the Hamiltonian of the water wave problem with surface tension
as for instance found in \cite{Alazard2011},
and written in terms of the Dirichlet-Neumann operator $G(\eta)$.
As shown in \cite{Nicholls}, the Dirichlet-Neumann operator is analytic
in a certain sense, and can be expanded as a power series as
%
\[
	G(\eta)\Phi = \sum_{j=0}^\infty G_j(\eta) \Phi
	.
\]
%
In order to proceed, we need to understand the first few terms in this series.
As shown in \cite{CS} and \cite{CG}, the first two terms in this series
can be written with the help of the operator $D = - i \partial_x$ as 
%
\[
	G_0(\eta)= D\tanh(D)
	, \qquad 
	G_1(\eta)= D\eta D - D\tanh(D) \eta D\tanh(D)
	. 
\]
%
%
Note that it can be shown that the terms $G_j(\eta)$ for $j\ge 2$ are of quadratic or
higher-order in $\eta$, and will therefore not be needed in the following
analysis.

It will be convenient for the present purpose to formulate the Hamiltonian
in terms of the dependent variable $u = \Phi_x$.
This new variable is proportional to the velocity of the fluid
tangential to the surface.
More precisely
\(
	u = \varphi_x + \eta_x \varphi_z
	= \varphi _{\tau} \sqrt{1 + \eta_x^2}
\)
where $\varphi _{\tau}$ is exactly the tangential
velocity component to the surface.
To this end, we define the operator $\mathcal{K}$ by
%
\begin{equation*}
	G(\eta) = D \mathcal{K}(\eta) D
	.
\end{equation*}
%
As was the case with $G(\eta)$, the operator $\mathcal{K}(\eta)$ 
can also be expanded in a Taylor series
with respect to powers of $\eta$ and $\eta_x$ as
%
\[
	\mathcal{K}(\eta) = \sum_{j=0}^\infty \mathcal{K}_j(\eta)
	, \qquad
	\mathcal{K}_j(\eta) = D^{-1}G_j(\eta)D^{-1}
	.
\]
%
In particular, note that $\mathcal{K}_0 = \tanh{D} / D$
and
$\mathcal{K}_1 = \eta - \tanh{D} (\eta \tanh{D})$.
After integrating by parts the Hamiltonian can be expressed as 
%
\begin{equation}
\label{Hamiltonian_eta_u}
	H  = \int_\R \Big[ \Sfrac{1}{2} \eta^2 + \Sfrac{1}{2} u \mathcal K(\eta) u
	+ \varkappa \Sfrac{ \eta_x^2 }{ 1 + \sqrt{1 + \eta_x^2} }	
	\Big] \, dx
	.
\end{equation}
%
The following analysis has the formal character of long-wave approximation.
Consider a wave-field having a characteristic non-dimensional wavelength
$\lambda$ and a characteristic non-dimensional amplitude $\alpha$.
We also introduce the small parameter $\mu = \frac{1}{\lambda}$.
To obtain different approximations of the discussed problem
the amplitude $\alpha$ is considered as a function of wave-number $\mu$.
Its behaviour at small wave-numbers defines different scaling regimes.
The long-wave approximation means the scale $\eta = O(\alpha)$,
$u = O(\alpha)$ and $D = - i \partial_x = O(\mu)$
where $\alpha = \alpha(\mu)$ depends on the small parameter $\mu$.
Now the Hamiltonian (\ref{Hamiltonian_eta_u}) may be simplified as follows
%
\begin{equation}
\label{Hamiltonian_expansion}
	H = H_g + H_c + O(\mu^2 \alpha^4)
\end{equation}
%
with the gravity term
%
\begin{equation}
\label{Hamiltonian0}
	H_g =
	\frac 12 \int_\R
	\Big[
		\eta^2 + u \Sfrac{\tanh D}{D} u + \eta u^2 -
		u \tanh{D} (\eta \tanh{D} u)
	\Big]
	dx
\end{equation}
%
and the capillary part
%
\begin{equation}
\label{Hamiltonian1}
	H_c
	=
	\varkappa \int_\R \Sfrac{ \eta_x^2 dx }{ 1 + \sqrt{1 + \eta_x^2} }	
	=
	\frac{\varkappa}{2} \int_\R \eta_x^2 dx 
	+ O(\mu^4 \alpha^4)
	.
\end{equation}
%
Before we continue with derivation of the Whitham equation we prove
the following lemma about integration by parts, which is certainly
well-known and we add it here only for completeness.
%
\begin{lemma}
\label{integration_by_parts}
	Let $f, g$ be real-valued square integrable functions
	on real axis $\mathbb R$.
	Regard $D = -i \partial_x$ as self-adjoint on $L^2(\mathbb R, \mathbb C)$
	and a real-valued function $\varphi$ that is measurable and
	almost everywhere finite with respect to Lebesgue measure.
	If $f, g$ lie in the domain of the operator $\varphi(D)$ then
	%
	\[
		\int F \varphi(D) g = \int G \varphi(-D) f 
	\]
\end{lemma}
%
\begin{proof}
	It is given two proofs below.
	The first one is to regard 
	%
	\(
		\varphi(D) = ( \mathcal F^{-1} \varphi ) *
	\)
	%
	as	the operator of convolution
	in the sense of distribution theory
	%
	\begin{multline*}
		\int f \varphi(D) g
		= 
		\int f(\xi) ( \mathcal F^{-1} \varphi ) (\xi - x) g(x) dx d\xi
		=
		\\
		= 
		\int f(\xi) ( \mathcal F^{-1} \varphi \circ (-id) ) (x - \xi) g(x) dx d\xi
		=		
		\int g \varphi(-D) g
		.	
	\end{multline*}
	%
	The second proof is to represent
	%
	\(
		\varphi(D) = \int \varphi dE
	\)
	%
	as the integral with respect to spectral measure $E$
	of the operator $D$.
	The corresponding projector of the interval $(\alpha, \beta)$
	is the convolution with the function
	%
	\(
		e _{(\alpha, \beta)}(x) =
		\frac{1}{2 \pi ix} ( e^{i \beta x} - e^{i \alpha x} )
	\).
	So the replacement of $f$ and $g$ changes the corresponding spectral
	complex measure of intervals as follows
	%
	\begin{multline*}
		\mu _{f,g}(\alpha, \beta)
		=
		(E(\alpha, \beta)f,g)
		=
		\int e _{(\alpha, \beta)}(x-y) f(y) g(x) dxdy
		=
		\\
		=
		\int e _{(-\beta, -\alpha)}(y-x) f(y) g(x) dxdy
		=
		\mu _{g,f}(-\beta, -\alpha)
	\end{multline*}
	%
	which implies the statement of the lemma
	%
	\[
		\int f \varphi(D) g
		= 
		\int \varphi (x)d\mu _{g,f}(x)
		=
		\int \varphi (-x)d\mu _{f,g}(x)
		=		
		\int g \varphi(-D) f
		.	
	\]
\end{proof}

% *************************************************************************************
\subsection{Derivation of the Whitham type evolution system.}
% *************************************************************************************
%\setcounter{equation}{0}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%

The water-wave problem can be rewritten as a Hamiltonian system using
the variational derivatives of $H$.
Making reference to \citep{Craig1994}, Craig2005} note that
the pair $( \eta, \Phi )$ represents the canonical variables
for the Hamiltonian function \eqref{Hamiltonian-eta-phi}.
However, it is more common to write the equations of motion
in the fluid dynamics of free surface in terms of $\eta$
and $u = \Phi_x$.
The transformation
\(
	( \eta, \Phi ) \mapsto ( \eta, u )
\)
is associated with the Jacobian
%
\[
	\frac{ \partial ( \eta, u ) }{ \partial ( \eta, \Phi )  }	
	=
	\begin{pmatrix}
		1 & 0
		\\
		0 & \partial_x
	\end{pmatrix}
	.
\]
%
Thus in terms of $\eta$ and $u$ the Hamiltonian equations have the form
%
\begin{equation}
\label{Hamilton_system_eta_u}
	\eta_t = - \partial_x \frac{\delta H}{\delta u} 
	, \qquad
	u_t = - \partial_x \frac{\delta H}{\delta \eta}
\end{equation}
%
that is not canonical since the associated structure map $J_{\eta,u}$ is symmetric:
%
\[
	J_{\eta,u}
	=
	\left(
		\frac{ \partial ( \eta, u ) }{ \partial ( \eta, \Phi )  }	
	\right)
	\begin{pmatrix}
		0 & 1
		\\
		-1 & 0
	\end{pmatrix}
	\left(
		\frac{ \partial ( \eta, u ) }{ \partial ( \eta, \Phi )  }	
	\right) ^*
	=
	\begin{pmatrix}
		0 & -\partial_x
		\\
		-\partial_x & 0
	\end{pmatrix}
	.
\]
%
We now proceed to derive a system of equations which is similar to the Whitham equation
\eqref{dimWhitham}, but allows bi-directional wave propagation. This system will be
a stepping stone on the way to a derivation of \eqref{dimWhitham}, but may also be
of independent interest.
The variational derivative ${\delta H_g} / {\delta u}$ is defined
by means of any real-valued square integrable function $h$ as follows
%
\begin{multline*}
	\int _{\mathbb R} \frac{\delta H_g}{\delta u} (x)h(x) dx
	=
	d_u H_g h
	=
	\left. \frac{d}{d\tau} \right| _{\tau = 0} H_g( u + \tau h, \eta )
	= \int _{\mathbb R} u \eta h dx
	+
	\\
	+
	\frac 12 \int_\R
	\Big[
		h \Sfrac{\tanh D}{D} u + u \Sfrac{\tanh D}{D} h -
		h \tanh{D} (\eta \tanh{D} u) - u \tanh{D} (\eta \tanh{D} h)
	\Big]
	dx
	.
\end{multline*}
%
Making use of integration by parts described in
Lemma~{\ref{integration_by_parts}} one obtains
%
\[
	\frac{\delta H_g}{\delta u}
	=
	\frac{\tanh D}{D} u + \eta u - \tanh{D} (\eta \tanh{D} u)
	=
	\frac{\tanh D}{D} u + \eta u + O(\mu^2 \alpha^2)
\]
%
and in the same way
%
\[
	\frac{\delta H_g}{\delta \eta}
	=
	\eta + \frac 12 u^2 + \frac 12 (\tanh{D} u)^2
	=
	\eta + \frac 12 u^2 + O(\mu^2 \alpha^2)
	.
\]
%
These variational derivatives were also obtained by Moldabayev and Kalisch
\cite{Moldabayev2015}.
The capillary part $H_c$ defined by (\ref{Hamiltonian1})
gives the pressure $P$ on the surface
%
\[
	\frac{\delta H_c}{\delta \eta}
	=
	- \varkappa \frac{ \eta_{xx} }{ ( 1 + \eta_x^2 )^{\frac 32} }	
	=
	- \varkappa \eta_{xx} + O(\mu^4 \alpha^3)
	.
\]
%
At last the Hamilton system (\ref{Hamilton_system_eta_u})
is simplified to the Whitham system
%
\begin{align}
\label{sys1}
	\eta_t &=
	- \frac{\tanh D}{D} u_x - (\eta u)_x +
	\tanh{D} (\eta \tanh{D} u)_x + O(\mu^3 \alpha^3)
	, \\
\label{sys2}
	u_t &=
	- \eta_x - u u_x - (\tanh{D} u) \tanh{D} u_x
	+ \varkappa \eta_{xxx} + O(\mu^3 \alpha^3)
\end{align}
%
which is in line with the system obtained in \cite{Moldabayev2015}
%
\begin{align*}
	\eta_t &=
	- \frac{\tanh D}{D} u_x - (\eta u)_x + O(\mu^3 \alpha^2)
	, \\
	u_t &=
	- \eta_x - u u_x + \varkappa \eta_{xxx} + O(\mu^3 \alpha^2)
	.
\end{align*}
%
%
%
% *************************************************************************************
\subsection{Derivation of Whitham type evolution equations.}
% *************************************************************************************
%\setcounter{equation}{0}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
It turns out that the Whitham system \eqref{sys1}, \eqref{sys2} might be
rewritten as a system of two independent equations by further simplification.
More precisely, they will be independent with respect to
the linear approximation of that system.
For this purpose we need to separate solutions corresponding
to waves moving in other directions.
In order to derive the Whitham equation for uni-directional wave propagation,
it is important to understand how one-way propagation works in the 
Whitham system \eqref{sys1}, \eqref{sys2}.
Regard the linearisation of this system
%
\begin{align}
\label{linear_sys1}
	\eta_t + \frac{\tanh D}{D} u_x &= 0
	, \\
\label{linear_sys2}
	u_t + ( 1 + \varkappa D^2 ) \eta_x &= 0
	.
\end{align}
%
Regarding solutions of this linear system in the wave form
%
\[
	\eta(x, t) = A e^{ i \xi x - i \omega t }
	, \qquad
	u(x, t) = B e^{ i \xi x - i \omega t }
\]
%
gives rise to the matrix equation
%
\[
	\begin{pmatrix}
		- \omega					&	\tanh \xi
		\cr
		\xi + \varkappa \xi^3	&	- \omega	
	\end{pmatrix}
	\begin{pmatrix}
		A
		\cr
		B
	\end{pmatrix}
	=
	\begin{pmatrix}
		0
		\cr
		0
	\end{pmatrix}
	.
\]
%
This equation has a non-trivial solution, provided
its determinant equals zero, so that
\(
	\omega^2 - ( \xi + \varkappa \xi^3 ) \tanh \xi = 0
\).
Defining the phase speed as $c = \omega(\xi) / \xi$ one obtains
the dispersion relation 
%
\[
	c^2(\xi) = ( 1 + \varkappa \xi^2 ) \frac{\tanh \xi}{\xi}
\]
%
which coincides, up to the sign of $c$, with Whitham dispersion
relation (\ref{symbol}).
Obviously, the choice $c > 0$ corresponds to right-going wave solutions
of the linear system \eqref{linear_sys1}, \eqref{linear_sys2}.
And the phase speed $c < 0$ gives left-going waves.
To split up these two kinds of waves we regard the following
transformation of variables
%
\begin{equation}
\label{variable_transformation}
	r = \frac 12 (\eta + K  u)
	, \qquad
	s = \frac 12 (\eta - K  u)
	. 
\end{equation}
%
It is supposed that $K$ is an invertible operator,
namely an invertible function of the differential operator $D$.
The inverse transformation has the form
%
\begin{equation}
\label{inverse_variable_transformation}
	\eta = r + s
	, \qquad
	u = K^{-1} (r - s)
	. 
\end{equation}
%
The question arises whether it is possible to choose
such operator $K$ that $r$ and $s$ correspond to
right- and left-going waves, respectively.
After applying the transformation (\ref{inverse_variable_transformation})
to the linear system \eqref{linear_sys1}, \eqref{linear_sys2} one
arrives to the system
%
\begin{align*}
	r_t + \partial _x \big( A(D,K)r + B(D,K)s \big) &= 0
	, \\
	s_t - \partial _x \big( A(D,K)s + B(D,K)r \big) &= 0
\end{align*}
%
where operators $A$ and $B$ depend on $D$ and $K$ as follows
\[
	A = \frac 12 \left( (1 + \varkappa D^2) K + \frac{\tanh D}{D} K^{-1} \right)
	, \qquad
	B = \frac 12 \left( (1 + \varkappa D^2) K - \frac{\tanh D}{D} K^{-1} \right)
	.
\]
%
So to achieve independence of the obtained two equations
we need to choose the transformation $K$ in the way $B(D,K) = 0$,
so that
\begin{equation}
\label{K_transformation}
	K = \sqrt{ \frac{1}{ 1 + \varkappa D^2 } \cdot \frac{\tanh D}{D} }
\end{equation}
%
whish leads to the two independent Whitham equations
%
\begin{align}
\label{right_linear_sys1}
	r_t + \partial _x Wr &= 0
	, \\
\label{left_linear_sys2}
	s_t - \partial _x Ws &= 0
\end{align}
%
where the Whitham operator $W = w(D) = A(D,K)$ was introduced at
the beginning of the paper by (\ref{symbol}).
If we again regard the wave solutions
\(
	r(x, t) = \exp({ i \xi x - i \omega_r t })
\)
and
\(
	s(x, t) = \exp({ i \xi x - i \omega_s t })
\)
then we conclude that the first equation (\ref{right_linear_sys1})
describes waves moving to the right with the phase
velocity $c_r = \omega_r / \xi = w(\xi)$
and the second equation (\ref{left_linear_sys2})
corresponds to the left-going waves with
$c_s = \omega_s / \xi = - w(\xi)$.

Now we regard the Hamiltonian (\ref{Hamiltonian_eta_u})
as a functional of $r$ and $s$
with the same long-wave approximation as before (\ref{Hamiltonian_expansion})
where obviously $r = O(\alpha)$ and
$s = O(\alpha)$.
The unperturbed Hamiltonian part (\ref{Hamiltonian0}) is
%
\begin{multline}
\label{Hamiltonian0_r_s}
	H_g =
	\frac 12 \int_\R
	\Big[
		(r + s)^2 + \big( K^{-1} (r - s) \big) \Sfrac{\tanh D}{D}
		K^{-1} (r - s) + (r + s) \big( K^{-1} (r - s) \big) ^2 -
	%\right.
	\\
	%\left.
		\big( K^{-1} (r - s) \big) \tanh{D} \big( (r + s)
		\tanh{D} K^{-1} (r - s) \big)
	\Big]
	dx
\end{multline}
%
and the surface tension adding (\ref{Hamiltonian1}) is
%
\begin{equation}
\label{Hamiltonian1_r_s}
	H_c
	=
	\frac{\varkappa}{2} \int_\R (r + s)_x^2 dx 
	+ O(\mu^4 \alpha^4)
	.
\end{equation}
%
According to the transformation theory detailed in \cite{Craig2005},
due to the changing of variables (\ref{variable_transformation})
or (\ref{inverse_variable_transformation}), the structure
map changes to
%
\[
	J_{r,s}
	=
	\left( \frac{\partial (r,s)}{\partial (\eta, u)} \right) J_{\eta, u}  
	\left( \frac{\partial (r,s)}{\partial (\eta, u)} \right)^*
	= 
	\begin{pmatrix}
		-\frac 12 \partial_x K & 0
		\\
		0 & \frac 12 \partial_x K
	\end{pmatrix}
	.
\]
%
The corresponding Hamiltonian system has the form
%
\begin{equation}
\label{Hamilton_system_r_s}
	r_t + \partial_x \left( \frac{K}{2} \frac{\delta H}{\delta r} \right) = 0
	, 
	\quad \quad 
	s_t - \partial_x \left( \frac{K}{2} \frac{\delta H}{\delta s} \right) = 0
	.
\end{equation}
%
These equations are equivalent to (\ref{Hamilton_system_eta_u}).
However, solutions $r(x,t)$ and $s(x,t)$ are the displacements
going right and left, respectively,
when a solution $u(x,t)$, representing the tangential velocity component
up to the curvature multiplier,
might not be imagined so easy.
As above we calculate the  G\^ateaux derivative of $H_g$
given by (\ref{Hamiltonian0_r_s})
with respect to $r$ at a real-valued square integrable function
%
\begin{multline*}
	\int _{\mathbb R} \frac{\delta H_g}{\delta r} (x)h(x) dx
	=
	d_r H_g h
	=
	\left. \frac{d}{d\tau} \right| _{\tau = 0} H_g( r + \tau h, s )
	=
	\int (r + s)h
	+
	\\
	+
	\frac 12 \int \big( K^{-1} h \big) \frac{\tanh D}{D} K^{-1} (r - s)
	+
	\frac 12 \int \big( K^{-1} (r - s) \big) \frac{\tanh D}{D} K^{-1} h
	+	
	\\
	+
	\frac 12 \int h	\big( K^{-1} (r - s) \big) ^2
	+
	\int (r + s) \big( K^{-1} (r - s) \big) K^{-1} h
	-
	\\
	-
	\frac 12 \int \big( K^{-1} h \big)
	\tanh{D} \big( (r + s) \tanh{D} K^{-1} (r - s) \big)
	-
	\\
	-
	\frac 12 \int \big( K^{-1} (r - s) \big)
	\tanh{D} \big( h \tanh{D} K^{-1} (r - s) \big)
	-
	\\
	-
	\frac 12 \int \big( K^{-1} (r - s) \big)
	\tanh{D} \big( (r + s) \tanh{D} K^{-1} h \big)
	.
\end{multline*}
%
Integrating by parts as in Lemma~{\ref{integration_by_parts}}
and taking into account that functions $\tanh D$ is odd while
$K$ is even (\ref{K_transformation}) with respect to $D$
one can obtain ${\delta H_g}/{\delta r}$,
and in the similar way ${\delta H_c}/{\delta r}$, ${\delta H_g}/{\delta s}$,
${\delta H_c}/{\delta s}$.
Thus as a result
%
\begin{multline}
\label{right_variational_derivative}
	\frac{K}{2} \frac{\delta H}{\delta r}
	=
	Wr + \frac 14 K \big( K^{-1} (r - s) \big)^2 +
	\frac 12 (r + s) K^{-1} (r - s)
	+
	\\
	+
	\frac 14 K \big( \tanh D K^{-1} (r - s) \big)^2 -
	\frac 12 \tanh D \big( (r + s) \tanh D K^{-1} (r - s) \big)
	+ O(\mu^2 \alpha^3)
\end{multline}
%
and
%
\begin{multline}
\label{left_variational_derivative}
	\frac{K}{2} \frac{\delta H}{\delta s}
	=
	Ws + \frac 14 K \big( K^{-1} (r - s) \big)^2 -
	\frac 12 (r + s) K^{-1} (r - s)
	+
	\\
	+
	\frac 14 K \big( \tanh D K^{-1} (r - s) \big)^2 +
	\frac 12 \tanh D \big( (r + s) \tanh D K^{-1} (r - s) \big)
	+ O(\mu^2 \alpha^3)
\end{multline}
%
together with (\ref{Hamilton_system_r_s}) are the Whitham system
describing the displacement $\eta(x, t)$ in terms of waves going
right and left.
This system is entirely equivalent to (\ref{sys1}), (\ref{sys2})
and at the same time gives rise to solutions with more clear physical meaning.
Another useful property of this system is that it can be enough to
regard only one equation if we are allowed to neglect the waves
going to a particular direction.
More precisely, regarding only right-going waves leads to
%
\begin{multline}
\label{only_right_variational_derivative}
	\frac{K}{2} \frac{\delta H}{\delta r}
	=
	Wr + \frac 14 K \big( K^{-1} r \big)^2 +
	\frac 12 r K^{-1} r
	+
	\\
	+
	\frac 14 K \big( \tanh D K^{-1} r \big)^2 -
	\frac 12 \tanh D \big( r \tanh D K^{-1} r \big)
	+ O(\alpha |s|) + O(\mu^2 \alpha^3)
\end{multline}
%
and only left-going waves gives
%
\begin{multline}
\label{only_left_variational_derivative}
	\frac{K}{2} \frac{\delta H}{\delta s}
	=
	Ws + \frac 14 K \big( K^{-1} s \big)^2 +
	\frac 12 s K^{-1} s
	+
	\\
	+
	\frac 14 K \big( \tanh D K^{-1} s \big)^2 -
	\frac 12 \tanh D \big( s \tanh D K^{-1} s \big)
	+ O(\alpha |r|) + O(\mu^2 \alpha^3)
	.
\end{multline}
%
As one can see these expressions are identical up to discarded parts.
Equality (\ref{only_right_variational_derivative}) together with the
first equation of (\ref{Hamilton_system_r_s}) corresponds to the
Whitham equation describing right-going surface waves.
Equality (\ref{only_left_variational_derivative}) together with the
second equation of (\ref{Hamilton_system_r_s}) corresponds to the
Whitham equation describing left-going surface waves.
Further simplifications can be made by studying
concrete regimes $\alpha(\mu)$.
As a matter of fact, examples of behaviour $\alpha(\mu)$ at small $\mu$
that we regard below need less accurate asymptotic.
So that operators $K$ and $\tanh D$ can still be simplified by
taking into account $D = O(\mu)$ in Equality
(\ref{only_right_variational_derivative}) as follows
%
\begin{multline}
\label{simplified_variational_derivative}
	\frac{K}{2} \frac{\delta H}{\delta r}
	=
	Wr + \frac 34 r^2 +
	\frac 14 \left( \varkappa - \frac 53 \right) r D^2 r
	-
	\frac 14 \left( \varkappa + \frac 43 \right) (D r)^2
	+
	\\
	+
	O(\mu^4 \alpha^2) + O(\alpha |s|) + O(\mu^2 \alpha^3)
	.
\end{multline}
%
It is in line with \cite{Moldabayev2015} for liquids
without surface tension $\varkappa = 0$.
There is the same expression for the other variational derivative
with replacement $r$ by $s$.
%
%
\subsection{Linear approximation.}
%
%
Let $r$ and $s$ be of the same order and $\alpha = o(1)$ as $\mu \to 0$.
Then (\ref{only_right_variational_derivative}),
(\ref{only_left_variational_derivative}) are simplified to
%
\[
	\frac{K}{2} \frac{\delta H}{\delta r} = Wr (1 + o(1))
	, \qquad
	\frac{K}{2} \frac{\delta H}{\delta s} = Ws (1 + o(1))
\]
%
that together with (\ref{Hamilton_system_r_s}) represent two
independent linear equations.
We again arrived to (\ref{right_linear_sys1}), (\ref{left_linear_sys2}).
This is the case when we do not have enough information about
order relation between right- and left-going waves.
%
%
\subsection{The shallow-water scaling regime.}
%
%
Let $\alpha = O(1)$ as $\mu \to 0$.
Assume also that left-going waves can be discarded $s = o(1)$.
In this case operators $W$ can also be simplified by
taking into account $D = O(\mu)$.
Expression (\ref{simplified_variational_derivative}) becomes
%
\[
	\frac{K}{2} \frac{\delta H}{\delta r}
	= r + \frac 34 r^2 + o(1)
\]
%
which leads to the shallow-water equation
%
\[
	r_t + r_x + \frac 32 r r_x = o(\mu)
	.
\]
%
%
\subsection{The Boussinesq scaling regime.}
%
%
Let $\alpha = O(\mu^2)$ and $s = o(\mu^2)$ as $\mu \to 0$.
Expression (\ref{simplified_variational_derivative}) becomes
%
\[
	\frac{K}{2} \frac{\delta H}{\delta r}
	= Wr + \frac 34 r^2 + o(\mu^4)
\]
%
that can be simplified further by two different ways
%
\[
	W = 1 + \frac 12 \left( \varkappa - \frac 13 \right) D^2 + O(\mu^4) 
	= \left( 1 - \frac 12 \left( \varkappa - \frac 13 \right) D^2 \right)
	^{-1} + O(\mu^4)
	.
\]
%
The first equality leads to the KdV equation
%
\[
	r_t + r_x - \frac 12 \left( \varkappa - \frac 13 \right) r_{xxx}
	+ \frac 32 r r_x = o(\mu^5)
	.
\]
%
The second equality gives rise to the BBM equation
%
\[
	r_t + r_x + \frac 12 \left( \varkappa - \frac 13 \right) r_{xxt}
	+ \frac 32 r r_x = o(\mu^5)
	.
\]
%
%
\subsection{The Pad\'e (2,2) approximation.}
%
%
Suppose $\alpha = O(\mu^4)$ and $s = o(\mu^4)$ as $\mu \to 0$.
Expression (\ref{simplified_variational_derivative}) becomes
%
\[
	\frac{K}{2} \frac{\delta H}{\delta r}
	= Wr + \frac 34 r^2 + o(\mu^8)
\]
%
that can be simplified further provided $\varkappa \neq 1/3$ by the way
%
\[
	W = \frac{ 1 + aD^2 }{ 1 + bD^2 } + O(\mu^6) 
\]
%
where constants $a$ and $b$ depend on $\varkappa$ as follows
%
\[
	a(\varkappa) = \frac
	{ 3 + 10 \varkappa - 45 \varkappa^2 }
	{ 20 (1 - 3 \varkappa) }
	,	
\]
%
\[
	b(\varkappa) = \frac
	{ 19 - 30 \varkappa - 45 \varkappa^2 }
	{ 60 (1 - 3 \varkappa) }
	.	
\]
%
The corresponding equation
%
\[
	r_t + r_x - a(\varkappa) r_{xxx} - b(\varkappa) r_{xxt}
	+ \frac 32 r r_x = o(\mu^9)
	.
\]
%
As one can see the order of this differential equation is the same
as the order of KdV or BBM, meanwhile the Pad\'e approximation
is more accurate.
In the case $\varkappa = 1/3$ one has to use
the usual Taylor approximation 
%
\[
	W = 1 + \frac 12 \left( \varkappa - \frac 13 \right) D^2
	+ \frac{1}{360} ( 19 - 30 \varkappa - 45 \varkappa^2 ) D^4 + O(\mu^6) 
\]
%
which gives rise to the equation of fifth order
%
\[
	r_t + r_x - \frac 12 \left( \varkappa - \frac 13 \right) r_{xxx}
	+ \frac{1}{360} ( 19 - 30 \varkappa - 45 \varkappa^2 ) r_{xxxxx}
	+ \frac 32 r r_x = o(\mu^9)
\]
that is the Kawahara equation \eqref{Kawahara}.
%
%
\subsection{The Whitham scaling regime.}
%
%
If we now assume $\alpha = O(\mu^N)$ for any positive integer $N$
and $s = o(\alpha)$ as $\mu \to 0$, then we arrive to an example
when the Whitham operator $W$ cannot be approximated
using a simple differential operator instead.
The simplest equation in this case is the Whitham equation
%
\[
	r_t + W r_x + \frac 32 r r_x = o(\mu \alpha^2)
	.
\]
%
An example of the function $\alpha(\mu)$ when the Whitham equation
works better than its approximations was given in \cite{Moldabayev2015}.
At the same time that function $\alpha(\mu)$ may be similar
to the Boussinesq scale at some wave-numbers $\mu$
as was pointed out in \cite{Moldabayev2015}.
%
%
% *************************************************************************************
\subsection{Numerical results.}
% *************************************************************************************
%\setcounter{equation}{0}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
The purpose of this section is to compare the performance of the Whitham
equation as a model for surface water waves to both the KdV equation \eqref{KdV}
and to the Kawahara equation \eqref{Kawahara}.
In other words, all these approximate models are compared to
the Euler system which is considered as giving the closest description
of an actual surface wave profile.
For this purpose initial data are imposed, the Whitham,
KdV and Kawahara equations are solved with periodic boundary conditions,
and the solutions are compared to the numerical solutions of the full Euler equations
with free-surface boundary conditions.
This matching is made in various scaling regimes
from small Stokes numbers to $\mathcal S \sim 1$, to large Stokes numbers.

\begin{table}
\begin{center}
%	\label{experiment_table}
	\begin{tabular}{| c | c | c | c |}
		\hline
		Experiment  &  Stokes number $\mathcal S$
		&
		Amplitude $\alpha$  &  Wavelength $\lambda$
		\\
		\hline
		A	&  1   &  	 0.1 &	$\sqrt{10}$
		\\	
		B	&  1   & 	 0.2 &	$\sqrt{5}$ 			
		\\
		C	&  10  & 	 0.1 &	10	
		\\
		D	&  10  & 	 0.2 &	$\sqrt{50}$	
		\\	
		E	&  0.1 & 	 0.1 &	$1$ 	
		\\
		F	&  0.1 & 	 0.2 &	$\sqrt{ 1/2 }$			
		\\
		\hline
	\end{tabular}
\end{center}
\caption{Summary of the Stokes number, nondimensional wavelength,
nondimensional amplitude of the initial data used in the numerical experiments.}
\end{table}

The numerical treatment of the three model equations is a standard spectral scheme,
such as used in \cite{Fornberg1978} and \cite{Ehrnstrom2013} for example. For the time stepping,
an efficient fourth-order implicit method developed in \cite{DeFrutos1992} is used. 
The numerical treatment of the free-surface problem for the Euler equations
is based on a conformal mapping of the fluid domain into a rectangle. 
In the case of transient dynamics, this method has roots
in the work of Ovsyannikov \cite{Ovsyannikov1974}, 
and was later used in \cite{Dyachenko1996} and \cite{Li2004}.
In the case of periodic
boundary conditions, a Fourier-spectral collocation method can be used
for the computations, and the particular method used for the numerical experiments
reported here is detailed in \cite{Mitsotakis2014}.

Initial conditions for the Euler equations are chosen in such a way that 
the solutions are expected to be right moving.
This can be achieved by
imposing an initial surface disturbance $\eta_0$ together with the initial
trace of the potential
\(
	\Phi(x) = \int_0^x \eta_0(\xi) \, d\xi
	.
\)
Indeed, the right-going wave condition is $s(x,t) = 0$ which together with
\eqref{variable_transformation} and \eqref{K_transformation} imply
%
\[
	\Phi(x) = \int_0^x u(\xi, 0) d\xi
	= \int_0^x K^{-1} \eta_0 (\xi) d\xi
	\approx \int_0^x \eta_0 (\xi) d\xi
	.
\]
%
This last provision makes our numerical experiments more natural,
since it is not assumed that the regarded surface waves are
strictly right-moving $s = O(\mu^2 \alpha)$ and $\eta = r + O(\mu^2 \alpha)$.
In Fig. 1 one can see the corresponding small
wave $s$ moving to the left given by solving the Euler system.
In order to normalize the data, we choose $\eta_0$ in such a way that
the average of $\eta_0$ over the computational domain is zero.
The experiments are performed with
several different amplitudes $\alpha$ and wavelengths $\lambda$.
For the purpose of this section, we define the 
wavelength $\lambda$ as the distance between 
the two points $x_1$ and $x_2$ at which $\eta_0(x_1) = \eta_0(x_2) = \alpha /2$.
Both positive and negative initial disturbances are considered.
Numerical experiments were performed with a range of parameters
for amplitude $\alpha$ and the wave-length $\lambda$.
The summary of experiments' settings is given in Table 1.
All experiments are made with initial wave of elevation and wave of depression, labeled as ``positive" and ``negative" respectively.
The domain for computations is $-L\leq x \leq L$, with $L = 100$. 
The ``positive" initial data is
%
\begin{equation}
\label{test_1}
	\eta_0(x) = a\cdot \sech^2(f(\lambda)x)-C, 
\end{equation}
%
where 
%
\[
	f(\lambda) = \frac{2}{ \lambda }
	\log \left( \frac{ 1 + \sqrt{1/2} }{ \sqrt{1/2} } \right)
	, \qquad
	C = \frac{1}{L}\frac{a}{f(\lambda)}\tanh \left(\frac{L}{f(\lambda)}\right)
	.
\]
%
Here $C$ and $f(\lambda)$ are chosen so that 
$\int_{-L}^{L}\eta_0(x)dx = 0$, and
the wave-length $\lambda$
is the distance between the two points $x_1$ and $x_2$
at which $\eta_0(x_1) = \eta_0(x_2) = a/2$.
The velocity potential in this case is:
%
\begin{equation}
\label{velocity1}
	\Phi(x) = \frac{a}{f(\lambda)}\tanh(f(\lambda)x)-Cx
	,
\end{equation}
%
The ``negative" case function is just the ``reverse" of the first one
%
\begin{equation}
\label{test_2}
	\eta_0(x) = - a\cdot \sech^2(f(\lambda)x) + C
	. 
\end{equation}
%
The definitions for $f(\lambda)$ and $C$ are the same.
And the velocity potential is
%
\begin{equation}
\label{velocity2}
	\Phi(x) = -\frac{a}{f(\lambda)}\tanh(f(\lambda)x)+Cx
	,
\end{equation}
%
We calculate solutions of the Whitham equation and the Euler system.
We also calculate solutions of the KdV for the capillarity $\varkappa = 1/2$,
and solutions of the Kawahara in the case $\varkappa = 1/3$.

In Fig. 1,
the time evolution of a wave with an initial narrow peak is shown according to
the Euler (black line), Whitham (red line) and KdV (blue line) equations. 
Here the amplitude $\alpha = 0.2$,
the wavelength $\lambda = \sqrt{5}$ and the capillary parameter $\varkappa = 1/2$
are used.
This case corresponds the Stokes number $\mathcal S = 1$.
It appears that the KdV equation
produces a significant amount of spurious oscillations
and the Whitham equation gives the closest approximation of
the corresponding Euler solution.
As one can expect the unidirectional models 
also lag in the description of
waves going to the left.

In order to compare the accuracy of each approximate model
we calculate the differences, firstly between the Whitham and Euler equations,
and secondly between the KdV (or Kawahara) and Euler equations.
These differences are measured in the integral $L^2$-norm
normalized by initial condition $\lVert \eta_0 \rVert$ as follows
%
\[
	\frac{ \lVert \eta_E(t) - \eta(t) \rVert }{ \lVert \eta_0 \rVert }
	=
	\sqrt
	{
		\frac{ \int ( \eta_E(x, t) - \eta(x, t) )^2 dx }{ \int \eta_0(x)^2 dx }
	}
\]
%
where $ \eta_E(x, t) $ is the solution for the Euler system
and $ \eta(x, t) $ corresponds either to the Whitham, KdV or Kawahara equation.
The next Figs. 2, 3, 4, 5 show the dependence of $L^2$-error on time for different
initial situations. Thus as one can see, the Whitham model
performs better then the KdV and Kawahara equations,
in nearly all situations except the cases with a negative initial wave of depression
and Stokes number approximately unity.

### 2017-02-04 Thesis/latex/paperB.tex
Bibliography.

@@
 where the convolution kernel of the operator
-\(
+\[
 	W \eta_x = w(-i \partial _x) \eta_x
 	= \big( {\mathcal F}^{-1} w \big) * \eta_x
-\)	
+\]	
 is given in terms of the Fourier transform by
-%
-\begin{equation}
-\label{symbol}
 %	\mathcal{F} \big( K_{h_0} \big) =
 %	{\textstyle{ \sqrt{\frac{g\tanh (h_0\xi)}{\xi}}}}
+%
+\begin{equation}
+\label{symbol}
 	w(\xi) = \sqrt{ ( 1 + \varkappa \xi^2 ) \Sfrac{ \tanh(\xi) }{ \xi }}.
@@
 
-The Whitham equation was proposed by Whitham \cite{Wh1} as an alternative to the
+The Whitham equation was proposed by Whitham \cite{Whitham1967} as an alternative to the
 well known Korteweg-de Vries (KdV) equation
@@
 The performance of the Whitham equation in the description of surface water
-waves has been investigated in \cite{BKN} in the steady case without surface tension.
+waves has been investigated in \cite{Borluk2016} in the steady case without surface tension.
 However, it appears that no study of the performance of the Whitham equation
@@
 and written in terms of the Dirichlet-Neumann operator $G(\eta)$.
-As shown in \cite{Nicholls}, the Dirichlet-Neumann operator is analytic
+As shown in \cite{Nicholls2001}, the Dirichlet-Neumann operator is analytic
 in a certain sense, and can be expanded as a power series as
@@
 In order to proceed, we need to understand the first few terms in this series.
-As shown in \cite{CS} and \cite{CG}, the first two terms in this series
+As shown in \cite{Craig1993} and \cite{Craig1994}, the first two terms in this series
 can be written with the help of the operator $D = - i \partial_x$ as 
@@
 where $\alpha = \alpha(\mu)$ depends on the small parameter $\mu$.
-Now the Hamiltonian (\ref{Hamiltonian_eta_u}) may be simplified as follows
+Now the Hamiltonian \eqref{Hamiltonian_eta_u} may be simplified as follows
 %
@@
 the variational derivatives of $H$.
-Making reference to \citep{Craig1994}, Craig2005} note that
+Making reference to \citep{Craig1994, Craig2005} note that
 the pair $( \eta, \Phi )$ represents the canonical variables
@@
 %
-\subsection{Linear approximation.}
+\subsubsection{Linear approximation.}
 %
@@
 %
-\subsection{The shallow-water scaling regime.}
+\subsubsection{The shallow-water scaling regime.}
 %
@@
 %
-\subsection{The Boussinesq scaling regime.}
+\subsubsection{The Boussinesq scaling regime.}
 %

### 2017-02-08 Thesis/latex/paperC.tex
TBD: paperB, paperC, bibliography
paperC - start

--- 2017-01-30 Thesis/latex/summary.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Summary of results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This chapter provides an overview of the results achieved in the course of research work. 
-Detailed 
-\input{paperA}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Whitham equation with surface tension}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Explicit solutions for a long-wave model with constant vorticity}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In nonlinear dispersive evolution equations, the competing effects of nonlinearity and dispersion 
+make a number of interesting phenomena possible. In the current work, the focus is on the numerical
+approximation of traveling-wave solutions of such equations.
+We describe our efforts to write a dedicated \textsf{Python}
+code which is able to compute traveling-wave solutions of nonlinear
+dispersive equations in a very general form.
+%\begin{equation*}
+%u_t + [f(u)]_{x} + \Lop u_x = 0,
+%\end{equation*}
+%%\vskip -0.05cm
+%where $\Lop$ is a self-adjoint operator, and $f$ is a real-valued function with 
+%$f(0) = 0$ and $f'(0) = 0$.
+
+The \textsf{SpecTraVVave} code uses a continuation method coupled with a spectral projection
+to compute approximations of steady symmetric solutions of this equation.
+The code is used in a number of situations to gain an
+understanding of traveling-wave solutions. The first case is the Whitham
+equation, where numerical evidence points to the conclusion that the main 
+bifurcation branch features three distinct points of interest, namely a turning point, 
+a point of stability inversion, and a terminal point which corresponds to a cusped wave.
+
+The second case is the so-called modified Benjamin--Ono equation where
+the interaction of two solitary waves is investigated. It is found
+that two solitary waves may interact in such a way
+that the smaller wave is annihilated. The third case concerns
+the Benjamin equation which features two competing dispersive operators.
+In this case, it is found that bifurcation curves of periodic traveling-wave
+solutions may cross and connect high up on the branch in the nonlinear regime.
+%Include keywords and mathematical subject classification numbers as needed.
+
+\section{Introduction}
+\label{intro}
+
+This paper is concerned with traveling wave solutions for a class of nonlinear dispersive
+equations of the form
+\begin{equation}\label{nonlocal}
+u_t + [f(u)]_x + \Lop u_x = 0,
+\end{equation}
+where $\Lop$ is a self-adjoint operator, and $f$ is a real-valued function 
+with $f(0)=0$ and $f'(0)=0$, and which satisfies certain growth conditions.
+Equations of this form arise routinely in the study of wave problems in fluid mechanics
+and many other contexts.
+A prototype of such an equation is the KdV equation that appears
+if $\Lop= I + \frac{1}{6} \partial_x^2$ and $f(u) = \frac{3}{4} u^2$.
+In the current work, the operator $\Lop$ is considered to be given as a Fourier multiplier operator,
+such as for instance in the Benjamin--Ono equation, which arises in the study of
+interfacial waves. In this case, the Fourier multiplier operator is given by
+$\Lop =  I - \Hop \partial_x$, where the Hilbert transform
+$\Hop$ is defined as
+\begin{equation}
+\Hop u(x) = \frac{1}{\pi} \, \mathrm{p.v.} \int_{-\infty}^{\infty} \frac{u(x-y)}{y} \, \dd y, 
+\qquad \widehat{\Hop u}(k) = -\ii \operatorname{sgn}(k)\widehat{u}(k).
+\end{equation}
+We also study in detail traveling wave solutions of the Whitham equation, which appears when 
+$\Lop$ is given by convolution with the integral kernel $K_{h_0}$ in the form
+\begin{equation}\label{symbol}
+\Lop u(x) = \int_{-\infty}^{\infty} K_{h_0} (y) u(x-y) \, \dd y,
+\qquad \widehat{K_{h_0}} (k) 
+= {\textstyle{ \sqrt{\frac{g\tanh (h_0 k)}{k}}}}, 
+\end{equation}
+and $f$ is the same function as in the KdV equation.
+
+The particular form of equation~\eqref{nonlocal} exhibits the competing effects of dispersion and nonlinearity, 
+which gives rise to a host of interesting phenomena. The most well known special phenomenon is the existence
+of solitary waves and of periodic traveling waves containing higher Fourier modes.
+Indeed, note that in the purely dispersive model $u_t + \Lop u_x = 0$, 
+the only possible permanent progressive waves are simple sinusoidal waves, while
+in the nonlinear model~\eqref{nonlocal} higher Fourier modes must be considered
+to obtain solutions.
+
+The order of the operator $\mathcal{L}$ appearing in~\eqref{nonlocal} 
+has a major effect on the types of solutions that may be found. 
+A higher-order operator, such as in the Korteweg--de Vries equation, 
+acts as a smoothing operator because of its effect of spreading different frequency components
+out due to a strongly varying phase speed~\cite{Kato}. Lower-order operators 
+such as the operator $K_{h_0}$ in~\eqref{symbol} appearing in the Whitham equation
+may allow solutions to develop singularities, such as derivative blow-up (see~\cite{Hur2015}, \cite{HurTao})
+and formation of cusps (see~\cite{EW2016}).
+
+On the other hand, highly nonlinear functions $f(u)$ may lead to $L^{\infty}$-blow-up.
+For instance, the generalized KdV equation which is written in normalized form as
+\begin{equation}
+\label{kdv}
+u_t + u^{p}u_x + u_x + u_{xxx} = 0,
+\end{equation}
+features global existence of solutions for $p=1,2,3$, 
+but the solutions blow-up in the critical case $p=4$ (the case $p \ge 5$ is open).
+In the case of the generalized Benjamin--Ono equation
+\begin{equation*}
+u_t + u^{p}u_x + u_x - \mathcal{H} u_{xx} = 0,
+\end{equation*}
+where $\mathcal{H}$ is the Hilbert transform, numerical evidence points to singularity formation 
+for $p > 2$~\cite{BoKa}, but no proofs are available at this time.
+
+In order to study different phenomena related to equations of the form~\eqref{nonlocal} 
+and their traveling wave solutions, a \textsf{Python}-based solver package \textsf{SpecTraVVave} 
+was developed by the authors~\cite{github}. 
+The general idea behind the solver is to use a numerical continuation method~\cite{Keller} 
+implemented with a pseudo-spectral algorithm. 
+Similar previous projects include \textsf{AUTO}~\cite{Doed97} and \textsf{Wavetrain}~\cite{Sherratt}.
+\textsf{AUTO} is written in \emph{C}, whereas \textsf{Wavetrain} is written in \emph{Fortran}. 
+Both programs are efficient and very general, as they are able to cover a wide range of problems
+involving bifurcation analyses. However, from a user's perspective, 
+such a generality coupled with low level programming languages
+may lead to some difficulties in utilizing these programs efficiently.
+
+\textsf{SpecTraVVave} is designed to provide researchers with a simple yet effective tool 
+for investigating problems on traveling waves. The package is flexible, 
+and its functionality can be easily expanded. 
+The availability of the \textsf{IPython} notebook \cite{IPython} makes the solver very interactive,
+so that it should be easier for new users to get started.
+
+In order to maximize ease of use, \textsf{SpecTraVVave} was designed
+to find even solutions of~\eqref{nonlocal}.
+Symmetry of steady solutions can be proved for some of the models
+in the form~\eqref{nonlocal}, but not for all~\cite{CB}.
+Some of these equations also admit non-smooth solutions,
+for instance as termination points of a bifurcation branch.
+This happens for example for the Whitham equation, which features bifurcation
+curves which terminate in a solution with a cusp~\cite{EW2016}.
+One of the goals of the present paper is to investigate the
+precise nature of the termination of the bifurcation curve.
+
+The content of the paper is structured as follows. A mathematical description of the numerical method of 
+\textsf{SpecTraVVave} is given in \autoref{sec:spectral}.
+\autoref{sec:experiments} presents results of different experiments carried out with the package.
+Concluding remarks are given in \autoref{sec:conclusion}.
+A method for finding initial guesses for the solver is described in \autoref{sec:stokes}.
+\autoref{sec:implementation} contains a schematic of program and a description of its workflow.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Spectral scheme and construction of nonlinear system.}
+\label{sec:spectral}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Cosine collocation method.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+To compute traveling wave solutions to the equation~\eqref{nonlocal} the following ansatz is employed:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation*}%\label{ansatz}
+u(x,t) = \phi(x-ct).
+\end{equation*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Thus, the equation takes the form
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation*}%\label{phi-nonlocal}
+\phi' + \left[f(\phi) \right]' + \Lop \phi' = 0,
+\end{equation*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+which can be integrated to give
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation}\label{int_gen1}
+-c \phi + f(\phi)  + \Lop \phi = B. 
+\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The constant $B$ is a priori undetermined. One may set the $B$ equal to 
+zero as a way of normalizing the solutions. Another option is to impose
+an additional condition, for example that the integral of $\phi$ over
+one wavelength be zero. In this case, $B$ will be found along with the solution $\phi$.
+
+We consider $\Lop$ as a Fourier multiplier operator with symbol
+$\alpha(k)$. We also assume that $f$ is at least twice differentiable,
+and we have $f(0)=0$ and $f'(0)=0$.
+When computing traveling-wave solutions we focus on even periodic solutions. 
+While it can be proved in some cases that solutions of~\eqref{int_gen1}
+must be even, this is not known for a general operator $\Lop$.
+Nevertheless, we make this assumption here in order to
+make the numerical procedure as uniform as possible.
+For even periodic solutions, 
+one may use a cosine collocation instead of a Fourier method. In particular, using
+the cosine functions as basis elements automatically removes the inherent symmetries
+due to reflective and translational symmetry. 
+Moreover, the number of unknowns is reduced by a factor of $2$, 
+and the problem of the asymmetric arrangement of nodes in the FFT
+is circumvented. Of course, all these problems could also be dealt with a collocation
+method based on the Fourier basis, but the cosine basis does all of the above automatically.
+In addition,  the \textsf{Python} cosine transform is based on an integrated algorithm, which
+relies on an optimized version of the discrete cosine transform (DCT).
+
+The following description of computation scheme was presented in detail in~\cite{EK2}, 
+but we will briefly repeat it here for consistency of the manuscript. For the purpose of clarity, 
+we will refer to full wavelength $L$ of a solution as the (full) wavelength, 
+and half of fundamental wavelength will be called half-wavelength. 
+Such a definition is required because the method computes a half of a solution profile, 
+the other half is automatically constructed due to symmetry.
+
+Traveling wave solutions to the equation~\eqref{int_gen1}
+are to be computed in the form of a linear combination of cosine functions of different
+wave-numbers, i.e., in the space
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+\SN = \mathrm{span}_{\R}\setc{\cos(lx)}{0 \leq l \leq N-1 }
+.
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+This is a subspace of $L^2(0,2\pi)$, and the collocation points 
+$x_n = \pi \frac{2n -1}{2N}$ for $n = 1,\dots, N$ are used to discretize the domain.  
+If the required full wavelength of solutions is to be $L \neq 2\pi$, 
+one can use a scaling on the $x$-variable. 
+Defining the new variable
+\begin{equation*}
+x' = \frac{L}{2\pi}x, 
+\end{equation*}
+yields collocation points $x_n'$ and wavenumbers $\kappa_l$ defined by
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+x_n' = \frac{L}{2} \frac{2n -1}{2N}, \quad\quad \kappa_l = \frac{2\pi}{L}l. 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+We are seeking a function $\phi_N \in \SN$ that satisfies the equations
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+-c\phi_N(x_n') + f(\phi_N)(x_n') + \Lop^N\phi_N(x_n') = 0, \label{int_gen-disc}
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+at the collocation points $x_n'$.
+The operator $\Lop^N$ is the discrete form of the operator $\Lop$, 
+and $\phi_N$ is the discrete cosine representation of $\phi$
+which is given by
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+\phi_N(x') &= \sum_{l = 0}^{N-1} \omega(\kappa_l)\Phi_N(\kappa_l)\cos(\kappa_{l}x'),\\
+\omega(\kappa_l) &= 
+\begin{cases}
+\sqrt{1/N}, \quad \kappa_l = 0, \\
+\sqrt{2/N}, \quad \kappa_l > 0,
+\end{cases}                  \\
+\Phi_N(\kappa_l) &= \omega(\kappa_l) \sum_{n = 1}^{N} \phi_N(x'_n) \cos(\kappa_l x'_n),
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+where $\kappa_l = 0,\frac{2\pi}{L},\ldots, \frac{2\pi}{L}(N-1)$ are the scaled wavenumbers,
+and $\Phi_N(\cdot)$ are the discrete cosine coefficients.
+As the equation~\eqref{int_gen-disc} is enforced at the collocation points $x_n'$, 
+one may evaluate the term $\Lop^N\phi_N$ using the matrix $\Lop^N(i,j)$ 
+defined by
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+\Lop^N\phi_N(x'_i) &= \sum_{j=1}^{N}\Lop^N(i,j) \phi_N(x'_j),\\
+\Lop^N(i,j) &= \sum_{l=0}^{N-1} \omega^2(\kappa_l) \alpha(\kappa_l) \cos(\kappa_l x'_i) \cos(\kappa_l x'_j),
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+where $\alpha(\cdot)$ is the Fourier multiplier function of the operator $\Lop$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Construction of nonlinear system.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+The equation~\eqref{int_gen-disc} enforced at $N$ collocation points 
+yields a nonlinear system of $N$ equations in $N$ unknowns, which can be written in shorthand as  
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+	\begin{equation*}
+		F(\phi_N) = 0.
+	\end{equation*}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+This system can be solved by a standard iterative method, such as Newton's method. 
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+	In this system, the value of phase speed $c$ has to be fixed for computing one particular solution. 
+	Such an approach becomes impractical when a turning point on the bifurcation curve appears.
+
+	In \textsf{SpecTraVVave} a different approach is employed: both the amplitude $a$ and the phase speed $c$ of a solution are treated as functions of a parameter $\theta$: $ a = a(\theta)$, $c = c(\theta)$.
+The parameter $\theta$ is to be computed from the system~\eqref{system-N3}.
+This construction makes it possible to follow turning points
+on the bifurcation branch with relative ease.
+	Having computed two solutions, i.e., two points on the bifurcation curve $P_1 = (c_1, a_1)$ and $P_2 = (c_2, a_2)$, one may find a direction vector $\mathbf{d}=(d^c, d^a)$ of the line that contains these points:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	\begin{align*}
+		\mathbf{d}:	\quad d^c = c_2 - c_1, \qquad	d^a = a_2-a_1.
+	\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	Then the point $P_3 = (c_3,a_3)$ is fixed at some (small) distance $s$ from the point $P_2$ in the direction $\mathbf{d}$.
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	\begin{align*}
+		P_3: \quad c_3 = c_2 + s \cdot d^c, \qquad	a_3 = a_2 + s \cdot d^a.
+	\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 The point $P_3$ plays the role of the initial guess for velocity and amplitude when computing the next solution $P_\ast = (c_\ast, a_\ast)$. 
+	The solution point $P_\ast$ is required to lay on the line with direction vector $\mathbf{d_{\bot}}=(d^c_{\bot}, d^a_{\bot})$, which is orthogonal to the vector $\mathbf{d}$.
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	\begin{align*}
+	\mathbf{d_{\bot}}: \quad d^c_{\bot} = - d^a, \qquad &d^a_{\bot} = d^c,  \\
+	P_\ast: \quad c_\ast = c_3 + \theta d^c_{\bot} \qquad &a_\ast = a_3 + \theta d^a_{\bot}.
+	\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 A schematic sketch of finding a new solution $P_\ast$ is given in \autoref{navig}.
+
+	The variable $\theta$ is computed by Newton's method from the extended system 
+%%%%%%%%%===========%%%%%%%%%%%%%%%	
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+F\left(
+\begin{matrix}
+\phi_N(x_1)\\ \vdots \\ \phi_N(x_N) \\ B \\ \theta
+\end{matrix}
+ \right)= \left(
+\begin{matrix}
+(-c+\mathcal{L}_N)\phi_N(x_1)+ f(\phi_N(x_1)) - B
+\\ \vdots 
+\\ (-c+\mathcal{L}_N)\phi_N(x_N)+ f(\phi_N(x_N)) - B
+\\ \Omega(\phi_N, c, a, B)
+\\ \phi_N(x_1)-\phi_N(x_N)-a
+\end{matrix}
+ \right)=
+\left(
+\begin{matrix}
+0
+\\ \vdots 
+\\ 0
+\\ 0
+\\ 0
+\end{matrix}
+ \right). \label{system-N3} 
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	Here a nonhomogeneous problem ($B \neq 0$) is considered.
+	The equation 	
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+	\phi_N(x_1)-\phi_N(x_N)-a = 0, 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 makes the waveheight of the computed solution to be that of $a$. The equation 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+	\Omega(\phi_N, c, a, B)=0, %\label{boundary}
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 is called the  \emph{boundary condition}. 
+	It allows to enforce different specifications on the computed traveling wave solution.   	
+	For example, if one sets  
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+	\Omega(\phi_N, c, a, B)= \phi_N(x_1)+\cdots+\phi_N(x_N), 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 then \emph{the mean of the computed wave over a period will have to be equal to zero}.
+	One may also experiment with	
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+	\Omega(\phi_N, c, a, B) = B, 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 to consider the homogeneous problem $(B=0)$.
+	It can be also interesting to set
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+	\Omega(\phi_N, c, a, B) = \phi_N(x_N). \label{solitary}
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%	
+	 This enables us to compute traveling wave solutions that mimic solitary wave solutions. 	
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}
+\centering
+		\includegraphics[width=0.35\textwidth]{navig}
+\caption{\small Navigation on the bifurcation curve.}
+\label{navig}
+\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+\subsection{Convergence.}
+In order to test the numerical implementation of the discretization,
+the method is applied to a case where the solution is known.
+One such case is the KdV equation
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+u_t + u_x + \frac{3}{2} u u_x + \frac{1}{6}u_{xxx} = 0, 
+%\label{KdV_sol*}
+\end{align*}
+which has a known solution, given in the form
+\begin{align*}
+u_{\mathrm{exact}}(x,t) = a \sech^2\Big( {\textstyle\sqrt{\frac{3a}{4}}} (x-c t)\Big),\\
+\end{align*}
+with $c=1+a/2$.
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%
+Using the boundary equation~\eqref{solitary}, \textsf{SpecTraVVave} is capable of computing 
+approximations to solitary wave solutions of nonlinear wave equations. 
+Solitary wave solutions are treated as traveling waves with sufficiently long wavelength 
+that have the wave trough at zero.
+In case of the KdV equation solitary wave solutions have exponential decay, 
+and therefore, considering the symmetry of solitary solutions, 
+the half-wavelength of $30$ is considered for the comparison. 
+\itodo[DM]{as mentioned before: should we call the (full-wavelength )/2 as half-wavelength OR just wavelength. 
+This may lead to confusions for reviewers.}
+Approximation errors are summarized in \autoref{t2}.
+%
+\begin{table}  [ht]
+\centering
+\begin{tabular}{rrrr}
+  \toprule
+Nb.\ of grid points & $\log_{10}(\|u_{\mathrm{exact}} - u\|_{L^{\infty}})$ & $\log_{10}(\|u_{\mathrm{exact}} - u\|_{L^2})$ & Ratio of $L^2$-errors\\
+\midrule
+32 & $-1.389$  & $-2.092$ & \\
+
+64 & $-3.705$  & $-4.549$ & $286.8$\\
+
+128 & $-8.809$  & $-9.508$ & $90935.0$\\
+
+256 & $-15.353$  & $-16.144$ & $4329670.9$\\
+
+512 & $-15.353$  & $-16.087$ & $0.9$\\
+\bottomrule
+\end{tabular}
+\caption{Estimates of error between the exact and computed solitary wave solutions for the KdV equation. Half-wavelength $30$, waveheight $a=1.2651$}
+\label{t2}
+\end{table}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.47\textwidth]{convergence}      
+\caption{\small Graph of error estimates given in \autoref{t2}.}
+\label{L2convergence}
+\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments with SpecTraVVave.}
+\label{sec:experiments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Termination of the waveheight-velocity bifurcation curve of the Whitham equation.}
+\label{sec-W-term}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The waveheight-velocity bifurcation curve of the Whitham equation 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+	u_t + \frac{3}{2} u u_x + Ku_x = 0, \qquad 
+	\widehat{Ku}(k) = \sqrt{\frac{\tanh(k)}{k}},  \label{W}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+was studied numerically in~\cite{EK2}. 	
+An attempt was made to identify the termination point of the Whitham bifurcation curve. 
+The investigation was limited by computational tools and complete results were not obtained. 
+In particular, the authors could not confirm that traveling wave solutions do not exist 
+past the point where the authors, based on pioneering work of Whitham
+\cite{Wh1} suspected a cusped solution.
+In this section a number of numerical results on nature of the bifurcation curve
+for the Whitham equation are presented.
+Solutions to the equation~\eqref{W}
+are computed in the form of traveling waves $u(x,t) = \phi(x - c t)$
+and the homogeneous $(B = 0)$ integrated version the equation is considered:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+-c \phi + \frac{3}{4} \phi^2 + K \phi = 0. \label{W-int}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+Special attention is given to relation between stability of solutions 
+and their waveheight and velocity parameters, i.e., their position on the bifurcation curve. 
+The following questions are under study: 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{enumerate}[label=\alph*)] % chktex 9 chktex 10
+\item Where does the bifurcation curve terminate?
+
+\item Where on the bifurcation curve do solutions change their stability?
+
+\item Is there any role that the turning point on the bifurcation curve plays?
+\end{enumerate}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\subsubsection{Implementation of the problem in SpecTraVVave.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The results presented here focus on $2\pi$-periodic solutions to the equation~\eqref{W-int}, 
+i.e., solutions of the system~\eqref{system-N3}. 
+\autoref{bif-curve-all} presents Whitham bifurcation curves with numbers of grid points 
+$N = 512$, $N=1024$ and $N=2048$. 
+The current implementation of the \textsf{SpecTraVVave} package allows fixing the number of grid points $N$ 
+and a so-called doubling parameter $\DD$, 
+i.e., the number by which $N$ is doubled as computations are made. 	
+This allows us to get sets of solutions with $N, 2N, \ldots, 2^{\DD} N$ grid points. 
+If $\DD = 1$ then only two sets of solutions are computed and they are regarded as 
+lower grid (lower resolution) and higher grid (higher resolution) solutions.
+While the system~\eqref{system-N3} is processed by Newton solver, 
+lower grid solutions are taken as initial guesses for higher grid solutions. 
+All curves shown in this manuscript have been produced
+after tests with a number of resolutions were run, 
+and the curves shown did not change significantly under further refinement.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.48\textwidth]{512_1024_2048}{}      
+\caption{\small Whitham bifurcation curve in different grid resolutions.}
+\label{bif-curve-all}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%\subsubsection{Specific solutions and their stability.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+\autoref{bif-L2-1024}\subref{sfig:bif-ch} % chktex 8
+presents the Whitham bifurcation curve computed by \textsf{SpecTraVVave} 
+with $N=1024$ and $\DD=1$. 
+There are three solutions which deserve to be singled out:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{enumerate}
+	\item Traveling wave solution with minimum velocity (rhombus);
+	\item Traveling wave solution with maximum $L^2$-norm (circle);
+	\item Cusped traveling wave solution (square).
+\end{enumerate}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+Profiles of the above listed solutions are given in \autoref{3waves}\subref{sfig:prof-sing}.
+The solution with minimum velocity corresponds to the turning point of the bifurcation curve.
+The solution with maximum $L^2$-norm is very close to the latter one, although it has a higher waveheight 
+and a different velocity.
+The solution marked by a square is called here the \emph{terminal solution}.
+As already mentioned, previous studies, such as~\cite{EK1,EK2} did not provide any conclusive
+analysis on the part of the bifurcation curve past the turning point.
+In particular, it was not clear whether solutions ceased to exist
+at or after the turning point, or whether solutions were stable
+or unstable after the turning point.
+
+Let us first focus on the stability of solutions.
+Note that \textsf{SpecTraVVave} has an evolution integrator routine, 
+which enables one to check the stability of computed solutions. 	
+The current version of the package uses the fourth-order method developed in~\cite{dFSS}.
+In addition one may use a more refined analysis, resting on the evaluation
+of invariant functionals. This analysis is based on the observation
+that the traveling waves can be thought of as solutions of a constrained
+minimization problem. This analysis is based on ideas developed by
+Boussinesq, first exploited in~\cite{B2}, and later used 
+in~\cite{BSS,TJB,NgK}, and many other works.
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[ht]
+\centering
+\subfigure[][Bifurcation curve in the wavespeed-waveheight parameter space]{\includegraphics[width=0.48\textwidth]{1024_2048}{}\label{sfig:bif-ch}}%
+~%
+\subfigure[][Bifurcation curve in a wavespeed-$L^2$-norm diagram]{\includegraphics[width=0.48\textwidth]{L2_1024_2048}{}} 
+\caption{\small
+Whitham bifurcation curves for $2\pi$-periodic solutions. 
+% $N = 1024, \mathcal{D}= 1$. 
+}
+\label{bif-L2-1024}
+\end{figure} 
+%trim = 1.2cm 4.7cm 0.6cm 4.5cm, clip,
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+Let us define two functionals $V$ and $E$:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+	V( \phi) = \frac{1}{2} \int_{-\infty}^{+\infty} \phi^2\,\dd\zeta, \qquad
+	E( \phi) = \int_{-\infty}^{+\infty} \left\{ \Sfrac{1}{2} \phi^3 - \phi \, K\!  \phi \right\}\, \dd\zeta.
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+The equation~\eqref{W-int} can be then written in terms of variational derivatives of $E$ and $V$
+as
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+	E' ( \phi ) - c V ' ( \phi ) = 0.
+\label{zero}
+\end{align} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+It is known from~\cite{BSS} that under certain conditions,
+the stability of solitary wave solutions depends 
+on convexity of the function $d(c) = E(\phi) - c V(\phi)$. 
+Solutions with values of $c$ for which $d''(c) > 0$  are stable solutions,
+and solutions with wave speeds for which $d''(c) <0$ are unstable solutions.
+
+The current numerical investigation can therefore be interpreted as an indication of the
+stability properties of traveling wave solutions of the Whitham equation.
+Note that differentiation of $d(c)$ yields
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+	d '(c) &= \underbrace{E'(\phi) - c V'(\phi)}_{=0} - V(\phi).
+\end{align*}
+Using~\eqref{zero} as indicated yields
+\begin{align}
+	d '(c) &= - V(\phi) = - \frac{1}{2} \int_{-\infty}^{+\infty} \phi ^2\,\dd\zeta = - \frac{1}{2}\|\phi\|_{L^2}^2 . 
+\label{d'c}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+Therefore, in order to understand the convexity of $d(c)$, it is sufficient
+to find points of maximum $L^2$-norm on the curve in the right panel
+of \autoref{bif-L2-1024}.
+It is straightforward to see that $d''(c)$ changes sign in the neighbourhood of the maximum point 
+of this curve, i.e., around the solution with maximum $L^2$-norm. 
+In particular, $d''(c) > 0$, i.e., solutions are stable to the left of the maximum point, 
+and $d''(c) < 0$, i.e., solutions are unstable to the right of the maximum point.
+
+In addition, the solutions were tested with the evolution integrator to confirm their stability/instability in time. 
+The solution with maximum $L^2$-norm and those on the left-hand side were always found to be stable in 
+the time-dependent computations.
+Solutions on the right-hand side do not preserve their shape and thus are unstable. 
+Examples are given in \autoref{time-evo}. 
+This analysis confirms that 
+the point corresponding to the minumum wave speed (the turning point), 
+and the point of stability inversion are two distinct points on the bifurcation curve.
+Moreover, the point of stability inversion
+is a little further up the branch from the turning point.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[t]
+\centering
+\subfigure[][Profiles of the solutions singled out in \autoref{bif-L2-1024}\subref{sfig:bif-ch}]{\includegraphics[width=0.48\textwidth]{1024_waves_1}{}\label{sfig:prof-sing}}
+~%
+\subfigure[][Profile after terminal solution]{\includegraphics[width=0.48\textwidth]{1024_waves_2}{}\label{sfig:prof-term}}
+\caption{\small Profiles of specific traveling wave solutions.}
+\label{3waves}
+\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+Next, we turn our attention to the analysis of the terminal point. There
+are two main questions. Does the branch terminate, and if so, 
+does the terminal point on the branch correspond to a cusped traveling wave.
+First of all, note that the solution, which is computed by \textsf{SpecTraVVave}, 
+past the terminal solution has two crests, no matter how small the 
+stepping on the bifurcation branch is taken.
+(see \autoref{3waves}\subref{sfig:prof-term}).   
+Secondly, as will be explained presently, the relation 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+\frac{c}{\sup_{x} \phi(x)} = \frac{3}{2} 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+holds for the terminal solution with a good degree of approximation.
+For the most accurate runs, we obtain $c / \sup_{x \in \R} \phi(x)$ $\approx$ $1.51$. 
+%
+To explain how this relation comes about note that the steady integrated form of the
+Whitham equation can be written as
+\begin{equation}\label{Whitham_Alt}
+\Big( \frac{c}{\sqrt{3}} - \frac{\sqrt{3}}{2}\phi \Big)^{2}  = \frac{1}{3} c^2 - K \phi.
+\end{equation}
+It is clear that for any $\phi < \frac{3}{2} c$, the relation~\eqref{Whitham_Alt}
+can used in a bootstrap argument to show that any continuous solution must be
+in fact smooth. However for the case $\phi =  \frac{3}{2} c$ this bootstrap
+argument fails since the left-hand side vanishes. It can be concluded that
+a solutions containing a cusp will have a maximum  value of $\frac{3}{2} c$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+As an additional check, the discrete cosine coefficients of the solutions were
+examined, and fitted to the following models:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+	\mathcal{E}(k) = \nu_1 e^{-\nu_2 k^n}, \qquad 
+	\mathcal{P}(k) = \frac{\mu_1}{\mu_2 + \mu_3 k^m},
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%
+where $\nu_1$, $\nu_2$, $\mu_1$, $\mu_2$, $\mu_3$, $n$ and $m$ are fitting parameters. 
+A smooth function is known to have discrete cosine coefficients with exponential decay in $k$. 
+On the other hand, if a function is not smooth, the discrete cosine coefficients feature polynomial decay. 
+To identify the best fit, two parameters were used: $L^2$-norm of the residual 
+and the Akaike information criterion (AIC) measure.
+
+From the data given in the \autoref{t1}, one can deduce that for solutions with minimum speed 
+and maximum $L^2$-norm exponential fit is better than polynomial. 
+That is not the case for the terminal solution. 
+Thus, the first two solutions are smooth and the terminal solution is nonsmooth. 
+In fact, the polynomial fit is better than exponential for solutions that 
+are between the maximum $L^2$-norm solution and the terminal solution.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[t]
+  \subfigure[][Minimum speed solution]{\includegraphics[width=0.31\textwidth]{1024_full_3p_1}}
+~%
+\subfigure[][Maximum $L^2$-norm solution]{\includegraphics[width=0.31\textwidth]{1024_full_3p_2}{}}
+~%
+\subfigure[][Terminal solution]{\includegraphics[width=0.31\textwidth]{1024_full_3p_3}{}}
+\caption{\small Evolution of specific solutions in time. Time range for each solution is three periods.}
+\label{time-evo}
+\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+The numerical evidence brought forward supports the conclusion 
+that the Whitham bifurcation branch terminates at the terminal point
+indicated in \autoref{bif-curve-all}. Of course, as already mentioned,
+this conclusion has now also been reached using tools of mathematical analysis
+\cite{EW2016}.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{table}	
+\centering
+\begin{tabular}{l|rr|rr|rr|}
+\cline{2-7}
+& \multicolumn{2}{ c     }{Min speed solution} & \multicolumn{2}{ c     }{Max $L^2$-norm solution}& \multicolumn{2}{ c |  }{Terminal solution} \\
+\multicolumn{1}{l|} {Model} &$\mathcal{E}(k)$&$\mathcal{P}(k)$&$\mathcal{E}(k)$&$\mathcal{P}(k)$&$\mathcal{E}(k)$&$\mathcal{P}(k)$ \\ 
+\midrule
+\multicolumn{1}{l|} {Residual's $L^2$ norm} & $5\times 10^{-5}$ &$4\times 10^{-3}$ & $7 \times 10^{-5}$ & $3\times 10^{-3}$ &  $6\times 10^{-3}$& $6\times 10^{-4}$ \\
+\multicolumn{1}{l|} {AIC} & -543 & -321 & -529 & -333 & -298 & -416 \\ 
+\bottomrule
+\end{tabular}
+\caption{\small Results for measures of fit.}
+\label{t1}
+\end{table}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Interaction of solitary wave solutions of modified Benjamin--Ono equation}
+\label{mB-O-interact}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this section, we utilize the \textsf{SpecTraVVave} package to obtain
+high-precision approximations to solitary-wave solutions of
+the modified Benjamin--Ono
+\begin{equation*}
+%\label{BO}
+u_t + u^2u_x + u_x - \mathcal{H} u_{xx} = 0,
+\end{equation*}
+which is a special case of the generalized Benjamin--Ono equation, with $p=2$.
+This case corresponds to the critical scaling, i.e., invariance of 
+the energy norm under the natural invariant scaling.
+
+The Benjamin--Ono equation was found by Benjamin~\cite{B1} as a model
+for long small-amplitude interfacial waves in deep water. 
+The validity of approximating the more physically correct configuration of a continuous
+density distribution by the two-layer approximation has recently been justified mathematically
+\cite{ChenWalsh2016}.
+
+Solitary-wave solutions of the modified Benjamin--Ono equation with $p=3$, $p=4$ and $p=5$
+were approximated in~\cite{BoKa} with a standard Newton scheme.
+The solutions in~\cite{BoKa} were not very accurate, 
+but since singularity formation of the evolution
+equations were under study,
+the accuracy of the solitary-wave approximation was not an important issue.
+The problem with the method of~\cite{BoKa} and some other works was that the fft
+used there was not purged of possible symmetries (translational and reflective).
+In the current code, since a cosine formulation is chosen, these symmetries
+are automatically eliminated, and the resulting computations are able to to
+render more accurate approximations.
+
+Solitary-wave solutions of these equation could be computed with higher accuracy
+using a type of Petviashvili method in~\cite{Pelinovsky}, but here we employ
+the \textsf{SpecTraVVave} package using the boundary equation~\eqref{solitary},
+and treating solitary waves as 
+traveling waves with sufficiently long wavelength 
+that have wave trough at zero. 
+Once these high-accuracy solutions are found, they are aligned in 
+an evolution code using a high-order time integrator, and the
+interaction of two waves is studied.
+
+Two questions are investigated. First, the interaction is investigated for evidence
+of integrability. Second, we are looking for possible annihilation of one of the waves,
+such as may happen in some other evolution equations~\cite{Tjon}.
+
+A possible approach to studying the question of complete integrability
+is analyzing the interaction of two solitary wave solutions of the equation, 
+such as carried out in~\cite{HK,HKJB} for other nonlocal equations.	
+In \autoref{solitons-full}, snapshots of interaction of two solitary waves at different times are shown.
+The time difference between two consecutive snapshots is constant.    
+As it may be observed, during the process of interaction, the two initial solitary waves combine 
+into a single wave, and an additional oscillation is produced. 
+This leads us to the conclusion that the interaction of solitary waves is not elastic 
+and the modified Benjamin--Ono equation may not be integrable. 
+In addition, it appears that the smaller wave disappears as most of its
+mass is acquired by the larger wave. Thus one may argue that the small
+wave is annihilated by the larger wave. It can also be observed that
+the larger wave starts growing, and it is likely that this growth
+will lead to finite-time blow-up. This question was not investigated further
+since blow-up phenomena have already been studied closely in~\cite{HKJB}.
+Indeed, very recently, the finite-time blow-up 
+has been proven mathematically in~\cite{MP}.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[t]
+\centering
+		\includegraphics[width=0.6\textwidth]{soliton_inter}      
+\caption{\small Interaction of two solitary waves of the modified Benjamin--Ono equation.}
+\label{solitons-full}
+\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Effect of competing dispersion in the Benjamin equation}
+\label{Benjamin_eq}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The Benjamin equation was found by Benjamin~\cite{B3} as a model for two-layer flow in the case
+when the interface is subject to surface tension. The approximation may not be a good model for
+a stratified situation, but more applicable to the case where two fluids are separated by a
+sharp interface. 
+The equation is 
+\begin{equation}\label{benjamin}
+u_t + u_x + uu_x - \mathcal{H} u_{xx} - \tau u_{xxx} = 0,
+\end{equation}
+where $\tau$ is a parameter similar to the Bond number in free surface flow~
+\cite{B3}, \cite{KalischBenjamin}, \cite{Walsh2014a}.
+
+%%%%%%%%\vskip 0.05in
+
+In this section, a study relating to the effects of competing dispersion operators 
+on the shape of periodic traveling waves in the Benjamin equation is presented.
+An in-depth study of solitary waves was carried out in~\cite{DDM}.
+As will come to light, the periodic case features some new phenomena, such as secondary
+bifurcations, connecting and crossing branches.
+For the purpose of this study, we fix the parameter $\tau=0.1$,
+so that the dispersion relation for the linearized equation is
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+c(k) = 1 - |k| +  0.1  k^2. 	\label{Ben_disp_rel}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+Traveling wave solutions with full wavelengths $L_1 = \pi/5$, $L_2 = 4\pi /19$ and $L_3 = 4\pi$ 
+are computed for the equation~\eqref{benjamin}.  
+The corresponding wavenumbers are $k_1 = 10$, $k_2 = 19/2$, and $k_3 = 0.5$, respectively.
+A plot of the dispersion relation~\eqref{Ben_disp_rel} is given in \autoref{disp-rel}.
+Bifurcation branches of traveling wave solutions with the selected wavelengths are given in \autoref{fig:ben-bifur}.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[ht]
+  \centering
+\includegraphics[width=0.47\textwidth]{{disp_rel_plot}}  \caption{Dispersion relation \eqref{Ben_disp_rel}}
+\label{disp-rel}
+\end{figure}
+
+\begin{figure}
+  \centering
+\includegraphics[width=0.47\textwidth]{Benj_bifur_plot}
+\caption{Bifurcation curves of the equation~\eqref{benjamin} with different wavelengths, $\blacktriangle$ --  points of bifurcation, $\bullet$ -- selected solutions (see \autoref{long-solutions}).}
+\label{fig:ben-bifur}
+\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+The branch denoted by $L_1$ originates at the bifurcation point 
+located at $c = 1$ and zero waveheight.
+The branches denoted by $L_2$ and $L_3$ originate from the same bifurcation point,
+located at $c = 0.525$ and zero waveheight. 
+These two branches continue in different directions, due to differences in wavelength.
+In particular, the $L_3$ branch contains waves with shorter wavelengths, and falls into the  
+capillary regime. On the other hand, the $L_2$ branch falls in the gravity regime.
+As the waveheight grows, solutions on the $L_3$ branch first cross the $L_1$ branch without
+connecting. Additional oscillations develop in the solutions until 
+a new fundamental wavelength $\pi/5$ is reached, and the branch terminates
+as it connects to the $L_1$ branch. The situation is depicted in \autoref{long-solutions}. 
+The point where the $L_1$ and $L_3$ branches meet
+is approximately $(c^*,a^*) = (0.945, 5.938)$.
+The corresponding profiles essentially overlap, as shown on \autoref{last-solutions}.
+This point can also be interpreted as a secondary bifurcation point 
+of the $L_1$ branch, where solutions with wavelengths that are multiples of $\pi/5$ develop. 
+We should note that similar phenomena concerning crossing and connecting
+branches were previously observed in \cite{RK}
+for the Whitham equation with surface tension which was introduced in~\cite{HurJohnson}.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[ht]
+\begin{subfigure}
+	\centering
+	\includegraphics[width=0.47\textwidth]{left_long_solutions}      
+	    \end{subfigure}%
+~%
+\begin{subfigure}
+        \centering
+		\includegraphics[width=0.47\textwidth]{right_long_solutions}{}      
+    \end{subfigure}%
+	\caption{\small Selected solutions of the equation~\eqref{benjamin}. 
+Labels preserved as shown in \autoref{fig:ben-bifur}.}
+\label{long-solutions}
+\end{figure}
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[ht]
+        \centering
+		\includegraphics[width=0.47\textwidth]{last2}{}      
+	\caption{\small Solution profiles at the point $(c^*, a^*)$ where the $L_1$ and $L_3$ branches meet (see \autoref{fig:ben-bifur}).}
+\label{last-solutions}
+\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%                                                  %%%%%%%%%%%%%%%%%
+%%%%%%%%%%                                                  %%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Conclusions and future work}
+\label{sec:conclusion}
+
+The numerical algorithm of \textsf{SpecTraVVave} features ample flexibility for researching 
+different aspects of nonlocal dispersive wave equations and their traveling wave solutions.
+The solver package is simpler in use when compared with programs such as \textsf{AUTO} and 
+\textsf{Wavetrain}, however it does not have the same level of generality. 
+Moreover, \textsf{AUTO} and \textsf{Wavetrain} are programmed in low-level programming languages
+and will therefore run more efficiently.
+\textsf{SpecTraVVave} is implemented in an object-oriented fashion~\cite{PyBook}, 
+which makes the program easily expandable.
+\textsf{IPython} provides means for interactive work with the package, 
+and enables users to create convenient notebook-programs.
+A parametric approach in defining amplitude and phase speed makes it possible to follow
+turning points on bifurcation curves. Specification of different boundary conditions allows 
+computing solutions with certain features, such as traveling waves with zero mean, 
+or approximations to solitary waves.
+
+In this work, the \textsf{SpecTraVVave} package has been put to use for the study of 
+a number on nonlinear evolution equations: the Whitham equation, the modified Benjamin--Ono equation
+and the Benjamin equation.
+For the chosen set of parameters, experiments on the Whitham equation resulted in numerical 
+confirmation of the conjecture on cusped solutions. It was also possible to identify the point of stability 
+inversion of traveling wave solutions of the equation and the termination point of its bifurcation curve.
+
+In case of the modified Benjamin--Ono equation, the study on solitary wave solutions lead us to conclude that 
+interaction process ended with annihilation of one of the two waves. 
+The experiment on the Benjamin equation showed one more example of the effect of competing dispersion.
+As the amplitude increased, traveling wave solutions of wavelength $4\pi$ developed
+additional oscillations, and later connected up with a branch of solutions with
+wavelength $\pi/5$.
+
+Future work on the \textsf{SpecTraVVave} package will be focused on development of its functionality and broadening the range of problems that can be studied.
+Possible extensions may include implementation of algorithms based on the Petviashvili method~\cite{Duran2014CNSNS}--\cite{Duran2014JCAM} and generalization to systems of equations.
+
+%\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\textbf{Acknowledgments.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+The authors would like to thank Mats Ehrnström and Erik Wahlén
+for fruitful discussions on the subject of the current paper. 
+This research was supported by the Research Council of Norway on grant no. 213474/F20
+and by the J C Kempe Memorial Fund on grant no. SMK-1238.
+
+\appendix
+\section*{Appendices}
+\addcontentsline{toc}{section}{Appendices}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Computing initial guesses from Stokes expansion.}
+\label{sec:stokes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+u_t+\left[f(u)\right]_{x} + \mathcal{L}u_x=0, \label{gen}
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 The goal of this section is to explain how the idea of Stokes's approximation works in providing the initial data (guess) on wave and phase velocity for solving the equation~\eqref{gen} numerically.
+
+	We will consider $\Lop$ being linear and self-adjoint Fourier multiplier operator, and a function $f$ that has degree of zeros $p \geq 2$:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+\widehat{\Lop u}(k) &= \alpha(k) \cdot \widehat{u}(k),\\
+\langle\Lop u, v \rangle_{L^2(0,L)} &= \langle u, \Lop v\rangle_{L^2(0,L)}, \\
+p &= \min_{k \in \NN}{f^{(k-1)}(0) =0 \text{ and } f^{(k)}(0)\neq 0}
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 Consider the equation~\eqref{gen} and its solution in the form $u(x,t) = \phi(x-ct)$, which is a traveling wave solution. Inserting $\phi(x-ct)$ into~\eqref{gen} leads to the equation:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+-c\phi' + f'(\phi)\phi'+\mathcal{L}\phi'= 0,
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 which can be integrated to give:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+-c\phi + f(\phi)+\mathcal{L}\phi= B, \quad B = \mathrm{const}. \label{int_gen-B}
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	Consider $B = 0$ in equation~\eqref{int_gen-B}, and expansions of $\phi$ and $c$:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+\phi = \xi = \varepsilon\xi_1 + \varepsilon^2\xi_2+ \cdots, \label{fi_exp}\\
+c = c_0 + \varepsilon c_1 + \varepsilon^2 c_2+\cdots. \label{c_exp}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 The next step is to insert~\eqref{fi_exp} and~\eqref{c_exp} to the equation~\eqref{int_gen-B} and write out the terms at powers of $\varepsilon$. 
+	The function $f(\phi)$ is expanded around zero and, therefore, will appear only in $\varepsilon^p$ terms. 
+	Thus, the term at the first power of $\varepsilon$ reads: 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+\varepsilon: \quad -c_0\xi_1 + \mathcal{L}\xi_1 = 0, \label{epsilon1}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 Hence, $c_0$ is an eigenvalue of the operator $\Lop$, regarded as defined on $L$-periodic functions. 
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 Taking the Fourier transform of the equation~\eqref{epsilon1} gives:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+-c_0\widehat{\xi_1}(k) + \alpha(k)\widehat{\xi_1}(k) = 0 . \label{eps1-transformed}
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\quad\rightarrow\quad \widehat{\xi_1}(k) \equiv 0 \mbox{ or } (-c_0+\widehat{\mathcal{L}}(k)) = 0. 
+	 The equation~\eqref{eps1-transformed} has two trivial solutions: either $\xi_1(k) \equiv 0$ or $\alpha(k) \equiv c_0$.
+If we assume non-trivial $\xi_1$ and $\alpha(k) \neq \text{const}$, the following solves the problem:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+  \widehat{\xi_1}(k) = 2\pi\delta(k - k_0), \quad \text{and} \quad c_0 = \alpha(k_0), \label{solution-xi}
+\end{align}
+	 for some $k_0 \in \R$. 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	Since $\xi_1$ is the first-order approximation to $\phi$, the corresponding wave number should be equal to 1.
+The $L$-periodicity condition entails that $k_0 = 2\pi/L\cdot 1$.
+The spatial variable $x$ has to be scaled to $x'= L/2\pi x$, accordingly.
+From the solutions in~\eqref{solution-xi} we have
+\begin{equation}
+	\xi_1(x') = e^{ik_0x'} = \cos(k_0x') + \sin(k_0x').
+\end{equation}	
+Considering the projection onto the space $\SN$, we are led to choose $\xi_1(x') = \cos(k_0x')$.
+%
+%
+%
+
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 For further analysis, let us define an operator $\Aop$ 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+	\Aop :=  - c_0 \mathcal{E} + \Lop,
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+where $\mathcal{E}$ is the identity operator. The operator $\Aop$ inherits the property of being 
+self-adjoint from $\Lop$. Moreover, it follows from~\eqref{epsilon1} that $\Aop \xi_1 = 0$ and $\xi_1 \in \ker(\Aop)$. 
+If $p>2$ then $f''(0)=0$ and the terms at $\epsilon^2$ are:
+\begin{equation}
+\Aop \xi_2 - c_1\xi_1 = 0.
+\end{equation}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 Taking scalar multiplication of the latter with $\xi_1$, one obtains:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+	&\langle\xi_1, \Aop\xi_2 \rangle_{L^2(0,L)} = c_1 \| \xi_1\|^2_{L^2(0,L)}, \\
+	&\langle\xi_1, \Aop\xi_2\rangle_{L^2(0,L)} = \langle\Aop \xi_1, \xi_2\rangle_{L^2(0,L)} = \langle0, \xi_2\rangle_{L^2(0,L)} = 0.
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 As a result, one has $c_1 \| \xi_1\|^2_{L^2(0,L)} = 0$ and, hence, $c_1 = 0$.
+	Repeating the same argument, it becomes clear that $c_k = 0$ for any $k \leq p-1$.
+	Besides that, $\xi_2$ is in the kernel of $\Aop$, so it may be assumed to be proportional to $\xi_1$.
+	The terms at order $\varepsilon^p$ are:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+  \Aop \xi_p - c_{p-1}\xi_1 + \frac{f^{(p)}(0)}{p!} \xi_1^p = 0. \label{epsp}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 Let us denote for brevity
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+	f_p := \frac{f^{(p)}(0)}{p!}.
+\end{align*}
+%%%%%%%%%			%%%%%%%%%%%%%%%
+	 Pairing~\eqref{epsp} with $\xi_1$ (and assuming $\| \xi_1 \|_{L^2(0,L)} = 1$) gives
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+	c_{p-1} = f_p	\cdot \langle\xi_1^p, \xi_1\rangle_{L^2(0,L)}, \label{cp-1}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 which gives us the value of $c_{p-1}$.
+	It only remains to solve the following problem numerically in order to obtain $\xi_p$:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+  \Aop \xi_p = c_{p-1}\xi_1 - f_p\xi_1^p \label{epsilonp}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+	For the last equation to be solved, the operator $\Aop$ has to be invertible. It is also required that $\langle\xi_1, \xi_p\rangle_{L^2(0,L)}=0$. 
+	Therefore the solution is sought in the space orthogonal to $\ker(\Aop)$.
+	Since $\Aop$ is still a Fourier multiplier operator, one can take the Fourier transform of the equation~\eqref{epsilonp} to find
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+	\widehat\Aop(k) \widehat\xi_p(k)& = c_{p-1}\widehat\xi_1(k) - f_p				
+	\widehat{\xi_1^p}(k), \\
+	\widehat\xi_p(k)& = \widehat\Aop(k)^{-1} \left(c_{p-1}\widehat\xi_1(k) - f_p	
+	\widehat{\xi_1^p}(k)\right).
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+Taking the inverse Fourier transform of $\widehat\xi_p(k)$ gives $\xi_p$.
+Since only even solutions of the problem are considered the cosine part of the Fourier transforms will be required.
+It is sufficient to use $\xi_1$ and $c_0$ as the initial guesses for the Newton method.
+However, it should be noted that for different values of $p$ the pair of parameters $\xi_p$ and $c_{p-1}$ are computed in different ways.
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{enumerate}[label=\alph*)]
+	\item If $p = 2$, then $\xi_2$ is computed from~\eqref{epsilonp}, but $c_{p-1}$ here becomes zero. Therefore one has to consider the next level of the expansion $\varepsilon^p$.
+	\item For odd values of $p$ the parameter $c_{p-1}$ can be computed from~\eqref{cp-1} and $\xi_p$ from~\eqref{epsilonp}.
+	\item For even values of $p \geq 4$ the parameter $\xi_p$ can be computed, but $c_{p-1}$ may not be non-zero in general. 
+	In such cases a different strategy of fixing the initial guess should be used.
+\end{enumerate}
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Presentation of \textsf{SpecTraVVave} and its workflow}
+\label{sec:implementation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Overview}
+There are several classes in the \textsf{SpecTraVVave} package. An overview of the program is shown 
+in \autoref{schematic}. The workflow begins with defining a flux function $f$ and the Fourier multiplier function $\alpha$ to set up an equation. 
+The traveling wave solution is characterized by the wavelength $L$ 
+and a boundary condition $\Omega(c, a, \phi_N, B)$. These parameters are fixed for a given problem.
+The defined equation is then discretized. 
+The \pyth{Discretization} object contains all required elements such as grid points, wave-numbers and the  discrete linear operator.
+
+The initial guess and the equation's residual are passed from the \pyth{Discretization} to the \pyth{Solver} object. 
+The \pyth{Navigation} object is responsible for finding good initial guesses for $c$ and $a$ that are  
+passed to the \pyth{Solver} object. 
+The \pyth{Solver} object applies Newton's method to find a solution to the system of equations~\eqref{system-N3}.
+
+The new solution is sent back to the  \pyth{Discretization} and \pyth{Navigation} objects, where variables get updated.
+All computed solutions are stored for further analysis. 
+This finishes one iteration. 
+For the next iteration the updated variables are used and a new solution is found. 
+The process may be continued as long as the Jacobian of the problem is non-singular.
+
+ %%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{figure}[ht]
+	\centering
+\includegraphics[width=0.8\textwidth]{block-scheme} 
+\caption{\small Overview of the \textsf{SpecTraVVave} package.}     
+\label{schematic}
+\end{figure} 
+ %%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ \subsection{Class Description}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We present an overview of the classes used in \textsf{SpecTraVVave} package.
+Note that, since the package is under continuous modification and development, we describe here only the  basic classes and functions the package.
+We refer to the package repository \cite{github} for up-do-date tutorials and installation instructions.
+
+ The \pyth{Equation} class is the general class for all model equations. 
+Its only role is to store a parameter $L$, the wavelength:
+
+\begin{python}
+class Equation(object):
+    def __init__(self, L):
+        self.length = L
+\end{python}
+A subclass of the \pyth{Equation} class has to implement two functions, \pyth{compute_kernel} and \pyth{flux}.
+
+The KdV model equation
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+	u_t + \frac{3}{2} u u_x + u_x + \frac{1}{6} u_{xxx} = 0
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+with $f(u) = \frac{3}{4} u^2$ and $\widehat{\Lop u}(k) = (1 - \frac{1}{6} k^2)\widehat{u}(k)$ is presented in the program as a subclass of the \pyth{Equation} class:
+\begin{python}
+class KDV(Equation):
+    def compute_kernel(self, k):
+        return 1-1/6*k*k
+
+    def flux(self, u):
+        return 3/4*u*u  
+\end{python}
+
+On can then create an object of the class \pyth{KDV} with the command:
+\begin{python}
+kdv = KDV(L=np.pi)
+\end{python} 
+%
+The solver will compute only a half of a solutions profile. The full wavelength of the solutions 
+of the defined equation will be equal to $2\pi$.
+
+In order to find solutions with specific features, boundary conditions are introduced as separate classes.
+For instance, the boundary condition specifying a constant of integration is implemented as follows:
+\begin{python}
+class Const(object):
+""" The boundary condition under which the constant of integration 
+(B) is always set to zero. """
+    def __init__(self, level=0):
+        self.level = level
+
+    def enforce(self, wave, variables, parameters):
+    """ Enforces the Const boundary condition. """
+        return np.hstack([variables[0] - self.level])
+
+    def variables_num(self):
+    """ The number of additional variables that are required to construct 
+    the Const boundary condition. """
+        return 1
+\end{python}
+A \pyth{Const} boundary condition object is created as follows:
+\begin{python}
+boundary_condition = Const()
+\end{python}
+
+The next step is to create an object of \pyth{Discretization} class, which is initialized with a model equation such as \pyth{kdv_model} and the number of grid points.
+The main parts of the class are the following:
+\begin{python}
+class Discretization(object):
+    def __init__(self, model_equation, grid_size):
+        self.equation = model_equation
+        self.size = grid_size
+
+    def operator(self, u):
+        u_ = scipy.fftpack.dct(u, norm='ortho')
+        Lv = self.fourier_multiplier()*u_
+        result = scipy.fftpack.idct(Lv, norm='ortho')
+        return result
+
+    def residual(self, u, wavespeed, const_B):
+        residual = - wavespeed*u + self.equation.flux(u) 
+                                        + self.operator(u) - const_B
+        return  residual
+\end{python}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The call \pyth{Discretization.operator(u)} computes $\Lop u $ as the inverse transform of a transformed convolution 
+\begin{equation}
+\Lop u = \mathcal{F}^{-1}[ \widehat{\Lop u}(k)] = \mathcal{F}^{-1}\left[ \alpha(k)\cdot\widehat{u}(k)\right].
+\end{equation}
+The result of the call \pyth{Discretization.residual(u, wavespeed, const_B)} is then used in the \pyth{Solver} class. 
+An object of the \pyth{Solver} class is initialized with an object of the \pyth{Discretization} class, and a boundary condition object.
+
+\begin{python}
+class Solver(object):
+    def __init__(self, discrete_problem, boundary_condition):
+        self.discretization = discrete_problem
+        self.boundary = boundary_condition
+
+    def solve(self, guess_wave, parameter_anchor, direction):
+    """ Runs a Newton solver on a system of nonlinear equations once. 
+    Takes the residual(vector) as the system to solve. parameter_anchor 
+    is the initial guess for (c,a) values and it is taken from the 
+    Navigation class. """
+        size = len(guess_wave)
+        self.discretization.size = size
+
+        def residual(vector):
+        """ Contructs a system of nonlinear equations. First part, 
+        main_residual, is from given wave equation; second part, 
+        boundary_residual, comes from the chosen boundary conditions. """
+        . . .   
+        return np.hstack([main_residual, boundary_residual, 
+                                          amplitude_residual])  
+    . . .
+    return new_wave, new_boundary_variables, new_parameter
+\end{python}
+%
+Some omitted parts in the above script are substituted by \textsf{'.
+.
+.'} sign.
+Each iteration on a \pyth{Solver} object is run from a \pyth{Navigation} object, which takes the \pyth{Solver} object for initialization.
+\begin{python}
+class Navigation(object):
+""" Runs the Solver and stores the results. """
+
+    def __init__(self, solver_object, size=32):
+    self.solve = solver_object.solve # function to run Newton method
+    self.size = size # size for navigation
+    . . . 
+    def run_solver(self, current_wave, pstar, direction):
+        new_wave, variables, p3 = self.solve(current_wave, pstar, direction)
+        return new_wave, variables, p3
+\end{python}
+
+All the above classes can be modified and developed further, new classes may be defined as well.
+
+\subsection{Detailed Workflow}
+
+The workflow with the package consists of three basic steps:
+\begin{enumerate}
+\item Once the necessary classes have been imported in the current namespace, generate all necessary objects:
+\begin{python}
+equation = KDV(L=np.pi)
+boundary_condition = Const()
+discretization = Discretization(equation, grid_size=64)
+solver = Solver(discretization, boundary_condition)
+navigator = Navigation(solver)    
+\end{python}
+\item Choose a number of iterations, i.e., the number of solutions to compute, and run the solver:
+\begin{python}
+n_iter = 50
+navigator.run(n_iter)
+\end{python}
+\item All computed solutions are stored in \pyth{navigation_object}
+\begin{python}
+last_computed = -1    
+wave_profile = navigator[last_computed]['solution']
+wave_speed = navigator[last_computed]['current'][0]
+wave_amplitude = navigator[last_computed]['current'][1]
+\end{python}
+\end{enumerate}
+
+We further refer to the code repository \url{https://github.com/olivierverdier/SpecTraVVave} for up-to-date instructions on how to run the code.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+% For two-column wide figures use
+%
+% For tables use
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+% BibTeX users please use one of
+%\bibliographystyle{spbasic}      % basic style, author-year citations
+%\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
+%\bibliography{}   % name your BibTeX data base
+
+% Non-BibTeX users please use
+\begin{thebibliography}{}
+%
+% and use \bibitem to create references. Consult the Instructions
+% for authors for reference list style.
+%
+
+\bibitem{ABFS} Abdelouhab, L. Bona, J.L., Felland, M. and Saut, J.-C. 
+     {\em Nonlocal models for nonlinear dispersive waves}.
+     Physica D \textbf{40},360-392 (1989)
+
+\bibitem{APR} Albert, J.P., Bona, J.L. and Restrepo, J.M. 
+{\em Solitary-wave solutions of the Benjamin equation}.
+SIAM J. Appl. Math. \textbf{59},2139--2161 (1999)
+
+\bibitem{B1} Benjamin, T.B.
+          {\em Internal Waves of permanent form in fluids of great depth}.
+          J. Fluid Mech.  \textbf{29}, 559--592 (1967)
+
+\bibitem{B2} Benjamin, T.B.
+               {\em The stability of solitary waves}.
+              Proc. Roy. Soc. London A 328, 153-183 (1972)
+
+\bibitem{B3} Benjamin, T.B. 
+          {\emph{A new kind of solitary wave}}.
+          J. Fluid Mech. \textbf{245}, 401--411  (1992)
+
+\bibitem{Duran2011JCAM} \'{A}lvarez, J. and  Dur\'{a}n, A.
+{\em A numerical scheme for periodic travelling-wave simulations in some nonlinear dispersive wave models}.
+J. Comput. Appl. Math. \textbf{235}, 1790--1797 (2011)
+
+\bibitem{Duran2014CNSNS} \'{A}lvarez, J. and  Dur\'{a}n, A.
+{\em An extended Petviashvili method for the numerical generation of traveling and localized waves}.
+Commun. Nonlinear Sci. Numer. Simul. \textbf{19}, 2272--2283 (2014)
+
+\bibitem{Duran2014JCAM} \'{A}lvarez, J. and  Dur\'{a}n, A.
+{\em, Petviashvili type methods for traveling wave computations: I. Analysis of convergence}.
+J. Comput. Appl. Math. \textbf{266}, 39--51 (2014)
+
+\bibitem{Duran2014JCAMCor} \'{A}lvarez, J. and  Dur\'{a}n, A.
+{\em Corrigendum to "Petviashvili type methods for traveling wave computations: I. Analysis of convergence},
+[J. Comput. Appl. Math. 266 (2014) 39--51]. J. Comput. Appl. Math. \textbf{277}, 215--216 (2015)
+
+\bibitem{BDKM} Bona, J.L., Dougalis, V.A.,  Karakashian, O.A. and McKinney, W.R.
+   {\em Conservative, high-order numerical schemes for the generalized
+   Korteweg--de Vries equation},
+   Philos. Trans. Royal Soc. London Ser. A, \textbf{351}, 107--164 (1995)
+
+\bibitem{BoKa} Bona, J.L. and Kalisch, H.
+              {\em Singularity formation in the generalized Benjamin--Ono equation}.
+              Discrete Contin. Dyn. Syst. \textbf{11}, 779--785 (2004)
+
+\bibitem{BSS}
+Bona, J.L., Souganidis, P.E. and Strauss, W.A.
+{\em Stability and instability of solitary waves of Korteweg--de Vries type}.
+Proc. R. Soc. Lond. A \textbf{411}, 395-412 (1987) 
+%
+%
+\bibitem{HKN} Borluk, H., Kalisch, H. and Nicholls, D.P. 
+{\em A numerical study of the Whitham equation as a model for steady surface water waves}.
+J. Comput. Appl. Math. \textbf{296}, 293--302 (2016)
+
+\bibitem{TJB}
+Bridges, T.J. 
+{\em Superharmonic instability, homoclinic torus bifurcation and water-wave breaking}.
+J. Fluid Mech. \textbf{505}, 153-162 (2004)
+
+\bibitem{Chen} Chen, H.
+{\em Existence of periodic travelling-wave solutions of nonlinear, dispersive wave equations}. 
+Nonlinearity \textbf{17}, 2041--2056 (2004)
+
+\bibitem{CB} Chen H. and Bona, J.L. 
+          {\em Existence and asymptotic properties of solitary-wave
+          solutions of Benjamin-type equations}.
+          Adv. Diff. Eq. \textbf{3},  51-84 (1998)
+
+%
+
+\bibitem{ChenWalsh2016} Chen, R.M. and Walsh, S. 
+{\em Continuous dependence on the density for stratified steady water waves}.
+Arch. Ration. Mech. Anal. \textbf{219}, 741--792 (2016)
+
+\bibitem{CC2} Choi, W. and Camassa, R.
+          {\em Fully nonlinear internal waves in a two-fluid system}.
+          J. Fluid Mech.  \textbf{396}, 1-36 (1999)
+
+\bibitem{dFSS}
+de Frutos, J. and Sanz-Serna, J. M.
+{\em An easily implementable fourth-order method for the time integration of wave problems}.
+J. Comp. Phys. \textbf{103}, 160--168 (1992)
+
+\bibitem{Doed97}
+Doedel, E.J., Champneys, A.R., Fairgrieve, T.F., Kuznetsov, Y.A., Sandstede, B. and Wang, X.-J., 
+{\em AUTO97 : Continuation and bifurcation software for ordinary differential equations}, 
+available at 
+\url{http://indy.cs.concordia.ca/auto/}, Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada, 1997.
+
+\bibitem{DDM} Dougalis, V.A., Duran, A. and Mitsotakis, D. 
+{\em Numerical solution of the Benjamin equation}.
+Wave Motion \textbf{52}, 194--215 (2015)
+
+\bibitem{EK1}
+Ehrnstr{\"o}m, M. and Kalisch, H. 
+{\em Traveling waves for the Whitham equation}. 
+Differential Integral Equations \textbf{22}, 1193--1210 (2009)
+
+\bibitem{EK2}
+Ehrnstr{\"o}m, M. and Kalisch, H. 
+{\em Global bifurcation for the Whitham equation}.
+Math. Mod. Nat. Phenomena \textbf{8}, 13--30 (2013)
+%
+\bibitem{EW2016}
+Ehrnstr{\"o}m, M. and  Wahl\'en, E. 
+{\em On Whitham's conjecture of a highest cusped wave for a nonlocal dispersive equation}. arXiv:1602.05384
+(2016).
+
+\bibitem{PyBook}
+F\"uhrer, C., Solem, J.E., Verdier, O. 
+Computing with Python: An introduction to Python for science and engineering, Pearson (2014)
+%
+\bibitem{GK} Gruji\'{c}, Z. and Kalisch, H.
+{\em Gevrey regularity for a class of water-wave models}.
+Nonlinear Analysis \textbf{71}, 1160--1170 (2009)
+
+\bibitem{Hur2015} Hur, V.M. 
+{\em Breaking in the Whitham equation for shallow water waves}.
+arXiv:1506.04075 (2015).
+
+\bibitem{HurJohnson}
+Hur, V.M. and Johnson, M. 
+{\em Modulational instability in the Whitham equation of water waves}.
+Studies in Applied Mathematics \textbf{134}, 120--143 (2015)
+
+\bibitem{HurTao} Hur, V.M. and Tao, L.
+{\em Wave breaking for the Whitham equation with fractional dispersion}.
+arXiv:1410.1570, 2014.
+
+\bibitem{FW}
+Fornberg, B. and Whitham, G. B. 
+{\em A numerical and theoretical study of certain nonlinear wave phenomena}. 
+Phil. Trans. Roy. Soc. A \textbf{289}, 373--404 (1978)
+
+%\bibitem{ipython}
+%\url{https://ipython.org/}
+
+\bibitem{HK}
+Kalisch, H. 
+{\em Error analysis of a spectral projection of the regularized Benjamin--Ono equation}.
+BIT Numerical Mathematics \textbf{45}, 69--89 (2005)
+
+\bibitem{KalischBenjamin}
+Kalisch, H.
+{\em Derivation and comparison of model equations for interfacial capillary-gravity waves in deep water}.
+Math. Comput. Simulation \textbf{74},168--178 (2007)
+%(2-3):
+
+\bibitem{HKJB}
+Kalisch, H. and Bona, J.L.
+{\em Models for internal waves in deep water}. 
+Discrete and Continuous Dynamical Systems \textbf{6}, 1--20 (2000)
+
+\bibitem{Kato} Kato, T.
+{\em On the Cauchy problem for the (generalized) Korteweg--de Vries equation}.
+Studies in applied mathematics, 93--128, Adv. Math. Suppl. Stud., \textbf{8} 	
+(Academic Press, New York, 1983).
+
+\bibitem{Keller} 
+Keller H.B., 
+{\em Numerical solution of bifurcation and nonlinear eigenvalue problems}. 
+Applications of Bifurcation Theory, Rabinowitz, P. H., Academic Press, (1977), pp. 359--384.
+
+\bibitem{LS} Lannes, D. and Saut, J.-C.
+{\em Remarks on the full dispersion Kadomtsev--Petviashvli equation}.
+Kinet. Relat. Models \textbf{6}, 989--1009 (2013)
+
+\bibitem{MM} Martel, Y. and Merle, F.
+{\em Blow up in finite time and dynamics of blow up solutions for the L2-critical generalized KdV equation}.
+J. Amer. Math. Soc. \textbf{15}, 617--664 (2002)
+
+\bibitem{MP} Martel, Y. and Pilod, D.
+{\em Construction of a minimal mass blow up solution of the modified Benjamin-Ono equation}.
+arXiv:1605.01837 (2016).
+
+\bibitem{MKD2015} Moldabayev, D., Kalisch, H. and  Dutykh, D. 
+{\em The Whitham Equation as a model for surface water waves}.
+Phys. D  \textbf{309}, 99--107 (2015)
+
+\bibitem{github}
+Moldabayev, D., Verdier, O. and Kalisch, H.
+{\em SpecTraVVave}, available at \newline
+\url{https://github.com/olivierverdier/SpecTraVVave}.
+
+\bibitem{NgK}
+Nguyen, N.T. and Kalisch, H.
+{\em Orbital stability of negative solitary waves}.
+Mathematics and Computers in Simulation \textbf{80}, 139-150 (2009)
+
+\bibitem{O} Ono, H.
+         {\emph{Algebraic Solitary Waves in Stratified Fluids}}.
+         J. Phys. Soc. of Japan \textbf{39}, 1082-1091 (1975)
+
+\bibitem{Pelinovsky} Pelinovsky, D.E. and Stepanyants, Y.A. 
+{\em Convergence of Petviashvili's iteration method for numerical approximation 
+of stationary solutions of nonlinear wave equations}.
+SIAM J. Numer. Anal. \textbf{42}, 1110--1127 (2004)
+
+\bibitem{IPython} Fernando Pérez, Brian E. Granger
+ \emph{IPython: A System for Interactive Scientific Computing}, 
+Computing in Science and Engineering, vol. 9, no. 3, pp. 21-29, May/June 2007.
+% doi:10.1109/MCSE.2007.53.
+URL: \url{http://ipython.org}
+
+\bibitem{Petviashvili} Petviashvili, V.I.
+{\em Equation of an extraordinary soliton}, 
+Sov. J. Plasma Phys. \textbf{2}, 257--258 (1976)
+
+\bibitem{RK}
+Remonato, F. and Kalisch, H.
+{\em Numerical bifurcation for the capillary Whitham equation}.
+arXiv:1604.08324 (2016).
+
+\bibitem{Sanford}
+Sanford, N., Kodama, K., Carter, J.D. and Kalisch, H. 
+{\em Stability of traveling wave solutions to the Whitham equation}. 
+Phys. Lett. A \textbf{378}, 2100--2107 (2014)
+
+\bibitem{Sherratt} Sherratt, J.A. 
+{\em Numerical continuation methods for studying periodic travelling wave (wavetrain) 
+solutions of partial differential equations}.
+Appl. Math. Comp. \textbf{218}, 4684--4694 (2012)
+
+\bibitem{Tjon} Courtenay Lewis, J. and Tjon, J.A.
+              {\em Resonant production of solitons in the RLW equation}.
+              Phys. Lett. A \textbf{73}, 275--279 (1979)
+
+\bibitem{Walsh2014a} Walsh, S. 
+{\em Steady stratified periodic gravity waves with surface tension I: Local bifurcation}.
+Discrete Contin. Dyn. Syst. \textbf{34}, 3241--3285 (2014)
+
+\bibitem{Walsh2014b} Walsh, S. 
+{\em Steady stratified periodic gravity waves with surface tension II: global bifurcation}. 
+Discrete Contin. Dyn. Syst. \textbf{34}, 3287--3315 (2014)
+
+\bibitem{Wh1} 
+Whitham, G.B. 
+{\em Variational methods and applications to water waves}.
+Proc. Roy. Soc. London A \textbf{299}, 6--25 (1967)
+
+\bibitem{Wh2}
+Whitham, G.B. 
+Linear and Nonlinear Waves. Wiley, New York, (1974)
+
+\end{thebibliography}
+
+\end{document}

### 2017-02-08 Thesis/latex/paperC.tex
paperC - next step

--- 2017-01-30 Thesis/latex/summary.tex
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Explicit solutions for a long-wave model with constant vorticity}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In nonlinear dispersive evolution equations, the competing effects of nonlinearity and dispersion 
+make a number of interesting phenomena possible. In the current work, the focus is on the numerical
+approximation of traveling-wave solutions of such equations.
+We describe our efforts to write a dedicated \textsf{Python}
+code which is able to compute traveling-wave solutions of nonlinear
+dispersive equations in a very general form.
+%\begin{equation*}
+%u_t + [f(u)]_{x} + \Lop u_x = 0,
+%\end{equation*}
+%%\vskip -0.05cm
+%where $\Lop$ is a self-adjoint operator, and $f$ is a real-valued function with 
+%$f(0) = 0$ and $f'(0) = 0$.
+
+The \textsf{SpecTraVVave} code uses a continuation method coupled with a spectral projection
+to compute approximations of steady symmetric solutions of this equation.
+The code is used in a number of situations to gain an
+understanding of traveling-wave solutions. The first case is the Whitham
+equation, where numerical evidence points to the conclusion that the main 
+bifurcation branch features three distinct points of interest, namely a turning point, 
+a point of stability inversion, and a terminal point which corresponds to a cusped wave.
+
+The second case is the so-called modified Benjamin--Ono equation where
+the interaction of two solitary waves is investigated. It is found
+that two solitary waves may interact in such a way
+that the smaller wave is annihilated. The third case concerns
+the Benjamin equation which features two competing dispersive operators.
+In this case, it is found that bifurcation curves of periodic traveling-wave
+solutions may cross and connect high up on the branch in the nonlinear regime.
+%Include keywords and mathematical subject classification numbers as needed.
+
+\subsection{Introduction}
+\label{intro}
+
+This paper is concerned with traveling wave solutions for a class of nonlinear dispersive
+equations of the form
+\begin{equation}\label{nonlocal}
+u_t + [f(u)]_x + \Lop u_x = 0,
+\end{equation}
+where $\Lop$ is a self-adjoint operator, and $f$ is a real-valued function 
+with $f(0)=0$ and $f'(0)=0$, and which satisfies certain growth conditions.
+Equations of this form arise routinely in the study of wave problems in fluid mechanics
+and many other contexts.
+A prototype of such an equation is the KdV equation that appears
+if $\Lop= I + \frac{1}{6} \partial_x^2$ and $f(u) = \frac{3}{4} u^2$.
+In the current work, the operator $\Lop$ is considered to be given as a Fourier multiplier operator,
+such as for instance in the Benjamin--Ono equation, which arises in the study of
+interfacial waves. In this case, the Fourier multiplier operator is given by
+$\Lop =  I - \Hop \partial_x$, where the Hilbert transform
+$\Hop$ is defined as
+\begin{equation}
+\Hop u(x) = \frac{1}{\pi} \, \mathrm{p.v.} \int_{-\infty}^{\infty} \frac{u(x-y)}{y} \, \dd y, 
+\qquad \widehat{\Hop u}(k) = -\ii \operatorname{sgn}(k)\widehat{u}(k).
+\end{equation}
+We also study in detail traveling wave solutions of the Whitham equation, which appears when 
+$\Lop$ is given by convolution with the integral kernel $K_{h_0}$ in the form
+\begin{equation}\label{symbol}
+\Lop u(x) = \int_{-\infty}^{\infty} K_{h_0} (y) u(x-y) \, \dd y,
+\qquad \widehat{K_{h_0}} (k) 
+= {\textstyle{ \sqrt{\frac{g\tanh (h_0 k)}{k}}}}, 
+\end{equation}
+and $f$ is the same function as in the KdV equation.
+
+The particular form of equation~\eqref{nonlocal} exhibits the competing effects of dispersion and nonlinearity, 
+which gives rise to a host of interesting phenomena. The most well known special phenomenon is the existence
+of solitary waves and of periodic traveling waves containing higher Fourier modes.
+Indeed, note that in the purely dispersive model $u_t + \Lop u_x = 0$, 
+the only possible permanent progressive waves are simple sinusoidal waves, while
+in the nonlinear model~\eqref{nonlocal} higher Fourier modes must be considered
+to obtain solutions.
+
+The order of the operator $\mathcal{L}$ appearing in~\eqref{nonlocal} 
+has a major effect on the types of solutions that may be found. 
+A higher-order operator, such as in the Korteweg--de Vries equation, 
+acts as a smoothing operator because of its effect of spreading different frequency components
+out due to a strongly varying phase speed~\cite{Kato1983}. Lower-order operators 
+such as the operator $K_{h_0}$ in~\eqref{symbol} appearing in the Whitham equation
+may allow solutions to develop singularities, such as derivative blow-up (see~\cite{Hur2015arxiv}, \cite{Tao2014})
+and formation of cusps (see~\cite{Wahlen2016}).
+
+On the other hand, highly nonlinear functions $f(u)$ may lead to $L^{\infty}$-blow-up.
+For instance, the generalized KdV equation which is written in normalized form as
+\begin{equation}
+\label{kdv}
+u_t + u^{p}u_x + u_x + u_{xxx} = 0,
+\end{equation}
+features global existence of solutions for $p=1,2,3$, 
+but the solutions blow-up in the critical case $p=4$ (the case $p \ge 5$ is open).
+In the case of the generalized Benjamin--Ono equation
+\begin{equation*}
+u_t + u^{p}u_x + u_x - \mathcal{H} u_{xx} = 0,
+\end{equation*}
+where $\mathcal{H}$ is the Hilbert transform, numerical evidence points to singularity formation 
+for $p > 2$~\cite{Bona2004}, but no proofs are available at this time.
+
+In order to study different phenomena related to equations of the form~\eqref{nonlocal} 
+and their traveling wave solutions, a \textsf{Python}-based solver package \textsf{SpecTraVVave} 
+was developed by the authors~\cite{github}. 
+The general idea behind the solver is to use a numerical continuation method~\cite{Keller1977} 
+implemented with a pseudo-spectral algorithm. 
+Similar previous projects include \textsf{AUTO}~\cite{Doedel1997} and \textsf{Wavetrain}~\cite{Sherratt2012}.
+\textsf{AUTO} is written in \emph{C}, whereas \textsf{Wavetrain} is written in \emph{Fortran}. 
+Both programs are efficient and very general, as they are able to cover a wide range of problems
+involving bifurcation analyses. However, from a user's perspective, 
+such a generality coupled with low level programming languages
+may lead to some difficulties in utilizing these programs efficiently.
+
+\textsf{SpecTraVVave} is designed to provide researchers with a simple yet effective tool 
+for investigating problems on traveling waves. The package is flexible, 
+and its functionality can be easily expanded. 
+The availability of the \textsf{IPython} notebook \cite{Perez2007} makes the solver very interactive,
+so that it should be easier for new users to get started.
+
+In order to maximize ease of use, \textsf{SpecTraVVave} was designed
+to find even solutions of~\eqref{nonlocal}.
+Symmetry of steady solutions can be proved for some of the models
+in the form~\eqref{nonlocal}, but not for all~\cite{Chen1998}.
+Some of these equations also admit non-smooth solutions,
+for instance as termination points of a bifurcation branch.
+This happens for example for the Whitham equation, which features bifurcation
+curves which terminate in a solution with a cusp~\cite{Wahlen2016}.
+One of the goals of the present paper is to investigate the
+precise nature of the termination of the bifurcation curve.
+
+The content of the paper is structured as follows. A mathematical description of the numerical method of 
+\textsf{SpecTraVVave} is given in \autoref{sec:spectral}.
+\autoref{sec:experiments} presents results of different experiments carried out with the package.
+Concluding remarks are given in \autoref{sec:conclusion}.
+A method for finding initial guesses for the solver is described in \autoref{sec:stokes}.
+\autoref{sec:implementation} contains a schematic of program and a description of its workflow.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Spectral scheme and construction of nonlinear system.}
+\label{sec:spectral}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Cosine collocation method.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+To compute traveling wave solutions to the equation~\eqref{nonlocal} the following ansatz is employed:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation*}%\label{ansatz}
+u(x,t) = \phi(x-ct).
+\end{equation*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Thus, the equation takes the form
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation*}%\label{phi-nonlocal}
+\phi' + \left[f(\phi) \right]' + \Lop \phi' = 0,
+\end{equation*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+which can be integrated to give
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation}\label{int_gen1}
+-c \phi + f(\phi)  + \Lop \phi = B. 
+\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The constant $B$ is a priori undetermined. One may set the $B$ equal to 
+zero as a way of normalizing the solutions. Another option is to impose
+an additional condition, for example that the integral of $\phi$ over
+one wavelength be zero. In this case, $B$ will be found along with the solution $\phi$.
+
+We consider $\Lop$ as a Fourier multiplier operator with symbol
+$\alpha(k)$. We also assume that $f$ is at least twice differentiable,
+and we have $f(0)=0$ and $f'(0)=0$.
+When computing traveling-wave solutions we focus on even periodic solutions. 
+While it can be proved in some cases that solutions of~\eqref{int_gen1}
+must be even, this is not known for a general operator $\Lop$.
+Nevertheless, we make this assumption here in order to
+make the numerical procedure as uniform as possible.
+For even periodic solutions, 
+one may use a cosine collocation instead of a Fourier method. In particular, using
+the cosine functions as basis elements automatically removes the inherent symmetries
+due to reflective and translational symmetry. 
+Moreover, the number of unknowns is reduced by a factor of $2$, 
+and the problem of the asymmetric arrangement of nodes in the FFT
+is circumvented. Of course, all these problems could also be dealt with a collocation
+method based on the Fourier basis, but the cosine basis does all of the above automatically.
+In addition,  the \textsf{Python} cosine transform is based on an integrated algorithm, which
+relies on an optimized version of the discrete cosine transform (DCT).
+
+The following description of computation scheme was presented in detail in~\cite{Ehrnstrom2013}, 
+but we will briefly repeat it here for consistency of the manuscript. For the purpose of clarity, 
+we will refer to full wavelength $L$ of a solution as the (full) wavelength, 
+and half of fundamental wavelength will be called half-wavelength. 
+Such a definition is required because the method computes a half of a solution profile, 
+the other half is automatically constructed due to symmetry.
+
+Traveling wave solutions to the equation~\eqref{int_gen1}
+are to be computed in the form of a linear combination of cosine functions of different
+wave-numbers, i.e., in the space
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+\SN = \mathrm{span}_{\R}\setc{\cos(lx)}{0 \leq l \leq N-1 }
+.
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+This is a subspace of $L^2(0,2\pi)$, and the collocation points 
+$x_n = \pi \frac{2n -1}{2N}$ for $n = 1,\dots, N$ are used to discretize the domain.  
+If the required full wavelength of solutions is to be $L \neq 2\pi$, 
+one can use a scaling on the $x$-variable. 
+Defining the new variable
+\begin{equation*}
+x' = \frac{L}{2\pi}x, 
+\end{equation*}
+yields collocation points $x_n'$ and wavenumbers $\kappa_l$ defined by
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+x_n' = \frac{L}{2} \frac{2n -1}{2N}, \quad\quad \kappa_l = \frac{2\pi}{L}l. 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+We are seeking a function $\phi_N \in \SN$ that satisfies the equations
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+-c\phi_N(x_n') + f(\phi_N)(x_n') + \Lop^N\phi_N(x_n') = 0, \label{int_gen-disc}
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+at the collocation points $x_n'$.
+The operator $\Lop^N$ is the discrete form of the operator $\Lop$, 
+and $\phi_N$ is the discrete cosine representation of $\phi$
+which is given by
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+\phi_N(x') &= \sum_{l = 0}^{N-1} \omega(\kappa_l)\Phi_N(\kappa_l)\cos(\kappa_{l}x'),\\
+\omega(\kappa_l) &= 
+\begin{cases}
+\sqrt{1/N}, \quad \kappa_l = 0, \\
+\sqrt{2/N}, \quad \kappa_l > 0,
+\end{cases}                  \\
+\Phi_N(\kappa_l) &= \omega(\kappa_l) \sum_{n = 1}^{N} \phi_N(x'_n) \cos(\kappa_l x'_n),
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+where $\kappa_l = 0,\frac{2\pi}{L},\ldots, \frac{2\pi}{L}(N-1)$ are the scaled wavenumbers,
+and $\Phi_N(\cdot)$ are the discrete cosine coefficients.
+As the equation~\eqref{int_gen-disc} is enforced at the collocation points $x_n'$, 
+one may evaluate the term $\Lop^N\phi_N$ using the matrix $\Lop^N(i,j)$ 
+defined by
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+\Lop^N\phi_N(x'_i) &= \sum_{j=1}^{N}\Lop^N(i,j) \phi_N(x'_j),\\
+\Lop^N(i,j) &= \sum_{l=0}^{N-1} \omega^2(\kappa_l) \alpha(\kappa_l) \cos(\kappa_l x'_i) \cos(\kappa_l x'_j),
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+where $\alpha(\cdot)$ is the Fourier multiplier function of the operator $\Lop$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Construction of nonlinear system.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+The equation~\eqref{int_gen-disc} enforced at $N$ collocation points 
+yields a nonlinear system of $N$ equations in $N$ unknowns, which can be written in shorthand as  
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+	\begin{equation*}
+		F(\phi_N) = 0.
+	\end{equation*}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+This system can be solved by a standard iterative method, such as Newton's method. 
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+	In this system, the value of phase speed $c$ has to be fixed for computing one particular solution. 
+	Such an approach becomes impractical when a turning point on the bifurcation curve appears.
+
+	In \textsf{SpecTraVVave} a different approach is employed: both the amplitude $a$ and the phase speed $c$ of a solution are treated as functions of a parameter $\theta$: $ a = a(\theta)$, $c = c(\theta)$.
+The parameter $\theta$ is to be computed from the system~\eqref{system-N3}.
+This construction makes it possible to follow turning points
+on the bifurcation branch with relative ease.
+	Having computed two solutions, i.e., two points on the bifurcation curve $P_1 = (c_1, a_1)$ and $P_2 = (c_2, a_2)$, one may find a direction vector $\mathbf{d}=(d^c, d^a)$ of the line that contains these points:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	\begin{align*}
+		\mathbf{d}:	\quad d^c = c_2 - c_1, \qquad	d^a = a_2-a_1.
+	\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	Then the point $P_3 = (c_3,a_3)$ is fixed at some (small) distance $s$ from the point $P_2$ in the direction $\mathbf{d}$.
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	\begin{align*}
+		P_3: \quad c_3 = c_2 + s \cdot d^c, \qquad	a_3 = a_2 + s \cdot d^a.
+	\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 The point $P_3$ plays the role of the initial guess for velocity and amplitude when computing the next solution $P_\ast = (c_\ast, a_\ast)$. 
+	The solution point $P_\ast$ is required to lay on the line with direction vector $\mathbf{d_{\bot}}=(d^c_{\bot}, d^a_{\bot})$, which is orthogonal to the vector $\mathbf{d}$.
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	\begin{align*}
+	\mathbf{d_{\bot}}: \quad d^c_{\bot} = - d^a, \qquad &d^a_{\bot} = d^c,  \\
+	P_\ast: \quad c_\ast = c_3 + \theta d^c_{\bot} \qquad &a_\ast = a_3 + \theta d^a_{\bot}.
+	\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 A schematic sketch of finding a new solution $P_\ast$ is given in Figure 1.
+
+	The variable $\theta$ is computed by Newton's method from the extended system 
+%%%%%%%%%===========%%%%%%%%%%%%%%%	
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+F\left(
+\begin{matrix}
+\phi_N(x_1)\\ \vdots \\ \phi_N(x_N) \\ B \\ \theta
+\end{matrix}
+ \right)= \left(
+\begin{matrix}
+(-c+\mathcal{L}_N)\phi_N(x_1)+ f(\phi_N(x_1)) - B
+\\ \vdots 
+\\ (-c+\mathcal{L}_N)\phi_N(x_N)+ f(\phi_N(x_N)) - B
+\\ \Omega(\phi_N, c, a, B)
+\\ \phi_N(x_1)-\phi_N(x_N)-a
+\end{matrix}
+ \right)=
+\left(
+\begin{matrix}
+0
+\\ \vdots 
+\\ 0
+\\ 0
+\\ 0
+\end{matrix}
+ \right). \label{system-N3} 
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	Here a nonhomogeneous problem ($B \neq 0$) is considered.
+	The equation 	
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+	\phi_N(x_1)-\phi_N(x_N)-a = 0, 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 makes the waveheight of the computed solution to be that of $a$. The equation 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+	\Omega(\phi_N, c, a, B)=0, %\label{boundary}
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 is called the  \emph{boundary condition}. 
+	It allows to enforce different specifications on the computed traveling wave solution.   	
+	For example, if one sets  
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+	\Omega(\phi_N, c, a, B)= \phi_N(x_1)+\cdots+\phi_N(x_N), 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 then \emph{the mean of the computed wave over a period will have to be equal to zero}.
+	One may also experiment with	
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+	\Omega(\phi_N, c, a, B) = B, 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+	 to consider the homogeneous problem $(B=0)$.
+	It can be also interesting to set
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation}
+	\Omega(\phi_N, c, a, B) = \phi_N(x_N). \label{solitary}
+\end{equation}
+%%%%%%%%%===========%%%%%%%%%%%%%%%	
+	 This enables us to compute traveling wave solutions that mimic solitary wave solutions. 	
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}
+%\centering
+%		\includegraphics[width=0.35\textwidth]{navig}
+%\caption{\small Navigation on the bifurcation curve.}
+%\label{navig}
+%\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+\subsection{Convergence.}
+In order to test the numerical implementation of the discretization,
+the method is applied to a case where the solution is known.
+One such case is the KdV equation
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+u_t + u_x + \frac{3}{2} u u_x + \frac{1}{6}u_{xxx} = 0, 
+%\label{KdV_sol*}
+\end{align*}
+which has a known solution, given in the form
+\begin{align*}
+u_{\mathrm{exact}}(x,t) = a \sech^2\Big( {\textstyle\sqrt{\frac{3a}{4}}} (x-c t)\Big),\\
+\end{align*}
+with $c=1+a/2$.
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%
+Using the boundary equation~\eqref{solitary}, \textsf{SpecTraVVave} is capable of computing 
+approximations to solitary wave solutions of nonlinear wave equations. 
+Solitary wave solutions are treated as traveling waves with sufficiently long wavelength 
+that have the wave trough at zero.
+In case of the KdV equation solitary wave solutions have exponential decay, 
+and therefore, considering the symmetry of solitary solutions, 
+the half-wavelength of $30$ is considered for the comparison. 
+%\itodo[DM]{as mentioned before: should we call the (full-wavelength )/2 as half-wavelength OR just wavelength. 
+%This may lead to confusions for reviewers.}
+Approximation errors are summarized in \autoref{t2}.
+%
+\begin{table}  [ht]
+\centering
+\begin{tabular}{rrrr}
+  \toprule
+Nb.\ of grid points & $\log_{10}(\|u_{\mathrm{exact}} - u\|_{L^{\infty}})$ & $\log_{10}(\|u_{\mathrm{exact}} - u\|_{L^2})$ & Ratio of $L^2$-errors\\
+\midrule
+32 & $-1.389$  & $-2.092$ & \\
+
+64 & $-3.705$  & $-4.549$ & $286.8$\\
+
+128 & $-8.809$  & $-9.508$ & $90935.0$\\
+
+256 & $-15.353$  & $-16.144$ & $4329670.9$\\
+
+512 & $-15.353$  & $-16.087$ & $0.9$\\
+\bottomrule
+\end{tabular}
+\caption{Estimates of error between the exact and computed solitary wave solutions for the KdV equation. Half-wavelength $30$, waveheight $a=1.2651$}
+\label{t2}
+\end{table}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=0.47\textwidth]{convergence}      
+%\caption{\small Graph of error estimates given in \autoref{t2}.}
+%\label{L2convergence}
+%\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiments with SpecTraVVave.}
+\label{sec:experiments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Termination of the waveheight-velocity bifurcation curve of the Whitham equation.}
+\label{sec-W-term}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The waveheight-velocity bifurcation curve of the Whitham equation 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+	u_t + \frac{3}{2} u u_x + Ku_x = 0, \qquad 
+	\widehat{Ku}(k) = \sqrt{\frac{\tanh(k)}{k}},  \label{W}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+was studied numerically in~\cite{Ehrnstrom2013}. 	
+An attempt was made to identify the termination point of the Whitham bifurcation curve. 
+The investigation was limited by computational tools and complete results were not obtained. 
+In particular, the authors could not confirm that traveling wave solutions do not exist 
+past the point where the authors, based on pioneering work of Whitham
+\cite{Whitham1967j} suspected a cusped solution.
+In this section a number of numerical results on nature of the bifurcation curve
+for the Whitham equation are presented.
+Solutions to the equation~\eqref{W}
+are computed in the form of traveling waves $u(x,t) = \phi(x - c t)$
+and the homogeneous $(B = 0)$ integrated version the equation is considered:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+-c \phi + \frac{3}{4} \phi^2 + K \phi = 0. \label{W-int}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+Special attention is given to relation between stability of solutions 
+and their waveheight and velocity parameters, i.e., their position on the bifurcation curve. 
+The following questions are under study: 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{enumerate}[label=\alph*)] % chktex 9 chktex 10
+\item Where does the bifurcation curve terminate?
+
+\item Where on the bifurcation curve do solutions change their stability?
+
+\item Is there any role that the turning point on the bifurcation curve plays?
+\end{enumerate}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\subsubsection{Implementation of the problem in SpecTraVVave.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The results presented here focus on $2\pi$-periodic solutions to the equation~\eqref{W-int}, 
+i.e., solutions of the system~\eqref{system-N3}. 
+Figure 3 presents Whitham bifurcation curves with numbers of grid points 
+$N = 512$, $N=1024$ and $N=2048$. 
+The current implementation of the \textsf{SpecTraVVave} package allows fixing the number of grid points $N$ 
+and a so-called doubling parameter $\DD$, 
+i.e., the number by which $N$ is doubled as computations are made. 	
+This allows us to get sets of solutions with $N, 2N, \ldots, 2^{\DD} N$ grid points. 
+If $\DD = 1$ then only two sets of solutions are computed and they are regarded as 
+lower grid (lower resolution) and higher grid (higher resolution) solutions.
+While the system~\eqref{system-N3} is processed by Newton solver, 
+lower grid solutions are taken as initial guesses for higher grid solutions. 
+All curves shown in this manuscript have been produced
+after tests with a number of resolutions were run, 
+and the curves shown did not change significantly under further refinement.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=0.48\textwidth]{512_1024_2048}{}      
+%\caption{\small Whitham bifurcation curve in different grid resolutions.}
+%\label{bif-curve-all}
+%\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%\subsubsection{Specific solutions and their stability.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+Figure 4(a) % chktex 8
+presents the Whitham bifurcation curve computed by \textsf{SpecTraVVave} 
+with $N=1024$ and $\DD=1$. 
+There are three solutions which deserve to be singled out:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{enumerate}
+	\item Traveling wave solution with minimum velocity (rhombus);
+	\item Traveling wave solution with maximum $L^2$-norm (circle);
+	\item Cusped traveling wave solution (square).
+\end{enumerate}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+Profiles of the above listed solutions are given in Figure 5(a).
+The solution with minimum velocity corresponds to the turning point of the bifurcation curve.
+The solution with maximum $L^2$-norm is very close to the latter one, although it has a higher waveheight 
+and a different velocity.
+The solution marked by a square is called here the \emph{terminal solution}.
+As already mentioned, previous studies, such as~\cite{Ehrnstrom2009,Ehrnstrom2013} did not provide any conclusive
+analysis on the part of the bifurcation curve past the turning point.
+In particular, it was not clear whether solutions ceased to exist
+at or after the turning point, or whether solutions were stable
+or unstable after the turning point.
+
+Let us first focus on the stability of solutions.
+Note that \textsf{SpecTraVVave} has an evolution integrator routine, 
+which enables one to check the stability of computed solutions. 	
+The current version of the package uses the fourth-order method developed in~\cite{DeFrutos1992}.
+In addition one may use a more refined analysis, resting on the evaluation
+of invariant functionals. This analysis is based on the observation
+that the traveling waves can be thought of as solutions of a constrained
+minimization problem. This analysis is based on ideas developed by
+Boussinesq, first exploited in~\cite{Benjamin1972j}, and later used 
+in~\cite{Bona1987,Bridges2004,Nguyen2009}, and many other works.
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[ht]
+%\centering
+%\subfigure[][Bifurcation curve in the wavespeed-waveheight parameter space]{\includegraphics[width=0.48\textwidth]{1024_2048}{}\label{sfig:bif-ch}}%
+%~%
+%\subfigure[][Bifurcation curve in a wavespeed-$L^2$-norm diagram]{\includegraphics[width=0.48\textwidth]{L2_1024_2048}{}} 
+%\caption{\small
+%Whitham bifurcation curves for $2\pi$-periodic solutions. 
+%% $N = 1024, \mathcal{D}= 1$. 
+%}
+%\label{bif-L2-1024}
+%\end{figure} 
+%trim = 1.2cm 4.7cm 0.6cm 4.5cm, clip,
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+Let us define two functionals $V$ and $E$:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+	V( \phi) = \frac{1}{2} \int_{-\infty}^{+\infty} \phi^2\,\dd\zeta, \qquad
+	E( \phi) = \int_{-\infty}^{+\infty} \left\{ \Sfrac{1}{2} \phi^3 - \phi \, K\!  \phi \right\}\, \dd\zeta.
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+The equation~\eqref{W-int} can be then written in terms of variational derivatives of $E$ and $V$
+as
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+	E' ( \phi ) - c V ' ( \phi ) = 0.
+\label{zero}
+\end{align} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+It is known from~\cite{Bona1987} that under certain conditions,
+the stability of solitary wave solutions depends 
+on convexity of the function $d(c) = E(\phi) - c V(\phi)$. 
+Solutions with values of $c$ for which $d''(c) > 0$  are stable solutions,
+and solutions with wave speeds for which $d''(c) <0$ are unstable solutions.
+
+The current numerical investigation can therefore be interpreted as an indication of the
+stability properties of traveling wave solutions of the Whitham equation.
+Note that differentiation of $d(c)$ yields
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+	d '(c) &= \underbrace{E'(\phi) - c V'(\phi)}_{=0} - V(\phi).
+\end{align*}
+Using~\eqref{zero} as indicated yields
+\begin{align}
+	d '(c) &= - V(\phi) = - \frac{1}{2} \int_{-\infty}^{+\infty} \phi ^2\,\dd\zeta = - \frac{1}{2}\|\phi\|_{L^2}^2 . 
+\label{d'c}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+Therefore, in order to understand the convexity of $d(c)$, it is sufficient
+to find points of maximum $L^2$-norm on the curve in the right panel
+of Figure 4.
+It is straightforward to see that $d''(c)$ changes sign in the neighbourhood of the maximum point 
+of this curve, i.e., around the solution with maximum $L^2$-norm. 
+In particular, $d''(c) > 0$, i.e., solutions are stable to the left of the maximum point, 
+and $d''(c) < 0$, i.e., solutions are unstable to the right of the maximum point.
+
+In addition, the solutions were tested with the evolution integrator to confirm their stability/instability in time. 
+The solution with maximum $L^2$-norm and those on the left-hand side were always found to be stable in 
+the time-dependent computations.
+Solutions on the right-hand side do not preserve their shape and thus are unstable. 
+Examples are given in Figure 6. 
+This analysis confirms that 
+the point corresponding to the minumum wave speed (the turning point), 
+and the point of stability inversion are two distinct points on the bifurcation curve.
+Moreover, the point of stability inversion
+is a little further up the branch from the turning point.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[t]
+%\centering
+%\subfigure[][Profiles of the solutions singled out in \autoref{bif-L2-1024}\subref{sfig:bif-ch}]{\includegraphics[width=0.48\textwidth]{1024_waves_1}{}\label{sfig:prof-sing}}
+%~%
+%\subfigure[][Profile after terminal solution]{\includegraphics[width=0.48\textwidth]{1024_waves_2}{}\label{sfig:prof-term}}
+%\caption{\small Profiles of specific traveling wave solutions.}
+%\label{3waves}
+%\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+Next, we turn our attention to the analysis of the terminal point. There
+are two main questions. Does the branch terminate, and if so, 
+does the terminal point on the branch correspond to a cusped traveling wave.
+First of all, note that the solution, which is computed by \textsf{SpecTraVVave}, 
+past the terminal solution has two crests, no matter how small the 
+stepping on the bifurcation branch is taken.
+(see Figure 5(b)).   
+Secondly, as will be explained presently, the relation 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{equation*}
+\frac{c}{\sup_{x} \phi(x)} = \frac{3}{2} 
+\end{equation*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+holds for the terminal solution with a good degree of approximation.
+For the most accurate runs, we obtain $c / \sup_{x \in \R} \phi(x)$ $\approx$ $1.51$. 
+%
+To explain how this relation comes about note that the steady integrated form of the
+Whitham equation can be written as
+\begin{equation}\label{Whitham_Alt}
+\Big( \frac{c}{\sqrt{3}} - \frac{\sqrt{3}}{2}\phi \Big)^{2}  = \frac{1}{3} c^2 - K \phi.
+\end{equation}
+It is clear that for any $\phi < \frac{3}{2} c$, the relation~\eqref{Whitham_Alt}
+can used in a bootstrap argument to show that any continuous solution must be
+in fact smooth. However for the case $\phi =  \frac{3}{2} c$ this bootstrap
+argument fails since the left-hand side vanishes. It can be concluded that
+a solutions containing a cusp will have a maximum  value of $\frac{3}{2} c$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+As an additional check, the discrete cosine coefficients of the solutions were
+examined, and fitted to the following models:
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align*}
+	\mathcal{E}(k) = \nu_1 e^{-\nu_2 k^n}, \qquad 
+	\mathcal{P}(k) = \frac{\mu_1}{\mu_2 + \mu_3 k^m},
+\end{align*}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%
+where $\nu_1$, $\nu_2$, $\mu_1$, $\mu_2$, $\mu_3$, $n$ and $m$ are fitting parameters. 
+A smooth function is known to have discrete cosine coefficients with exponential decay in $k$. 
+On the other hand, if a function is not smooth, the discrete cosine coefficients feature polynomial decay. 
+To identify the best fit, two parameters were used: $L^2$-norm of the residual 
+and the Akaike information criterion (AIC) measure.
+
+From the data given in the \autoref{t1}, one can deduce that for solutions with minimum speed 
+and maximum $L^2$-norm exponential fit is better than polynomial. 
+That is not the case for the terminal solution. 
+Thus, the first two solutions are smooth and the terminal solution is nonsmooth. 
+In fact, the polynomial fit is better than exponential for solutions that 
+are between the maximum $L^2$-norm solution and the terminal solution.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[t]
+%  \subfigure[][Minimum speed solution]{\includegraphics[width=0.31\textwidth]{1024_full_3p_1}}
+%~%
+%\subfigure[][Maximum $L^2$-norm solution]{\includegraphics[width=0.31\textwidth]{1024_full_3p_2}{}}
+%~%
+%\subfigure[][Terminal solution]{\includegraphics[width=0.31\textwidth]{1024_full_3p_3}{}}
+%\caption{\small Evolution of specific solutions in time. Time range for each solution is three periods.}
+%\label{time-evo}
+%\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+The numerical evidence brought forward supports the conclusion 
+that the Whitham bifurcation branch terminates at the terminal point
+indicated in Figure 3. Of course, as already mentioned,
+this conclusion has now also been reached using tools of mathematical analysis
+\cite{Wahlen2016}.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{table}	
+\centering
+\begin{tabular}{l|rr|rr|rr|}
+\cline{2-7}
+& \multicolumn{2}{ c     }{Min speed solution} & \multicolumn{2}{ c     }{Max $L^2$-norm solution}& \multicolumn{2}{ c |  }{Terminal solution} \\
+\multicolumn{1}{l|} {Model} &$\mathcal{E}(k)$&$\mathcal{P}(k)$&$\mathcal{E}(k)$&$\mathcal{P}(k)$&$\mathcal{E}(k)$&$\mathcal{P}(k)$ \\ 
+\midrule
+\multicolumn{1}{l|} {Residual's $L^2$ norm} & $5\times 10^{-5}$ &$4\times 10^{-3}$ & $7 \times 10^{-5}$ & $3\times 10^{-3}$ &  $6\times 10^{-3}$& $6\times 10^{-4}$ \\
+\multicolumn{1}{l|} {AIC} & -543 & -321 & -529 & -333 & -298 & -416 \\ 
+\bottomrule
+\end{tabular}
+\caption{\small Results for measures of fit.}
+\label{t1}
+\end{table}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Interaction of solitary wave solutions of modified Benjamin--Ono equation}
+\label{mB-O-interact}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this section, we utilize the \textsf{SpecTraVVave} package to obtain
+high-precision approximations to solitary-wave solutions of
+the modified Benjamin--Ono
+\begin{equation*}
+%\label{BO}
+u_t + u^2u_x + u_x - \mathcal{H} u_{xx} = 0,
+\end{equation*}
+which is a special case of the generalized Benjamin--Ono equation, with $p=2$.
+This case corresponds to the critical scaling, i.e., invariance of 
+the energy norm under the natural invariant scaling.
+
+The Benjamin--Ono equation was found by Benjamin~\cite{Benjamin1967} as a model
+for long small-amplitude interfacial waves in deep water. 
+The validity of approximating the more physically correct configuration of a continuous
+density distribution by the two-layer approximation has recently been justified mathematically
+\cite{Chen2016}.
+
+Solitary-wave solutions of the modified Benjamin--Ono equation with $p=3$, $p=4$ and $p=5$
+were approximated in~\cite{Bona2004} with a standard Newton scheme.
+The solutions in~\cite{Bona2004} were not very accurate, 
+but since singularity formation of the evolution
+equations were under study,
+the accuracy of the solitary-wave approximation was not an important issue.
+The problem with the method of~\cite{Bona2004} and some other works was that the fft
+used there was not purged of possible symmetries (translational and reflective).
+In the current code, since a cosine formulation is chosen, these symmetries
+are automatically eliminated, and the resulting computations are able to to
+render more accurate approximations.
+
+Solitary-wave solutions of these equation could be computed with higher accuracy
+using a type of Petviashvili method in~\cite{Pelinovsky2004}, but here we employ
+the \textsf{SpecTraVVave} package using the boundary equation~\eqref{solitary},
+and treating solitary waves as 
+traveling waves with sufficiently long wavelength 
+that have wave trough at zero. 
+Once these high-accuracy solutions are found, they are aligned in 
+an evolution code using a high-order time integrator, and the
+interaction of two waves is studied.
+
+Two questions are investigated. First, the interaction is investigated for evidence
+of integrability. Second, we are looking for possible annihilation of one of the waves,
+such as may happen in some other evolution equations~\cite{Tjon1979}.
+
+A possible approach to studying the question of complete integrability
+is analyzing the interaction of two solitary wave solutions of the equation, 
+such as carried out in~\cite{Kalisch2005,Kalisch2000} for other nonlocal equations.	
+In Figure 7, snapshots of interaction of two solitary waves at different times are shown.
+The time difference between two consecutive snapshots is constant.    
+As it may be observed, during the process of interaction, the two initial solitary waves combine 
+into a single wave, and an additional oscillation is produced. 
+This leads us to the conclusion that the interaction of solitary waves is not elastic 
+and the modified Benjamin--Ono equation may not be integrable. 
+In addition, it appears that the smaller wave disappears as most of its
+mass is acquired by the larger wave. Thus one may argue that the small
+wave is annihilated by the larger wave. It can also be observed that
+the larger wave starts growing, and it is likely that this growth
+will lead to finite-time blow-up. This question was not investigated further
+since blow-up phenomena have already been studied closely in~\cite{Kalisch2000}.
+Indeed, very recently, the finite-time blow-up 
+has been proven mathematically in~\cite{Martel2016}.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[t]
+%\centering
+%		\includegraphics[width=0.6\textwidth]{soliton_inter}      
+%\caption{\small Interaction of two solitary waves of the modified Benjamin--Ono equation.}
+%\label{solitons-full}
+%\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Effect of competing dispersion in the Benjamin equation}
+\label{Benjamin_eq}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The Benjamin equation was found by Benjamin~\cite{Benjamin1992} as a model for two-layer flow in the case
+when the interface is subject to surface tension. The approximation may not be a good model for
+a stratified situation, but more applicable to the case where two fluids are separated by a
+sharp interface. 
+The equation is 
+\begin{equation}\label{benjamin}
+u_t + u_x + uu_x - \mathcal{H} u_{xx} - \tau u_{xxx} = 0,
+\end{equation}
+where $\tau$ is a parameter similar to the Bond number in free surface flow~
+\cite{Benjamin1992}, \cite{Kalisch2007}, \cite{Walsh2014}.
+
+%%%%%%%%\vskip 0.05in
+
+In this section, a study relating to the effects of competing dispersion operators 
+on the shape of periodic traveling waves in the Benjamin equation is presented.
+An in-depth study of solitary waves was carried out in~\cite{Dougalis2015}.
+As will come to light, the periodic case features some new phenomena, such as secondary
+bifurcations, connecting and crossing branches.
+For the purpose of this study, we fix the parameter $\tau=0.1$,
+so that the dispersion relation for the linearized equation is
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+\begin{align}
+c(k) = 1 - |k| +  0.1  k^2. 	\label{Ben_disp_rel}
+\end{align}
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+Traveling wave solutions with full wavelengths $L_1 = \pi/5$, $L_2 = 4\pi /19$ and $L_3 = 4\pi$ 
+are computed for the equation~\eqref{benjamin}.  
+The corresponding wavenumbers are $k_1 = 10$, $k_2 = 19/2$, and $k_3 = 0.5$, respectively.
+A plot of the dispersion relation~\eqref{Ben_disp_rel} is given in Figure 8.
+Bifurcation branches of traveling wave solutions with the selected wavelengths are given in Figure 9.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[ht]
+%  \centering
+%\includegraphics[width=0.47\textwidth]{{disp_rel_plot}}  \caption{Dispersion relation \eqref{Ben_disp_rel}}
+%\label{disp-rel}
+%\end{figure}
+%
+%\begin{figure}
+%  \centering
+%\includegraphics[width=0.47\textwidth]{Benj_bifur_plot}
+%\caption{Bifurcation curves of the equation~\eqref{benjamin} with different wavelengths, $\blacktriangle$ --  points of bifurcation, $\bullet$ -- selected solutions (see \autoref{long-solutions}).}
+%\label{fig:ben-bifur}
+%\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+The branch denoted by $L_1$ originates at the bifurcation point 
+located at $c = 1$ and zero waveheight.
+The branches denoted by $L_2$ and $L_3$ originate from the same bifurcation point,
+located at $c = 0.525$ and zero waveheight. 
+These two branches continue in different directions, due to differences in wavelength.
+In particular, the $L_3$ branch contains waves with shorter wavelengths, and falls into the  
+capillary regime. On the other hand, the $L_2$ branch falls in the gravity regime.
+As the waveheight grows, solutions on the $L_3$ branch first cross the $L_1$ branch without
+connecting. Additional oscillations develop in the solutions until 
+a new fundamental wavelength $\pi/5$ is reached, and the branch terminates
+as it connects to the $L_1$ branch. The situation is depicted in Figure 10. 
+The point where the $L_1$ and $L_3$ branches meet
+is approximately $(c^*,a^*) = (0.945, 5.938)$.
+The corresponding profiles essentially overlap, as shown on Figure 11.
+This point can also be interpreted as a secondary bifurcation point 
+of the $L_1$ branch, where solutions with wavelengths that are multiples of $\pi/5$ develop. 
+We should note that similar phenomena concerning crossing and connecting
+branches were previously observed in \cite{Remonato2016}
+for the Whitham equation with surface tension which was introduced in~\cite{Hur2015}.
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[ht]
+%\begin{subfigure}
+%	\centering
+%	\includegraphics[width=0.47\textwidth]{left_long_solutions}      
+%	    \end{subfigure}%
+%~%
+%\begin{subfigure}
+%        \centering
+%		\includegraphics[width=0.47\textwidth]{right_long_solutions}{}      
+%    \end{subfigure}%
+%	\caption{\small Selected solutions of the equation~\eqref{benjamin}. 
+%Labels preserved as shown in \autoref{fig:ben-bifur}.}
+%\label{long-solutions}
+%\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[ht]
+%        \centering
+%		\includegraphics[width=0.47\textwidth]{last2}{}      
+%	\caption{\small Solution profiles at the point $(c^*, a^*)$ where the $L_1$ and $L_3$ branches meet (see \autoref{fig:ben-bifur}).}
+%\label{last-solutions}
+%\end{figure} 
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%                                                  %%%%%%%%%%%%%%%%%
+%%%%%%%%%%                                                  %%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Conclusions and future work}
+\label{sec:conclusion}
+
+The numerical algorithm of \textsf{SpecTraVVave} features ample flexibility for researching 
+different aspects of nonlocal dispersive wave equations and their traveling wave solutions.
+The solver package is simpler in use when compared with programs such as \textsf{AUTO} and 
+\textsf{Wavetrain}, however it does not have the same level of generality. 
+Moreover, \textsf{AUTO} and \textsf{Wavetrain} are programmed in low-level programming languages
+and will therefore run more efficiently.
+\textsf{SpecTraVVave} is implemented in an object-oriented fashion~\cite{Verdier2014}, 
+which makes the program easily expandable.
+\textsf{IPython} provides means for interactive work with the package, 
+and enables users to create convenient notebook-programs.
+A parametric approach in defining amplitude and phase speed makes it possible to follow
+turning points on bifurcation curves. Specification of different boundary conditions allows 
+computing solutions with certain features, such as traveling waves with zero mean, 
+or approximations to solitary waves.
+
+In this work, the \textsf{SpecTraVVave} package has been put to use for the study of 
+a number on nonlinear evolution equations: the Whitham equation, the modified Benjamin--Ono equation
+and the Benjamin equation.
+For the chosen set of parameters, experiments on the Whitham equation resulted in numerical 
+confirmation of the conjecture on cusped solutions. It was also possible to identify the point of stability 
+inversion of traveling wave solutions of the equation and the termination point of its bifurcation curve.
+
+In case of the modified Benjamin--Ono equation, the study on solitary wave solutions lead us to conclude that 
+interaction process ended with annihilation of one of the two waves. 
+The experiment on the Benjamin equation showed one more example of the effect of competing dispersion.
+As the amplitude increased, traveling wave solutions of wavelength $4\pi$ developed
+additional oscillations, and later connected up with a branch of solutions with
+wavelength $\pi/5$.
+
+Future work on the \textsf{SpecTraVVave} package will be focused on development of its functionality and broadening the range of problems that can be studied.
+Possible extensions may include implementation of algorithms based on the Petviashvili method~\citep{Alvarez2014, Alvarez2014j} and generalization to systems of equations.
+
+%\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\textbf{Acknowledgments.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+The authors would like to thank Mats Ehrnström and Erik Wahlén
+for fruitful discussions on the subject of the current paper. 
+This research was supported by the Research Council of Norway on grant no. 213474/F20
+and by the J C Kempe Memorial Fund on grant no. SMK-1238.
+
+%\appendix
+%\section*{Appendices}
+%\addcontentsline{toc}{section}{Appendices}
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Computing initial guesses from Stokes expansion.}
+%\label{sec:stokes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{equation}
+%u_t+\left[f(u)\right]_{x} + \mathcal{L}u_x=0, \label{gen}
+%\end{equation}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 The goal of this section is to explain how the idea of Stokes's approximation works in providing the initial data (guess) on wave and phase velocity for solving the equation~\eqref{gen} numerically.
+%
+%	We will consider $\Lop$ being linear and self-adjoint Fourier multiplier operator, and a function $f$ that has degree of zeros $p \geq 2$:
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align*}
+%\widehat{\Lop u}(k) &= \alpha(k) \cdot \widehat{u}(k),\\
+%\langle\Lop u, v \rangle_{L^2(0,L)} &= \langle u, \Lop v\rangle_{L^2(0,L)}, \\
+%p &= \min_{k \in \NN}{f^{(k-1)}(0) =0 \text{ and } f^{(k)}(0)\neq 0}
+%\end{align*}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 Consider the equation~\eqref{gen} and its solution in the form $u(x,t) = \phi(x-ct)$, which is a traveling wave solution. Inserting $\phi(x-ct)$ into~\eqref{gen} leads to the equation:
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{equation*}
+%-c\phi' + f'(\phi)\phi'+\mathcal{L}\phi'= 0,
+%\end{equation*}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 which can be integrated to give:
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{equation}
+%-c\phi + f(\phi)+\mathcal{L}\phi= B, \quad B = \mathrm{const}. \label{int_gen-B}
+%\end{equation}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	Consider $B = 0$ in equation~\eqref{int_gen-B}, and expansions of $\phi$ and $c$:
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align}
+%\phi = \xi = \varepsilon\xi_1 + \varepsilon^2\xi_2+ \cdots, \label{fi_exp}\\
+%c = c_0 + \varepsilon c_1 + \varepsilon^2 c_2+\cdots. \label{c_exp}
+%\end{align}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 The next step is to insert~\eqref{fi_exp} and~\eqref{c_exp} to the equation~\eqref{int_gen-B} and write out the terms at powers of $\varepsilon$. 
+%	The function $f(\phi)$ is expanded around zero and, therefore, will appear only in $\varepsilon^p$ terms. 
+%	Thus, the term at the first power of $\varepsilon$ reads: 
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align}
+%\varepsilon: \quad -c_0\xi_1 + \mathcal{L}\xi_1 = 0, \label{epsilon1}
+%\end{align}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 Hence, $c_0$ is an eigenvalue of the operator $\Lop$, regarded as defined on $L$-periodic functions. 
+%%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 Taking the Fourier transform of the equation~\eqref{epsilon1} gives:
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{equation}
+%-c_0\widehat{\xi_1}(k) + \alpha(k)\widehat{\xi_1}(k) = 0 . \label{eps1-transformed}
+%\end{equation}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%\quad\rightarrow\quad \widehat{\xi_1}(k) \equiv 0 \mbox{ or } (-c_0+\widehat{\mathcal{L}}(k)) = 0. 
+%	 The equation~\eqref{eps1-transformed} has two trivial solutions: either $\xi_1(k) \equiv 0$ or $\alpha(k) \equiv c_0$.
+If we assume non-trivial $\xi_1$ and $\alpha(k) \neq \text{const}$, the following solves the problem:
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align}
+%  \widehat{\xi_1}(k) = 2\pi\delta(k - k_0), \quad \text{and} \quad c_0 = \alpha(k_0), \label{solution-xi}
+%\end{align}
+%	 for some $k_0 \in \R$. 
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	Since $\xi_1$ is the first-order approximation to $\phi$, the corresponding wave number should be equal to 1.
+The $L$-periodicity condition entails that $k_0 = 2\pi/L\cdot 1$.
+The spatial variable $x$ has to be scaled to $x'= L/2\pi x$, accordingly.
+From the solutions in~\eqref{solution-xi} we have
+%\begin{equation}
+%	\xi_1(x') = e^{ik_0x'} = \cos(k_0x') + \sin(k_0x').
+%\end{equation}	
+%Considering the projection onto the space $\SN$, we are led to choose $\xi_1(x') = \cos(k_0x')$.
+%%
+%%
+%%
+%
+%%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 For further analysis, let us define an operator $\Aop$ 
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{equation*}
+%	\Aop :=  - c_0 \mathcal{E} + \Lop,
+%\end{equation*}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%where $\mathcal{E}$ is the identity operator. The operator $\Aop$ inherits the property of being 
+%self-adjoint from $\Lop$. Moreover, it follows from~\eqref{epsilon1} that $\Aop \xi_1 = 0$ and $\xi_1 \in \ker(\Aop)$. 
+%If $p>2$ then $f''(0)=0$ and the terms at $\epsilon^2$ are:
+%\begin{equation}
+%\Aop \xi_2 - c_1\xi_1 = 0.
+%\end{equation}
+%%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 Taking scalar multiplication of the latter with $\xi_1$, one obtains:
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align*}
+%	&\langle\xi_1, \Aop\xi_2 \rangle_{L^2(0,L)} = c_1 \| \xi_1\|^2_{L^2(0,L)}, \\
+%	&\langle\xi_1, \Aop\xi_2\rangle_{L^2(0,L)} = \langle\Aop \xi_1, \xi_2\rangle_{L^2(0,L)} = \langle0, \xi_2\rangle_{L^2(0,L)} = 0.
+%\end{align*}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 As a result, one has $c_1 \| \xi_1\|^2_{L^2(0,L)} = 0$ and, hence, $c_1 = 0$.
+%	Repeating the same argument, it becomes clear that $c_k = 0$ for any $k \leq p-1$.
+%	Besides that, $\xi_2$ is in the kernel of $\Aop$, so it may be assumed to be proportional to $\xi_1$.
+%	The terms at order $\varepsilon^p$ are:
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align}
+%  \Aop \xi_p - c_{p-1}\xi_1 + \frac{f^{(p)}(0)}{p!} \xi_1^p = 0. \label{epsp}
+%\end{align}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 Let us denote for brevity
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align*}
+%	f_p := \frac{f^{(p)}(0)}{p!}.
+%\end{align*}
+%%%%%%%%%%			%%%%%%%%%%%%%%%
+%	 Pairing~\eqref{epsp} with $\xi_1$ (and assuming $\| \xi_1 \|_{L^2(0,L)} = 1$) gives
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align}
+%	c_{p-1} = f_p	\cdot \langle\xi_1^p, \xi_1\rangle_{L^2(0,L)}, \label{cp-1}
+%\end{align}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 which gives us the value of $c_{p-1}$.
+%	It only remains to solve the following problem numerically in order to obtain $\xi_p$:
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align}
+%  \Aop \xi_p = c_{p-1}\xi_1 - f_p\xi_1^p \label{epsilonp}
+%\end{align}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%
+%	For the last equation to be solved, the operator $\Aop$ has to be invertible. It is also required that $\langle\xi_1, \xi_p\rangle_{L^2(0,L)}=0$. 
+%	Therefore the solution is sought in the space orthogonal to $\ker(\Aop)$.
+%	Since $\Aop$ is still a Fourier multiplier operator, one can take the Fourier transform of the equation~\eqref{epsilonp} to find
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align*}
+%	\widehat\Aop(k) \widehat\xi_p(k)& = c_{p-1}\widehat\xi_1(k) - f_p				
+%	\widehat{\xi_1^p}(k), \\
+%	\widehat\xi_p(k)& = \widehat\Aop(k)^{-1} \left(c_{p-1}\widehat\xi_1(k) - f_p	
+%	\widehat{\xi_1^p}(k)\right).
+%\end{align*}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%	 
+%Taking the inverse Fourier transform of $\widehat\xi_p(k)$ gives $\xi_p$.
+%Since only even solutions of the problem are considered the cosine part of the Fourier transforms will be required.
+%It is sufficient to use $\xi_1$ and $c_0$ as the initial guesses for the Newton method.
+However, it should be noted that for different values of $p$ the pair of parameters $\xi_p$ and $c_{p-1}$ are computed in different ways.
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{enumerate}[label=\alph*)]
+%	\item If $p = 2$, then $\xi_2$ is computed from~\eqref{epsilonp}, but $c_{p-1}$ here becomes zero. Therefore one has to consider the next level of the expansion $\varepsilon^p$.
+%	\item For odd values of $p$ the parameter $c_{p-1}$ can be computed from~\eqref{cp-1} and $\xi_p$ from~\eqref{epsilonp}.
+%	\item For even values of $p \geq 4$ the parameter $\xi_p$ can be computed, but $c_{p-1}$ may not be non-zero in general. 
+%	In such cases a different strategy of fixing the initial guess should be used.
+%\end{enumerate}
+%
+%
+%
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Presentation of \textsf{SpecTraVVave} and its workflow}
+%\label{sec:implementation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 
+%\subsection{Overview}
+%There are several classes in the \textsf{SpecTraVVave} package. An overview of the program is shown 
+%in \autoref{schematic}. The workflow begins with defining a flux function $f$ and the Fourier multiplier function $\alpha$ to set up an equation. 
+%The traveling wave solution is characterized by the wavelength $L$ 
+%and a boundary condition $\Omega(c, a, \phi_N, B)$. These parameters are fixed for a given problem.
+%The defined equation is then discretized. 
+%The \pyth{Discretization} object contains all required elements such as grid points, wave-numbers and the  discrete linear operator.
+%		
+%The initial guess and the equation's residual are passed from the \pyth{Discretization} to the \pyth{Solver} object. 
+%The \pyth{Navigation} object is responsible for finding good initial guesses for $c$ and $a$ that are  
+%passed to the \pyth{Solver} object. 
+%The \pyth{Solver} object applies Newton's method to find a solution to the system of equations~\eqref{system-N3}. 
+%	
+%The new solution is sent back to the  \pyth{Discretization} and \pyth{Navigation} objects, where variables get updated.
+%All computed solutions are stored for further analysis. 
+%This finishes one iteration. 
+%For the next iteration the updated variables are used and a new solution is found. 
+%The process may be continued as long as the Jacobian of the problem is non-singular. 
+% 
+% %%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{figure}[ht]
+%	\centering
+%\includegraphics[width=0.8\textwidth]{block-scheme} 
+%\caption{\small Overview of the \textsf{SpecTraVVave} package.}     
+%\label{schematic}
+%\end{figure} 
+% %%%%%%%%%===========%%%%%%%%%%%%%%%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsection{Class Description}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 
+% 
+%We present an overview of the classes used in \textsf{SpecTraVVave} package.
+Note that, since the package is under continuous modification and development, we describe here only the  basic classes and functions the package.
+%We refer to the package repository \cite{github} for up-do-date tutorials and installation instructions.
+% 
+% The \pyth{Equation} class is the general class for all model equations. 
+%Its only role is to store a parameter $L$, the wavelength:
+%
+%\begin{python}
+%class Equation(object):
+%    def __init__(self, L):
+%        self.length = L
+%\end{python}
+%A subclass of the \pyth{Equation} class has to implement two functions, \pyth{compute_kernel} and \pyth{flux}. 
+%
+%The KdV model equation
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%\begin{align}
+%	u_t + \frac{3}{2} u u_x + u_x + \frac{1}{6} u_{xxx} = 0
+%\end{align}
+%%%%%%%%%%===========%%%%%%%%%%%%%%%
+%with $f(u) = \frac{3}{4} u^2$ and $\widehat{\Lop u}(k) = (1 - \frac{1}{6} k^2)\widehat{u}(k)$ is presented in the program as a subclass of the \pyth{Equation} class:
+%\begin{python}
+%class KDV(Equation):
+%    def compute_kernel(self, k):
+%        return 1-1/6*k*k
+%
+%    def flux(self, u):
+%        return 3/4*u*u  
+%\end{python} 
+%
+% 
+% 
+% 
+%On can then create an object of the class \pyth{KDV} with the command:
+%\begin{python}
+%kdv = KDV(L=np.pi)
+%\end{python} 
+%%
+%The solver will compute only a half of a solutions profile. The full wavelength of the solutions 
+%of the defined equation will be equal to $2\pi$. 
+%
+%In order to find solutions with specific features, boundary conditions are introduced as separate classes.
+%For instance, the boundary condition specifying a constant of integration is implemented as follows:
+%\begin{python}
+%class Const(object):
+%""" The boundary condition under which the constant of integration 
+%(B) is always set to zero. """
+%    def __init__(self, level=0):
+%        self.level = level
+%
+%    def enforce(self, wave, variables, parameters):
+%    """ Enforces the Const boundary condition. """
+%        return np.hstack([variables[0] - self.level])
+%
+%    def variables_num(self):
+%    """ The number of additional variables that are required to construct 
+%    the Const boundary condition. """
+%        return 1
+%\end{python}
+%A \pyth{Const} boundary condition object is created as follows:
+%\begin{python}
+%boundary_condition = Const()
+%\end{python}
+%
+%The next step is to create an object of \pyth{Discretization} class, which is initialized with a model equation such as \pyth{kdv_model} and the number of grid points.
+The main parts of the class are the following:
+%\begin{python}
+%class Discretization(object):
+%    def __init__(self, model_equation, grid_size):
+%        self.equation = model_equation
+%        self.size = grid_size
+%
+%    def operator(self, u):
+%        u_ = scipy.fftpack.dct(u, norm='ortho')
+%        Lv = self.fourier_multiplier()*u_
+%        result = scipy.fftpack.idct(Lv, norm='ortho')
+%        return result
+%
+%    def residual(self, u, wavespeed, const_B):
+%        residual = - wavespeed*u + self.equation.flux(u) 
+%                                        + self.operator(u) - const_B
+%        return  residual
+%\end{python}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%The call \pyth{Discretization.operator(u)} computes $\Lop u $ as the inverse transform of a transformed convolution 
+%\begin{equation}
+%\Lop u = \mathcal{F}^{-1}[ \widehat{\Lop u}(k)] = \mathcal{F}^{-1}\left[ \alpha(k)\cdot\widehat{u}(k)\right].
+%\end{equation}
+%The result of the call \pyth{Discretization.residual(u, wavespeed, const_B)} is then used in the \pyth{Solver} class. 
+%An object of the \pyth{Solver} class is initialized with an object of the \pyth{Discretization} class, and a boundary condition object.
+%
+%\begin{python}
+%class Solver(object):
+%    def __init__(self, discrete_problem, boundary_condition):
+%        self.discretization = discrete_problem
+%        self.boundary = boundary_condition
+%
+%    def solve(self, guess_wave, parameter_anchor, direction):
+%    """ Runs a Newton solver on a system of nonlinear equations once. 
+%    Takes the residual(vector) as the system to solve. parameter_anchor 
+%    is the initial guess for (c,a) values and it is taken from the 
+%    Navigation class. """
+%        size = len(guess_wave)
+%        self.discretization.size = size
+%
+%        def residual(vector):
+%        """ Contructs a system of nonlinear equations. First part, 
+%        main_residual, is from given wave equation; second part, 
+%        boundary_residual, comes from the chosen boundary conditions. """
+%        . . .   
+%        return np.hstack([main_residual, boundary_residual, 
+%                                          amplitude_residual])  
+%    . . .
+%    return new_wave, new_boundary_variables, new_parameter
+%\end{python}
+%%
+%Some omitted parts in the above script are substituted by \textsf{'.
+.
+.'} sign.
+Each iteration on a \pyth{Solver} object is run from a \pyth{Navigation} object, which takes the \pyth{Solver} object for initialization.
+%\begin{python}
+%class Navigation(object):
+%""" Runs the Solver and stores the results. """
+%
+%    def __init__(self, solver_object, size=32):
+%    self.solve = solver_object.solve # function to run Newton method
+%    self.size = size # size for navigation
+%    . . . 
+%    def run_solver(self, current_wave, pstar, direction):
+%        new_wave, variables, p3 = self.solve(current_wave, pstar, direction)
+%        return new_wave, variables, p3
+%\end{python}
+%
+%
+%All the above classes can be modified and developed further, new classes may be defined as well.
+%
+%\subsection{Detailed Workflow}
+%
+%The workflow with the package consists of three basic steps:
+%\begin{enumerate}
+%\item Once the necessary classes have been imported in the current namespace, generate all necessary objects:
+%\begin{python}
+%equation = KDV(L=np.pi)
+%boundary_condition = Const()
+%discretization = Discretization(equation, grid_size=64)
+%solver = Solver(discretization, boundary_condition)
+%navigator = Navigation(solver)    
+%\end{python}
+%\item Choose a number of iterations, i.e., the number of solutions to compute, and run the solver:
+%\begin{python}
+%n_iter = 50
+%navigator.run(n_iter)
+%\end{python}
+%\item All computed solutions are stored in \pyth{navigation_object}
+%\begin{python}
+%last_computed = -1    
+%wave_profile = navigator[last_computed]['solution']
+%wave_speed = navigator[last_computed]['current'][0]
+%wave_amplitude = navigator[last_computed]['current'][1]
+%\end{python}
+%\end{enumerate} 
+%
+%We further refer to the code repository \url{https://github.com/olivierverdier/SpecTraVVave} for up-to-date instructions on how to run the code.
+%
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+% For two-column wide figures use
+%
+% For tables use
+
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+%%%%%%%%%===========%%%%%%%%%%%%%%%
+
+% BibTeX users please use one of
+%\bibliographystyle{spbasic}      % basic style, author-year citations
+%\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
+%\bibliography{}   % name your BibTeX data base
+
+% Non-BibTeX users please use
+%\begin{thebibliography}{}
+%%
+%% and use \bibitem to create references. Consult the Instructions
+%% for authors for reference list style.
+%%
+%
+%\bibitem{ABFS} Abdelouhab, L. Bona, J.L., Felland, M. and Saut, J.-C. 
+%     {\em Nonlocal models for nonlinear dispersive waves}.
+%     Physica D \textbf{40},360-392 (1989) 
+%
+%
+%\bibitem{APR} Albert, J.P., Bona, J.L. and Restrepo, J.M. 
+%{\em Solitary-wave solutions of the Benjamin equation}.
+%SIAM J. Appl. Math. \textbf{59},2139--2161 (1999) 
+%
+%
+%\bibitem{B1} Benjamin, T.B.
+%          {\em Internal Waves of permanent form in fluids of great depth}.
+%          J. Fluid Mech.  \textbf{29}, 559--592 (1967)
+%
+%
+%\bibitem{B2} Benjamin, T.B.
+%               {\em The stability of solitary waves}.
+%              Proc. Roy. Soc. London A 328, 153-183 (1972) 
+%
+%\bibitem{B3} Benjamin, T.B. 
+%          {\emph{A new kind of solitary wave}}.
+%          J. Fluid Mech. \textbf{245}, 401--411  (1992)
+%
+%\bibitem{Duran2011JCAM} \'{A}lvarez, J. and  Dur\'{a}n, A.
+%{\em A numerical scheme for periodic travelling-wave simulations in some nonlinear dispersive wave models}.
+%J. Comput. Appl. Math. \textbf{235}, 1790--1797 (2011)
+%
+%\bibitem{Duran2014CNSNS} \'{A}lvarez, J. and  Dur\'{a}n, A.
+%{\em An extended Petviashvili method for the numerical generation of traveling and localized waves}.
+%Commun. Nonlinear Sci. Numer. Simul. \textbf{19}, 2272--2283 (2014)
+%
+%\bibitem{Duran2014JCAM} \'{A}lvarez, J. and  Dur\'{a}n, A.
+%{\em, Petviashvili type methods for traveling wave computations: I. Analysis of convergence}.
+%J. Comput. Appl. Math. \textbf{266}, 39--51 (2014)
+%
+%\bibitem{Duran2014JCAMCor} \'{A}lvarez, J. and  Dur\'{a}n, A.
+%{\em Corrigendum to "Petviashvili type methods for traveling wave computations: I. Analysis of convergence},
+%[J. Comput. Appl. Math. 266 (2014) 39--51]. J. Comput. Appl. Math. \textbf{277}, 215--216 (2015)
+%
+%
+%\bibitem{BDKM} Bona, J.L., Dougalis, V.A.,  Karakashian, O.A. and McKinney, W.R.
+%   {\em Conservative, high-order numerical schemes for the generalized
+%   Korteweg--de Vries equation},
+%   Philos. Trans. Royal Soc. London Ser. A, \textbf{351}, 107--164 (1995)
+%
+%
+%\bibitem{BoKa} Bona, J.L. and Kalisch, H.
+%              {\em Singularity formation in the generalized Benjamin--Ono equation}.
+%              Discrete Contin. Dyn. Syst. \textbf{11}, 779--785 (2004)
+%
+%
+%\bibitem{BSS}
+%Bona, J.L., Souganidis, P.E. and Strauss, W.A.
+%{\em Stability and instability of solitary waves of Korteweg--de Vries type}.
+%Proc. R. Soc. Lond. A \textbf{411}, 395-412 (1987) 
+%%
+%%
+%\bibitem{HKN} Borluk, H., Kalisch, H. and Nicholls, D.P. 
+%{\em A numerical study of the Whitham equation as a model for steady surface water waves}.
+%J. Comput. Appl. Math. \textbf{296}, 293--302 (2016) 
+%
+%
+%\bibitem{TJB}
+%Bridges, T.J. 
+%{\em Superharmonic instability, homoclinic torus bifurcation and water-wave breaking}.
+%J. Fluid Mech. \textbf{505}, 153-162 (2004)
+%
+%
+%\bibitem{Chen} Chen, H.
+%{\em Existence of periodic travelling-wave solutions of nonlinear, dispersive wave equations}. 
+%Nonlinearity \textbf{17}, 2041--2056 (2004)
+%
+%
+%\bibitem{CB} Chen H. and Bona, J.L. 
+%          {\em Existence and asymptotic properties of solitary-wave
+%          solutions of Benjamin-type equations}.
+%          Adv. Diff. Eq. \textbf{3},  51-84 (1998)
+%
+%%
+%
+%\bibitem{ChenWalsh2016} Chen, R.M. and Walsh, S. 
+%{\em Continuous dependence on the density for stratified steady water waves}.
+%Arch. Ration. Mech. Anal. \textbf{219}, 741--792 (2016)
+%
+%
+%\bibitem{CC2} Choi, W. and Camassa, R.
+%          {\em Fully nonlinear internal waves in a two-fluid system}.
+%          J. Fluid Mech.  \textbf{396}, 1-36 (1999)
+%
+%\bibitem{dFSS}
+%de Frutos, J. and Sanz-Serna, J. M.
+%{\em An easily implementable fourth-order method for the time integration of wave problems}.
+%J. Comp. Phys. \textbf{103}, 160--168 (1992)
+%
+%
+%
+%\bibitem{Doed97}
+%Doedel, E.J., Champneys, A.R., Fairgrieve, T.F., Kuznetsov, Y.A., Sandstede, B. and Wang, X.-J., 
+%{\em AUTO97 : Continuation and bifurcation software for ordinary differential equations}, 
+%available at 
+%\url{http://indy.cs.concordia.ca/auto/}, Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada, 1997. 
+%
+%
+%\bibitem{DDM} Dougalis, V.A., Duran, A. and Mitsotakis, D. 
+%{\em Numerical solution of the Benjamin equation}.
+%Wave Motion \textbf{52}, 194--215 (2015)
+%
+%
+%
+%\bibitem{EK1}
+%Ehrnstr{\"o}m, M. and Kalisch, H. 
+%{\em Traveling waves for the Whitham equation}. 
+%Differential Integral Equations \textbf{22}, 1193--1210 (2009)
+%
+%\bibitem{EK2}
+%Ehrnstr{\"o}m, M. and Kalisch, H. 
+%{\em Global bifurcation for the Whitham equation}.
+%Math. Mod. Nat. Phenomena \textbf{8}, 13--30 (2013)
+%%
+%\bibitem{EW2016}
+%Ehrnstr{\"o}m, M. and  Wahl\'en, E. 
+%{\em On Whitham's conjecture of a highest cusped wave for a nonlocal dispersive equation}. arXiv:1602.05384
+%(2016).
+%
+%\bibitem{PyBook}
+%F\"uhrer, C., Solem, J.E., Verdier, O. 
+%Computing with Python: An introduction to Python for science and engineering, Pearson (2014)
+%%
+%\bibitem{GK} Gruji\'{c}, Z. and Kalisch, H.
+%{\em Gevrey regularity for a class of water-wave models}.
+%Nonlinear Analysis \textbf{71}, 1160--1170 (2009)
+%
+%\bibitem{Hur2015} Hur, V.M. 
+%{\em Breaking in the Whitham equation for shallow water waves}.
+%arXiv:1506.04075 (2015).
+%
+%
+%\bibitem{HurJohnson}
+%Hur, V.M. and Johnson, M. 
+%{\em Modulational instability in the Whitham equation of water waves}.
+%Studies in Applied Mathematics \textbf{134}, 120--143 (2015)
+%
+%
+%\bibitem{HurTao} Hur, V.M. and Tao, L.
+%{\em Wave breaking for the Whitham equation with fractional dispersion}.
+%arXiv:1410.1570, 2014.
+%
+%
+%\bibitem{FW}
+%Fornberg, B. and Whitham, G. B. 
+%{\em A numerical and theoretical study of certain nonlinear wave phenomena}. 
+%Phil. Trans. Roy. Soc. A \textbf{289}, 373--404 (1978)
+%
+%
+%%\bibitem{ipython}
+%%\url{https://ipython.org/}
+%
+%\bibitem{HK}
+%Kalisch, H. 
+%{\em Error analysis of a spectral projection of the regularized Benjamin--Ono equation}.
+%BIT Numerical Mathematics \textbf{45}, 69--89 (2005)
+%
+%
+%\bibitem{KalischBenjamin}
+%Kalisch, H.
+%{\em Derivation and comparison of model equations for interfacial capillary-gravity waves in deep water}.
+%Math. Comput. Simulation \textbf{74},168--178 (2007)
+%%(2-3):
+%
+%\bibitem{HKJB}
+%Kalisch, H. and Bona, J.L.
+%{\em Models for internal waves in deep water}. 
+%Discrete and Continuous Dynamical Systems \textbf{6}, 1--20 (2000)
+%
+%\bibitem{Kato} Kato, T.
+%{\em On the Cauchy problem for the (generalized) Korteweg--de Vries equation}.
+%Studies in applied mathematics, 93--128, Adv. Math. Suppl. Stud., \textbf{8} 	
+%(Academic Press, New York, 1983).
+%
+% 
+%\bibitem{Keller} 
+%Keller H.B., 
+%{\em Numerical solution of bifurcation and nonlinear eigenvalue problems}. 
+%Applications of Bifurcation Theory, Rabinowitz, P. H., Academic Press, (1977), pp. 359--384. 
+%
+%
+%\bibitem{LS} Lannes, D. and Saut, J.-C.
+%{\em Remarks on the full dispersion Kadomtsev--Petviashvli equation}.
+%Kinet. Relat. Models \textbf{6}, 989--1009 (2013)
+%
+%
+%
+%\bibitem{MM} Martel, Y. and Merle, F.
+%{\em Blow up in finite time and dynamics of blow up solutions for the L2-critical generalized KdV equation}.
+%J. Amer. Math. Soc. \textbf{15}, 617--664 (2002)
+%
+%
+%\bibitem{MP} Martel, Y. and Pilod, D.
+%{\em Construction of a minimal mass blow up solution of the modified Benjamin-Ono equation}.
+%arXiv:1605.01837 (2016).
+%
+%
+%\bibitem{MKD2015} Moldabayev, D., Kalisch, H. and  Dutykh, D. 
+%{\em The Whitham Equation as a model for surface water waves}.
+%Phys. D  \textbf{309}, 99--107 (2015)
+%
+%
+%\bibitem{github}
+%Moldabayev, D., Verdier, O. and Kalisch, H.
+%{\em SpecTraVVave}, available at \newline
+%\url{https://github.com/olivierverdier/SpecTraVVave}. 
+%
+%
+%\bibitem{NgK}
+%Nguyen, N.T. and Kalisch, H.
+%{\em Orbital stability of negative solitary waves}.
+%Mathematics and Computers in Simulation \textbf{80}, 139-150 (2009) 
+%
+%
+%\bibitem{O} Ono, H.
+%         {\emph{Algebraic Solitary Waves in Stratified Fluids}}.
+%         J. Phys. Soc. of Japan \textbf{39}, 1082-1091 (1975)
+%
+%
+%\bibitem{Pelinovsky} Pelinovsky, D.E. and Stepanyants, Y.A. 
+%{\em Convergence of Petviashvili's iteration method for numerical approximation 
+%of stationary solutions of nonlinear wave equations}.
+%SIAM J. Numer. Anal. \textbf{42}, 1110--1127 (2004)
+%
+%\bibitem{IPython} Fernando Pérez, Brian E. Granger
+% \emph{IPython: A System for Interactive Scientific Computing}, 
+%Computing in Science and Engineering, vol. 9, no. 3, pp. 21-29, May/June 2007.
+%% doi:10.1109/MCSE.2007.53.
+%URL: \url{http://ipython.org}
+%
+%\bibitem{Petviashvili} Petviashvili, V.I.
+%{\em Equation of an extraordinary soliton}, 
+%Sov. J. Plasma Phys. \textbf{2}, 257--258 (1976) 
+%
+%
+%\bibitem{RK}
+%Remonato, F. and Kalisch, H.
+%{\em Numerical bifurcation for the capillary Whitham equation}.
+%arXiv:1604.08324 (2016).
+%
+%
+%\bibitem{Sanford}
+%Sanford, N., Kodama, K., Carter, J.D. and Kalisch, H. 
+%{\em Stability of traveling wave solutions to the Whitham equation}. 
+%Phys. Lett. A \textbf{378}, 2100--2107 (2014)
+%
+%
+%\bibitem{Sherratt} Sherratt, J.A. 
+%{\em Numerical continuation methods for studying periodic travelling wave (wavetrain) 
+%solutions of partial differential equations}.
+%Appl. Math. Comp. \textbf{218}, 4684--4694 (2012)
+%
+%
+%\bibitem{Tjon} Courtenay Lewis, J. and Tjon, J.A.
+%              {\em Resonant production of solitons in the RLW equation}.
+%              Phys. Lett. A \textbf{73}, 275--279 (1979)
+%
+%
+%
+%\bibitem{Walsh2014} Walsh, S. 
+%{\em Steady stratified periodic gravity waves with surface tension I: Local bifurcation}.
+%Discrete Contin. Dyn. Syst. \textbf{34}, 3241--3285 (2014)
+%
+%
+%
+%\bibitem{Walsh2014j} Walsh, S. 
+%{\em Steady stratified periodic gravity waves with surface tension II: global bifurcation}. 
+%Discrete Contin. Dyn. Syst. \textbf{34}, 3287--3315 (2014)
+%
+%
+%
+%\bibitem{Whitham1967j} 
+%Whitham, G.B. 
+%{\em Variational methods and applications to water waves}.
+%Proc. Roy. Soc. London A \textbf{299}, 6--25 (1967)
+%
+%
+%\bibitem{Whitham74}
+%Whitham, G.B. 
+%Linear and Nonlinear Waves. Wiley, New York, (1974)
+%
+%
+%
+%\end{thebibliography}
+%
+%\end{document}

### 2017-02-08 Thesis/latex/pythonlisting.tex
paperC - next step

% Python listing setup

\usepackage{color}
\usepackage[procnames]{listings}
\usepackage{textcomp}
\usepackage{setspace}
%\usepackage{palatino}
\usepackage{verbatim}
\renewcommand{\lstlistlistingname}{Code Listings}
\renewcommand{\lstlistingname}{Code Listing}
\definecolor{gray}{gray}{0.5}
\definecolor{green}{rgb}{0,0.5,0}
\definecolor{lightgreen}{rgb}{0,0.7,0}
\definecolor{purple}{rgb}{0.5,0,0.5}
\definecolor{darkred}{rgb}{0.5,0,0}
\lstnewenvironment{python}[1][]{
\lstset{
language=python,
basicstyle=\ttfamily\footnotesize\setstretch{1},
stringstyle=\color{green},
showstringspaces=false,
alsoletter={1234567890},
otherkeywords={\ , \}, \{},
keywordstyle=\color{blue},
emph={access,and,as,break,class,continue,def,del,elif,else,%
except,exec,finally,for,from,global,if,import,in,is,%
lambda,not,or,pass,print,raise,return,try,while,assert},
emphstyle=\color{green}\bfseries,
emph={[2]self},
emphstyle=[2]\color{gray},
emph={[4]ArithmeticError,AssertionError,AttributeError,BaseException,%
DeprecationWarning,EOFError,Ellipsis,EnvironmentError,Exception,%
False,FloatingPointError,FutureWarning,GeneratorExit,IOError,%
ImportError,ImportWarning,IndentationError,IndexError,KeyError,%
KeyboardInterrupt,LookupError,MemoryError,NameError,None,%
NotImplemented,NotImplementedError,OSError,OverflowError,%
PendingDeprecationWarning,ReferenceError,RuntimeError,RuntimeWarning,%
StandardError,StopIteration,SyntaxError,SyntaxWarning,SystemError,%
SystemExit,TabError,True,TypeError,UnboundLocalError,UnicodeDecodeError,%
UnicodeEncodeError,UnicodeError,UnicodeTranslateError,UnicodeWarning,%
UserWarning,ValueError,Warning,ZeroDivisionError,abs,all,any,apply,%
basestring,bool,buffer,callable,chr,classmethod,cmp,coerce,compile,%
complex,copyright,credits,delattr,dict,dir,divmod,enumerate,eval,%
execfile,exit,file,filter,float,frozenset,getattr,globals,hasattr,%
hash,help,hex,id,input,int,intern,isinstance,issubclass,iter,len,%
license,list,locals,long,map,max,min,object,oct,open,ord,pow,property,%
quit,range,raw_input,reduce,reload,repr,reversed,round,set,setattr,%
slice,sorted,staticmethod,str,sum,super,tuple,type,unichr,unicode,%
vars,xrange,zip},
emphstyle=[4]\color{purple}\bfseries,
upquote=true,
morecomment=[s][\color{purple}]{"""}{"""},
commentstyle=\color{red}\slshape,
literate={>>>}{\textbf{\textcolor{darkred}{>{>}>}}}3%
         {...}{{\textcolor{gray}{...}}}3,
procnamekeys={def,class},
procnamestyle=\color{blue}\textbf,
framexleftmargin=1mm, framextopmargin=1mm, 
rulesepcolor=\color{blue},#1
}}{}

### 2017-02-08 Thesis/latex/paperC.tex
paperC compiled

@@
 
-In nonlinear dispersive evolution equations, the competing effects of nonlinearity and dispersion 
-make a number of interesting phenomena possible. In the current work, the focus is on the numerical
-approximation of traveling-wave solutions of such equations.
-We describe our efforts to write a dedicated \textsf{Python}
-code which is able to compute traveling-wave solutions of nonlinear
-dispersive equations in a very general form.
-%\begin{equation*}
-%u_t + [f(u)]_{x} + \Lop u_x = 0,
-%\end{equation*}
-%%\vskip -0.05cm
-%where $\Lop$ is a self-adjoint operator, and $f$ is a real-valued function with 
-%$f(0) = 0$ and $f'(0) = 0$.
-
-The \textsf{SpecTraVVave} code uses a continuation method coupled with a spectral projection
-to compute approximations of steady symmetric solutions of this equation.
-The code is used in a number of situations to gain an
-understanding of traveling-wave solutions. The first case is the Whitham
-equation, where numerical evidence points to the conclusion that the main 
-bifurcation branch features three distinct points of interest, namely a turning point, 
-a point of stability inversion, and a terminal point which corresponds to a cusped wave.
-
-The second case is the so-called modified Benjamin--Ono equation where
-the interaction of two solitary waves is investigated. It is found
-that two solitary waves may interact in such a way
-that the smaller wave is annihilated. The third case concerns
-the Benjamin equation which features two competing dispersive operators.
-In this case, it is found that bifurcation curves of periodic traveling-wave
-solutions may cross and connect high up on the branch in the nonlinear regime.
-%Include keywords and mathematical subject classification numbers as needed.
+%In nonlinear dispersive evolution equations, the competing effects of nonlinearity and dispersion 
+%make a number of interesting phenomena possible. In the current work, the focus is on the numerical
+%approximation of traveling-wave solutions of such equations.
+%We describe our efforts to write a dedicated \textsf{Python}
+%code which is able to compute traveling-wave solutions of nonlinear
+%dispersive equations in a very general form.
+%%\begin{equation*}
+%%u_t + [f(u)]_{x} + \Lop u_x = 0,
+%%\end{equation*}
+%%%\vskip -0.05cm
+%%where $\Lop$ is a self-adjoint operator, and $f$ is a real-valued function with 
+%%$f(0) = 0$ and $f'(0) = 0$.
+%
+%The \textsf{SpecTraVVave} code uses a continuation method coupled with a spectral projection
+%to compute approximations of steady symmetric solutions of this equation.
+%The code is used in a number of situations to gain an
+%understanding of traveling-wave solutions. The first case is the Whitham
+%equation, where numerical evidence points to the conclusion that the main 
+%bifurcation branch features three distinct points of interest, namely a turning point, 
+%a point of stability inversion, and a terminal point which corresponds to a cusped wave.
+%
+%The second case is the so-called modified Benjamin--Ono equation where
+%the interaction of two solitary waves is investigated. It is found
+%that two solitary waves may interact in such a way
+%that the smaller wave is annihilated. The third case concerns
+%the Benjamin equation which features two competing dispersive operators.
+%In this case, it is found that bifurcation curves of periodic traveling-wave
+%solutions may cross and connect high up on the branch in the nonlinear regime.
+%%Include keywords and mathematical subject classification numbers as needed.
 
@@
 The content of the paper is structured as follows. A mathematical description of the numerical method of 
-\textsf{SpecTraVVave} is given in \autoref{sec:spectral}.
-\autoref{sec:experiments} presents results of different experiments carried out with the package.
-Concluding remarks are given in \autoref{sec:conclusion}.
-A method for finding initial guesses for the solver is described in \autoref{sec:stokes}.
-\autoref{sec:implementation} contains a schematic of program and a description of its workflow.
+\textsf{SpecTraVVave} is given in section \ref{sec:spectral}.
+Section \ref{sec:experiments} presents results of different experiments carried out with the package.
+Concluding remarks are given in section \ref{sec:conclusion}.
+%A method for finding initial guesses for the solver is described in \ref{sec:stokes}. \ref{sec:implementation} contains a schematic of program and a description of its workflow.
 
@@
 \begin{equation}
-\SN = \mathrm{span}_{\R}\setc{\cos(lx)}{0 \leq l \leq N-1 }
+\SN = \mbox{span}_{\R}\{ cos(lx) | 0 \leq l \leq N-1 \}
 .
@@
 %This may lead to confusions for reviewers.}
-Approximation errors are summarized in \autoref{t2}.
+Approximation errors are summarized in \ref{t2}.
 %
@@
 %\includegraphics[width=0.47\textwidth]{convergence}      
-%\caption{\small Graph of error estimates given in \autoref{t2}.}
+%\caption{\small Graph of error estimates given in \autoref{Ben_disp_rel}ref{t2}.}
 %\label{L2convergence}
@@
 %%%%%%%%%===========%%%%%%%%%%%%%%%
-\begin{enumerate}[label=\alph*)] % chktex 9 chktex 10
-\item Where does the bifurcation curve terminate?
-
-\item Where on the bifurcation curve do solutions change their stability?
-
-\item Is there any role that the turning point on the bifurcation curve plays?
-\end{enumerate}
+\begin{itemize} % chktex 9 chktex 10
+\item[a)] Where does the bifurcation curve terminate?
+
+\item[b)] Where on the bifurcation curve do solutions change their stability?
+
+\item[c)] Is there any role that the turning point on the bifurcation curve plays?
+\end{itemize}
 %%%%%%%%%===========%%%%%%%%%%%%%%%
@@
 
-From the data given in the \autoref{t1}, one can deduce that for solutions with minimum speed 
+From the data given in the \ref{t1}, one can deduce that for solutions with minimum speed 
 and maximum $L^2$-norm exponential fit is better than polynomial. 

### 2017-02-19 Babenko/BabenkoEquation/lecture/pythonlisting.tex
Babenko folder added.

% Python listing setup

\usepackage{color}
\usepackage[procnames]{listings}
\usepackage{textcomp}
\usepackage{setspace}
\usepackage{palatino}
\renewcommand{\lstlistlistingname}{Code Listings}
\renewcommand{\lstlistingname}{Code Listing}
\definecolor{gray}{gray}{0.5}
\definecolor{green}{rgb}{0,0.5,0}
\definecolor{lightgreen}{rgb}{0,0.7,0}
\definecolor{purple}{rgb}{0.5,0,0.5}
\definecolor{darkred}{rgb}{0.5,0,0}
\lstnewenvironment{python}[1][]{
\lstset{
language=python,
basicstyle=\ttfamily\tiny\setstretch{1},
stringstyle=\color{green},
showstringspaces=false,
alsoletter={1234567890},
otherkeywords={\ , \}, \{},
keywordstyle=\color{blue},
emph={access,and,as,break,class,continue,def,del,elif,else,%
except,exec,finally,for,from,global,if,import,in,is,%
lambda,not,or,pass,print,raise,return,try,while,assert},
emphstyle=\color{orange}\bfseries,
emph={[2]self},
emphstyle=[2]\color{gray},
emph={[4]ArithmeticError,AssertionError,AttributeError,BaseException,%
DeprecationWarning,EOFError,Ellipsis,EnvironmentError,Exception,%
False,FloatingPointError,FutureWarning,GeneratorExit,IOError,%
ImportError,ImportWarning,IndentationError,IndexError,KeyError,%
KeyboardInterrupt,LookupError,MemoryError,NameError,None,%
NotImplemented,NotImplementedError,OSError,OverflowError,%
PendingDeprecationWarning,ReferenceError,RuntimeError,RuntimeWarning,%
StandardError,StopIteration,SyntaxError,SyntaxWarning,SystemError,%
SystemExit,TabError,True,TypeError,UnboundLocalError,UnicodeDecodeError,%
UnicodeEncodeError,UnicodeError,UnicodeTranslateError,UnicodeWarning,%
UserWarning,ValueError,Warning,ZeroDivisionError,abs,all,any,apply,%
basestring,bool,buffer,callable,chr,classmethod,cmp,coerce,compile,%
complex,copyright,credits,delattr,dict,dir,divmod,enumerate,eval,%
execfile,exit,file,filter,float,frozenset,getattr,globals,hasattr,%
hash,help,hex,id,input,int,intern,isinstance,issubclass,iter,len,%
license,list,locals,long,map,max,min,object,oct,open,ord,pow,property,%
quit,range,raw_input,reduce,reload,repr,reversed,round,set,setattr,%
slice,sorted,staticmethod,str,sum,super,tuple,type,unichr,unicode,%
vars,xrange,zip},
emphstyle=[4]\color{purple}\bfseries,
upquote=true,
morecomment=[s][\color{lightgreen}]{"""}{"""},
commentstyle=\color{red}\slshape,
literate={>>>}{\textbf{\textcolor{darkred}{>{>}>}}}3%
         {...}{{\textcolor{gray}{...}}}3,
procnamekeys={def,class},
procnamestyle=\color{blue}\textbf,
framexleftmargin=1mm, framextopmargin=1mm, frame=shadowbox,
rulesepcolor=\color{blue},#1
}}{}

### 2017-02-19 Babenko/BabenkoEquation/lecture/statement.tex
Babenko folder added.

\section{The Babenko equation}
\subsection{Statement of the problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The Babenko equation: statement of the problem}
Equations of the following form are considered:

\begin{equation}
u_t+ f(u)_x + \Lop u_x=0 
\end{equation}

where $f(\eta)$ is a non-linear flux function and $\mathcal{L}$ is a linear dispersion operator:

$$\widehat{\Lop u}(k)=\alpha(k)~\widehat{u}(k). $$

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%---------------------------------------------------------------------%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The Babenko equation: statement of the problem}
Equations of the following form are considered:

   \onslide<1-2>{The Whitham equation:

\begin{equation}
u_t+\frac{3}{2}u u_x+K\ast u_x=0. \label{Whitham}
\end{equation}

$$
\widehat{K\ast u}(k) = \sqrt{\frac{\tanh(k)}{k}}~\widehat{u}(k).
$$}
   \onslide<2-2>{The KdV equation:
\begin{equation}
u_t+\frac{3}{2}u u_x+ u_x + \frac{1}{6} u_{xxx}=0.
\end{equation}}
\end{frame}

### 2017-02-19 Babenko/BabenkoEquation/lecture/filename.py

@@ -1,0 +1,22 @@
+import time
+import numpy as np
+import matplotlib.pyplot as plt
+
+fig=plt.figure()
+plt.axis([0,100,0,1])
+
+i=0
+x=list()
+y=list()
+
+plt.ion()
+plt.show()
+
+while i <100:
+    temp_y=np.random.random()
+    x.append(i)
+    y.append(temp_y)
+    plt.scatter(i,temp_y)
+    i+=1
+    plt.draw()
+    time.sleep(0.05)

### 2017-02-19 Babenko/gsemthesis-master/dev/bundle-for-ctan.sh

@@ -1,0 +1,7 @@
+mkdir gsemthesis
+cp README.md gsemthesis/README
+cp src/gsemthesis.ins gsemthesis/
+cp src/gsemthesis.dtx gsemthesis/
+cp src/gsemthesis.pdf gsemthesis/
+zip -r gsemthesis.zip gsemthesis
+rm -r gsemthesis

### 2017-02-19 Babenko/gsemthesis-master/dev/update-test.sh

@@ -1,0 +1,29 @@
+classname=gsemthesis
+
+# generates new class files
+cd src/
+if [ -e $classname.cls ]
+then
+  rm $classname.cls
+fi
+latex $classname.ins
+pdflatex $classname.dtx
+cd ..
+
+# copy files into the /cls directory
+cp src/gsemthesis.cls cls/
+cp src/gsemthesis.pdf cls/
+
+# test example file
+cp src/gsemthesis.cls example/
+cd example/
+pdflatex phdthesis-example.tex # better to run it manually
+biber phdthesis-example
+cd ..
+
+# test example-minimal file
+cp src/gsemthesis.cls example-minimal/
+cd example-minimal/
+pdflatex phdthesis-example-minimal.tex # better to run it manually
+biber phdthesis-example-minimal
+cd ..

### 2017-03-26 Thesis/latex/outline.tex
New run

@@
 
-\noindent Henrik Kalisch, Daulet Moldabayev, Olivier Verdier: \textit{A numerical study of nonlinear dispersive wave models with SpecTraVVave}, \textcolor{red}{specify status of the paper.}
+\noindent Kalisch, H., Moldabayev, D., Verdier, O.: \textit{A numerical study of nonlinear dispersive wave models with SpecTraVVave}, Electronic Journal of Differential Equations~ \textbf{2017}, No.
+62, pp.
+1--23 (2017), http://ejde.math.txstate.edu or http://ejde.math.unt.edu.
 

### 2017-03-26 Thesis/latex/preface.tex
New run

@@
 
-The advisory committee has consisted of Henrik Kalisch (University of Bergen, Bergen, Norway),  Dmitry Pelinovsky (McMaster University, Hamilton, Ontario, Canada) and Angel Dur\'an (Universidad de Valladolid, Valladolid, Spain).
+The advisory committee has consisted of Henrik Kalisch and Guttorm Alendal (University of Bergen, Bergen, Norway),  Dmitry Pelinovsky (McMaster University, Hamilton, Ontario, Canada) and Angel Dur\'an (Universidad de Valladolid, Valladolid, Spain).
